### 2019-10-10 appendix.tex
Initial Overleaf Import

In this supplementary material we highlight the use cases that \noah
is suited for, the in-situ chart representations within
the overview, \noah’s underlying system architecture,
and statistical significance results from our user study.
\section*{\noah Use Cases}
As mentioned in Section~\ref{sec:design}, the goal of designing 
an overview interface is to 
not only enable seamless navigation 
but also to facilitate functionalities to help users visually seek information about the data. Therefore, we draw parallels to the typology of abstract visualization tasks~\cite{brehmer2013multi} to identify the use cases that \noah is best suited for. 
We are specifically interested in exploring the different task purposes mentioned in the typology. The task purposes can be specified at three different levels: consume and produce (high-level), search (mid-level), and query (low-level). Table~\ref{tab:scope} captures all three levels and highlights where \noah can be used.

At a very high level, users may utilize \noah to verify some phenomenon in the data (\code{discover}) or consume the data for casual interest (\code{enjoy}). Example of \code{discover} includes a journalist trying to verify whether listings in Airbnb are run as hotels (see Section~\ref{sec:usage}). 
Moreover, users may utilize the overview to communicate details about the data (\code{present}). 
However, visualization tools are more suitable for such a purpose. 
Finally, users can customize (\code{create}) the overview to fit their exploration purpose. Here, the high-level goal is to \emph{Produce} a new representation. 
However, the present version of \noah does not allow users to add annotations, \eg visual cues, texts, to the overview (\code{annotate}).

Regardless of the high-level purposes, users may want to find elements of interest in the overview (\emph{search}) and seek additional details about those elements (\emph{query}). 
In Section~\ref{sec:ui}, we discussed the features and interactions supported by \noah and showed how these interactions enable users to \emph{search} and \emph{query} the data. Users can \code{browse} and \code{explore} the data using navigational operations, \eg clicking and zooming, to find elements of interest. 
Note that if users know either the identity or location of the information they are searching (\eg the name of the city in the Airbnb data), 
they can simply use traditional spreadsheet operations like \emph{VLOOKUP} (\code{lookup}) or \emph{CTRL+F} (\code{locate}) (see Table~\ref{tab:scope}). 
On the other hand, clicking or zooming is suitable when the identity or location of the information is unknown. Therefore, NOAH complements spreadsheets in accomplishing users’ \emph{search} goal via navigational operations. 
Once the user finds the elements of interest, they can further \code{summarize} the elements via aggregate columns, \code{compare} those elements, and \code{identify} the raw data corresponding to the summaries in the spreadsheet.

\section*{Explaining the Chart Representations}

In this section, we explain the chart representation of all the
formula categories discussed in Section~\ref{sec:ui},
also displayed in Figure~\ref{fig:agg}.

\stitle{Summary.} The result of a \emph{summary} formula,
\eg \code{AVERAGE}, is depicted using a horizontal bar.
The bar represents
the range of the data subset the bin spans
with the minimum and maximum values
annotated within the chart. A vertical line is used to highlight
where the result lies within the range.

\stitle{Conditional.} The result of a \emph{conditional} formulae,
\eg \code{COUNTIF}, is depicted using a histogram.
The histogram captures the distribution of the attribute,
on which the formula has been applied,
\eg the \emph{availability} attribute discussed
in Section~\ref{sec:usage}.
Shading is used to de-emphasize data ranges
that do not satisfy the condition. A vertical line is used to highlight
where the result lies within the distribution.

\stitle{Frequency.} The result of a \emph{frequency} formula,
\eg \code{mode}, is also depicted using a histogram.
The bin in the histogram
that contains the result is rendered with ``orange'' color.
A vertical line is used to highlight
where the result lies within the distribution.

\stitle{Spread.} Finally, the result of a \emph{spread} formula,
\eg \code{mode}, is depicted using a histogram.
A similar shading technique as the \emph{conditional} formula is used to highlight the standard deviation. The mean of the is highlighted using a vertical line.

In the chart representation mode,
each entry of the aggregate column
contains one additional visual cue---a color bar
with shades of green on
the right of the chart (see Figure~\ref{fig:customize}).
The darker the color, the higher the value corresponding to that entry. 
Users can utilize the color intensity
to compare the results among different aggregate column entries.

\section*{Implementation and  Architecture}
\label{sec:sys}
In this section, we provide an overview of the infrastructure of \noah. 
We integrate \noah as a data exploration plugin within{\scshape DataSpread}~\cite{dataspread}, 
an open-source scalable web-based spreadsheet.

\subsection{Underlying Data Structures}
As explained in Section~\ref{sec:ui}, the underlying data structure representing the overview is an in-memory equi-depth histogram. 
\noah constructs the histogram on demand based on the navigation attribute. 
In the beginning, only the highest granularity bins are constructed. 
As users perform ad-hoc interactions on the data, the interface is updated on the fly. 
For example, when a user zooms into a specific bin, 
\noah again constructs an equi-depth histogram 
on the data corresponding to that bin on demand.
To enable seamless integration of the overview 
with the spreadsheet data, we leverage the hierarchical positional indexes
used by {\scshape DataSpread}~\cite{datamodels} to access
the spreadsheet data. 
The index is essentially an
order statistics tree~\cite{datamodels} 
built on the position (\eg row number) of the spreadsheet data. 
For any given navigation attribute, a new positional index is constructed first. 
\noah then leverages the positional mapping to access the underlying data corresponding to the navigation attribute and constructs the histogram depicting the overview. 
Each bin in the histogram maintains positional information regarding its elements, \ie starting and ending index of each unique element in the bin (\eg cities in the Airbnb data).  
Therefore, \noah can be integrated into any 
spreadsheet and requires only access to the positional mapping structure of that spreadsheet.

\subsection*{System Architecture}

\begin{figure}
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 4cm 0 3cm,clip,width=\linewidth]{images/arch.pdf}
   \caption{System architecture.}
   \label{fig:arch}
 \end{figure}

We now explain the system architecture of \noah. 
The \noah client is a web-based front-end that captures user input and renders both the navigation interface and the spreadsheet 
based on the results returned by the back-end. 
The front-end is responsible for capturing user input and rendering components of the navigation interface, \ie overview, aggregate column, and context bar. 
Given any interaction by the user on the front end, 
the \emph{request processor} issues a request to back end.
The back end navigation controller receives the request from the front end. 
After processing the request that corresponds to some front-end user interaction, 
the \emph{request processor} sends a response to the front-end encoded in \emph{json}. 
For requests involving the spreadsheet, \eg scrolling, 
the \emph{request processor} leverages the \emph{positional index} 
to access the spreadsheet data. 
For requests involving the navigation interface, \eg zoom in/out, 
the \emph{request processor} leverages the \emph{compute engine} to manage the equi-depth histogram on demand. 
The compute engine is also responsible for processing analytical operations. We leverage {\scshape DataSpread}’s built-in formula engine to support the analytical operations.

\section*{Results}
In this section, we evaluate the statistical significance of the task performance results and survey responses presented in Section~\ref{sec:results}. 
\begin{figure}
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 57 0 145,clip,width=\linewidth]{images/userinfo-ip-1.pdf}
   \caption{Intra-participant submission time differences between \noah and Excel across the quiz tasks.}
   \label{fig:ipdiff}
 \end{figure}

\subsection{Intra-participant differences.}
Figure~\ref{fig:ipdiff} depicts the intra-participant submission time differences between \noah and Excel, across all the quiz tasks.
For a given task and a participant, the corresponding circle denotes by what percentage (between $0.30\%$ to $94.33\%$) a system is faster than the competing system. The color of the circle denotes which system was faster (green: \noah faster, yellow: Excel faster). The larger the circle the faster the corresponding system is. For example, participant $P9$'s submission time was $91.78\%$ faster with \noah for the \cmpB task and $72.62\%$ faster with Excel for the \cmpA task. Figure~\ref{fig:ipdiff} confirms that majority of the submission times using \noah were faster compared to Excel---nineteen out of the twenty participants completed at least four tasks in less time using \noah compared to Excel. The submission time difference is even more apparent for the steer, identify, and \cmpB tasks.

For \cmpA, customize, and the second steer task, less than a third of the participants' submission times were faster using Excel. 
For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah, utilized the \emph{autosum} shortcut for the birdstrikes dataset. For the \cmpA tasks, a number of participants ($N=3$) that required more time to submit answer using \noah, first used the bin customization feature to view all the unique bins which contributed to the higher submission time compared to Excel. On the other hand, as explained earlier, the unfamiliarity with the bin customization feature contributed to higher submission time using \noah for six participants (yellow circles).

\subsection{Statistical Significance: Task Performance}
Since we conducted our user study on a small population (20 participants), we further evaluated the statistical significance of the task performance results, \ie accuracy and completion time. 
To measure the significance of the task completion times, we ran \emph{Mann-Whitney's U test} (as completion times did not follow a normal distribution). \emph{For all of the tasks except the customize task, we found a significant effect of the tools, \ie the response times for the tasks significantly differed by the choice of the tool} (see Table~\ref{tab:statTest}). 
On the other hand, we ran the \emph{Fisher's exact test} that measures the statistical significance of categorical data (0/1 accuracy): \emph{only for steer tasks, the percentage of the accuracy of submissions significantly differed by the choice of the tool}.

\begin{table}[!htb]{}
\vspace{-10pt}
\scriptsize
\caption{Statistical significance of submission time and accuracy comparisons between \noah and Excel. (*) indicates statistically significant.}
\label{tab:statTest}
\centering
\begin{tabular}{cccc}
\hline
Question & Category &  Time   &    Accuracy \\
& & ($p$ value) & ($p$ value) \\ \hline
$Q1$ & Steer & $\mathbf{0.0007}$ (*) &  $\mathbf{0.0033}$ (*)  \\
$Q2$ & Identify &   $\mathbf{2.49 \times 10^{-5}}$ (*) & $0.7475$ \\  
$Q3$ & Steer & $\mathbf{0.0043}$ (*) & $\mathbf{0.0202}$ (*) \\
$Q4$  & \cmpA & $\mathbf{0.0154}$ (*) & $1$ \\  
$Q5$ & \cmpB & $\mathbf{5.83 \times 10^{-6}}$ (*) & $0.48$\\
$Q6$ & Customize & $0.1207$ &    $0.0959$ \\\hline
\end{tabular}
\vspace{-10pt}
\end{table}

\subsection{Statistical Significance: Subjective Ratings}
We further conducted a statistical significance test---the \emph{Wilcoxon Signed-rank} test---on the survey responses which showed that for all the metrics, the ratings significantly differed by the choice of the tool, \ie \noah or Excel. The distribution of the ratings for none of the criteria followed a normal distribution.

\begin{table}[!htb]{\scriptsize}
\caption{Survey results. (*) indicates statistical significance. }
\label{tab:survey}
\centering
\begin{tabular}{|l|l|l|l|}
\hline
Metric &    \noah   &    Excel   & $p$ value \\ \hline
Ease of Learning &  $\mu = 5.75$, &    $\mu = 4.22$, & $\mathbf{1.49 \times 10^{-7}}$ (*) \\
& $\sigma = 1.02$ & $\sigma = 1.41$ & \\\hline
Speed of Use &    $\mu = 6.03$, & $\mu = 4.22$, & $\mathbf{1.68 \times 10^{-7}}$ (*)\\
& $\sigma = 0.99$ & $\sigma = 1.65$ & \\\hline  
Ease of Use    & $\mu = 5.88$, & $\mu = 4.33$, & $\mathbf{7.85 \times 10^{-6}}$ (*) \\
& $\sigma = 0.90$ & $\sigma = 1.71$ & \\\hline
Confidence    & $\mu = 5.50$, & $\mu = 4.60$, & $\mathbf{0.0096}$ (*)\\
& $\sigma = 1.79$ & $\sigma = 1.50$ & \\\hline  
Comprehensibility & $\mu = 5.60$, & $\mu = 4.48$, & $\mathbf{0.0006}$ (*)\\
& $\sigma = 1.27$ & $\sigma = 1.65$ & \\\hline
Satisfaction & $\mu = 5.48$, &    $\mu = 4.52$, & $\mathbf{0.0018}$ (*)\\
& $\sigma = 1.16$ & $\sigma = 1.49$ & \\\hline
\end{tabular}
\end{table}

### 2019-10-10 design.tex
Initial Overleaf Import

\section{\noah: Design Considerations}
\label{sec:design}
In this section,
we outline our design considerations
for a spreadsheet navigation interface.
%In our supplementary material,
%we additionally characterize
%the use-cases that such a navigation %interface
%is good for and not good for.
Our design considerations were
informed by prior work on
information visualization~\cite{shneiderman2003eyes,brehmer2013multi}, overview+detail interfaces~\cite{cockburn2009review}, multiple-coordinated views~\cite{wang2000guidelines},
and refined through our experiences
across multiple design iterations.

\para{DC1. Construct the overview in-situ}
An overview helps users get a high-level picture
of the data.
However, maintaining the overview in a separate
location from the data can be cumbersome;
instead, having it co-located with the data
can help users make rapid glances to
explore information between a
bird's-eye view
and a close-up detail~\cite{grudin2001partitioning}.

\para{DC2. Ensure reduced
visual discontinuity while providing details on demand}
Users often need to access subsets of data,
and study their properties in detail (\eg via steering);
navigating back and forth \saj{between different subsets of data}
can lead to increased visual discontinuity.
The interface should allow users to compute such details for
various data subsets on demand~\cite{shneiderman2003eyes}.
The interface should maintain visual continuity
as users navigate to a different subset,
recomputing the details for the new subset.

\para{DC3. Balance the screen space afforded to the overview}
As the overview
has limited screen-space available,
we need to consider the trade-off between visual discontinuity (\emph{DC2}) and clarity.
\saj{Displaying a fine-grained overview improves visual clarity while increasing visual discontinuity---users need to scroll through the overview to access distant
subsets of data.
Displaying a coarse-grained overview decreases visual discontinuity at cost of reduced visual clarity---the overview may span too many data subsets and appear visually cluttered}.
The interface should further allow users to control the screen-space allocated
to the overview.

\para{DC4. Enable coordination between the spreadsheet and overview}
Since users can view the overview
and the spreadsheet simultaneously,
interactions on both need to be linked~\cite{roberts2007state},
\ie an interaction on one should be reflected on
the other interface~\cite{wang2000guidelines}.
For example, as a user scrolls through
the spreadsheet, the user's current focus
should be highlighted on the overview.
However, not all interactions need to be interlinked,
\eg changing the font size of a spreadsheet cell
need not lead to a change in the overview.

\para{DC5. Facilitate customization of the overview}
As the overview is automatically generated,
it may not reflect domain-specific context
known only to the user~\cite{Raman99scalablespreadsheets}.
For example, an overview constructed on a grading
spreadsheet by binning nearby scores
may not match the letter grade ranges
that the instructors have in mind.
Allowing users to customize the overview is therefore essential.

\para{DC6. Display contextual and historical information regarding users' navigation}
The interface should record navigation history,
allowing users to revisit previously visited locations~\cite{shneiderman2003eyes},
while also displaying their current navigation path for context.

\section{User Interface}
\label{sec:ui}
We now explain the design of \noah's components
and implementation details.

\subsection{In-situ Overview}

\noah constructs the overview in-situ ({\bf DC1})
next to the spreadsheet
on an attribute of the spreadsheet dataset
called the {\em navigation attribute},
selected by the user.
Any attribute type that can be ordered
can be a navigation attribute, \eg
text, numbers.
The overview is constructed at multiple granularities.
Each granularity is divided into non-overlapping
groups of data called {\em bins}.
As shown in Figure~\ref{fig:concept}d,
an overview of the Airbnb data
on the navigation attribute ``city''
has two levels of granularities.
The highest level of the granularity
consists of four bins.
Figure~\ref{fig:concept}a
depicts \noah showing the first four bins,
the first of which is {\em Ashville-Boston}.
Each bin contains summary information
regarding the data subset/region
it spans, \eg starting row and
ending row number,
the total number of rows the region spans.
Each bin displays an overview
of the next (finer) granularity (if any)
with embedded bar charts.
For example, in Figure~\ref{fig:concept}d,
the topmost bin ({\em Ash-Bos}) spans three cities
({\em Ashville}, {\em Austin}, {\em Boston}), each of which is a bin
in the next (finer) granularity.
Correspondingly, Figure~\ref{fig:concept}(a)
shows three horizontal bar charts for the
first {\em Ash-Bos} bin,
one for each bin in the next granularity.
On the other hand,
since the third bin from the top
(\emph{LA}) spans only one city,
no bar chart is embedded.
Users can perform different
operations on the bins, \eg clicking to pan and
semantic zooming in/out~\cite{perlin1993pad}.
\noah supports other interactions,
\eg customization and aggregation.
We discuss these interactions
in the context of our design considerations in Section~\ref{sec:overview_operations}.

\stitle{Why a Multi-granularity Binned Overview?}
\saj{A standard overview approach
is to show a spatially partitioned
collection of thumbnails \saj{on the left of the detailed view},
similar to Microsoft PowerPoint or
Adobe Reader.
However, displaying too many thumbnails
results in increased scrolling
to access distant thumbnails,
increasing visual discontinuity.
On the other hand,
displaying too few thumbnails
reduces visual discontinuity,
but at the cost of visual
clarity---the thumbnails appear cluttered
and fail to represent the
underlying data clearly~\cite{cockburn2009review}.
To strike a balance between these two objectives
({\bf DC3}) we designed a multi-granularity overview
that abstracts the data at varying levels of detail.
Multi-granularity interfaces
provide an alternative to a single-granularity
spatial partition of the data space,
by varying the scale at which the data is displayed,
enabling multiple views at varying granularities~\cite{cockburn2009review}.
Users can resize the overview to control the amount of spreadsheet data that remains visible.
\toappendix{Microsoft PowerPoint also uses a
similar technique to balance the screen
space between overview and detail.}
Moreover, users can also hide the overview if required.}

\saj{As explained earlier,
we divide each granularity
of the overview into bins.
The highest (coarsest) level of the granularity
provides a global view with less visual clarity
and reduced visual discontinuity,
As the bins become more fine-grained, visual clarity improves while visual discontinuity increases.
The binned representation of the overview
using equi-depth histograms
enables faster navigation
by reducing visual discontinuity. 
The bins of the histogram
act as landmarks
in the overview, enabling 
users to skip irrelevant bins
and quickly navigate to the desired subset of data.
The equi-depth histogram 
can be constructed on 
any data types
that can be ordered, \eg text, numbers, dates. For example, 
in the usage scenario explained in Section~\ref{sec:usage},
the journalist grouped the data into cities
for ease of navigation 
when exploring the larger cities
in the Airbnb dataset.}

\stitle{Overview Construction.}
The data structure underlying the overview
is an equi-depth histogram constructed
on the values
in the navigation attribute column.
Equi-depth histograms are commonly
used for summarizing the statistical properties
of data, with applications such as
database query optimizers~\cite{chaudhuri1998random}.
Each bin in the equi-depth histogram
contains the same number of items,
where each item is a value.
For example,
when constructing the overview
on city,
each value in the city column
is assigned to a bin.
The bins are constructed top-down
(see Figure~\ref{fig:concept}d). \noah divides
each of the bins at level $k$ into new bins
to construct the next lower level $k+1$, again,
by applying the same concept of equi-depth histograms.
If each value of the navigation attribute
column was unique, \eg if it was a numerical ID,
then construction of the histogram would be easy:
each bin of the equi-depth histogram
would contain almost the same number of items,
where each item corresponds to one unique value of
the attribute.
Unfortunately, in practice, for many attributes,
the same value is often repeated.
For example,
there are multiple listings
per city.
Therefore,
an equi-depth histogram
on the
attribute city
will result in consecutive bins
sharing items of
the same unique city value,
resulting in undesirable overlap.
Instead,
we construct a best effort
equi-depth
histogram that is as close to an equi-depth
histogram as possible, while ensuring
that the ranges represented by each bin have no overlap.
\toappendix{\agp{We don't have time but for the future, this is
something that would have been great to describe the algorithm on
for technical depth.}}

\begin{figure}[t]
        \centering
        %\vspace{-10pt} %trim=left bottom right top
        \includegraphics[width=0.45\textwidth,trim={0 0 400 0},clip]{images/navigationOp.pdf}
\vspace{-10pt}
   \caption{Navigational operations. (a) The overview at the highest level of granularity. (b) A zoomed in view of the \emph{Ashville-Boston} bin. (c) As the user clicks on the \emph{Boston} bin, the Boston listings are displayed on the sheet. The \emph{Boston} bin is highlighted in gray to indicate user’s current focus. (d) Conceptualizing the multi-granularity overview. }
\vspace{-18pt}
   \label{fig:concept}
 \end{figure}

\subsubsection{Operations and Interactions}
\label{sec:overview_operations}
We now discuss the operations and interactions that can be performed on the overview.

\stitle{Navigational Operation: Clicking.}
When a user clicks on a specific bin,
\noah displays the corresponding spreadsheet data;
users can use this to jump to a specific spreadsheet location
without having to scroll endlessly.
For example, in Figure~\ref{fig:concept}b,
as the user clicks on the \emph{Boston} bin,
the data corresponding to Boston is displayed
(Figure~\ref{fig:concept}c).
Note that the click operation is different from the traditional
spreadsheet \code{Filter} operation.
\code{Filter} hides spreadsheet data
that do not satisfy the filtering condition while
clicking brings the desired subset of data
in view without hiding the rest.
Users are free to navigate to other portions
through scrolling
even after clicking a bin, unlike filtering,
where users need to issue another \code{Filter}
to bring other data into view.

\stitle{Navigational Operation: Semantic Zooming.}  
Users can zoom into a specific bin
to view more fine-grained information or zoom out
to view more coarse-grained information,
via semantic zooming~\cite{perlin1993pad}.
For example, in Figure~\ref{fig:concept}a,
from the bin \emph{Ashville-Boston}
when the user zooms in to the next level,
\noah displays the bins \emph{Ashville}, \emph{Austin},
and \emph{Boston} (Figure~\ref{fig:concept}b).
If the user zooms out of the current granularity, again \noah displays the bins \emph{Ashville-Boston}, \emph{Chicago-Denver}, and others.
Users can only zoom into any bin that
contains multiple unique values.
For example, in Figure~\ref{fig:concept}d, at level 2,
each bin corresponds to one city.
Therefore, users can only click on those bins
to bring that data into view, and cannot zoom in further.
One issue with zooming interactions
is discoverability \saj{of the zoom operation}~\cite{cockburn2009review}.
We circumvent this (see Figure~\ref{fig:concept}c)
by providing the root of the bin under selection
for zoom out, and arrows for
clicking to zoom in (``\textcolor{blue}{$\rangle$}'') and out (``\textcolor{blue}{$\langle$}'').

\stitle{Customizing the Overview.}
As \noah constructs the overview automatically,
the overview binning or organization
may not capture domain-specific context
or user needs.
\noah enables users to customize\toappendix{ (\code{create} in Table~\ref{tab:scope})} this organization (\textbf{DC5}).
At any granularity, users can merge multiple consecutive bins
into a single bin, or split a bin into multiple bins.
\toappendix{The other operations include merging all the bins into one single bin and collapsing all the bins into singular bins, \ie one unique value per bin.}
Say the user wants to compare summary statistics
of Boston and Chicago.
In the current organization
these two cities are  
in two different bins (see Figure~\ref{fig:customize}a).
Using the bin customization feature,
the user can merge the two bins
\emph{Ashville-Boston} and \emph{Chicago-Denver} to create a new bin \emph{Ashville-Denver}.
Users can now zoom into this bin and
compare summary statistics of the cities in the same view.
The interactions for splitting a
bin depend on the data type.
If the navigation attribute is textual,
any bin can be split into as many bins as the
number of unique values that bin contains.
If the navigation attribute is numeric,
users can split the bin into any arbitrary number of bins.
Note that \noah does not allow users to rearrange the
order of the bins.\toappendix{ Users can only customize the boundary of the bins.} Since the overview represents a histogram, the bins are ordered---reshuffling the bins violates that order.

\begin{figure}[!htb]
        \centering
        \vspace{-8pt}
        \includegraphics[width=0.45\textwidth,trim={0 0 130 0},clip]{images/customize.pdf}
\vspace{-10pt}
   \caption{(a) Chart view of the aggregate column. (b) A new bin is created by merging the top two bins. (c) Zooming into the newly created bin. }
\vspace{-10pt}
   \label{fig:customize}
 \end{figure}

\subsubsection{Coordination Between Overview and Spreadsheet}
\noah supports coordination
between the overview and the
corresponding spreadsheet data (\textbf{DC4}),
\ie interaction on the overview
may be reflected on the spreadsheet and vice-versa.
Indicating the navigation attribute on the spreadsheet
using color (see the lime green column in Figure~\ref{fig:ux}c)
as user constructs the overview, is an example of such coordination. However, not all overview interactions are coupled
with the spreadsheet and vice versa.
The coupling depends on the user's current
focus---\emph{\saj{to ensure consistency between the overview and the spreadsheet,} any interaction on
either interface
that changes the current focus must be reflected
on the other interface}.
We now provide examples of both
coupled and decoupled interactions.

\stitle{Coupled interactions.}
Clicking a bin is an example of a coupled interaction
as the user actively changes the focus
to another bin on the overview.
To reflect the change, \noah populates
the corresponding spreadsheet data
on the screen.
As the user scrolls on the spreadsheet,
again the current focus changes
and the corresponding bin on the overview is highlighted.
For example, in Figure~\ref{fig:concept}c,
as the user clicks on the Boston bin,
the spreadsheet displays the Boston listings.
Conversely, as the user scrolls up,
both Austin and Boston listings appear in the current window of the spreadsheet.
Therefore, both the Austin and Boston overview
bins are highlighted
(see Figure~\ref{fig:concept}d).
%The zoom in operation, on the other hand,
%can be either coupled or decoupled based on the setting.
%Whenever a user zooms into a bin
%that is not the current focus,
%the spreadsheet is updated to show data
%for that bin.

\stitle{Decoupled interactions.}
When a user zooms into a bin
that is already in the user's current focus,
the spreadsheet view does not change.
For example, in Figure~\ref{fig:concept}a
the user zooms into the \emph{Ashville-Boston} bin;
here, the spreadsheet view stays the same
(see Figure~\ref{fig:concept}b).
Similarly, the zoom out operation is decoupled.
When a user zooms out,
the overview displays
a coarser granularity view
of the user's current focus.
Since the focus stays the same,
there's no need to update the spreadsheet view.
Similarly, operations like panning on the overview without clicking
and customizing the overview
do not change user's current focus and are decoupled.
Online maps also adopt similar
decoupling of the overview and detail~\cite{cockburn2009review}.
However, their goal is to reduce network
and computational overload, whereas in our case,
the decoupling is based on the user's current focus.

\subsection{Aggregate Columns}
Users can issue spreadsheet formulae
on the overview
to compute aggregates for the data
in each bin\toappendix{ (\code{summarize} in Table~\ref{tab:scope})}.
The results
are displayed as an {\em aggregate column} (see Figure~\ref{fig:ux}b). Each entry in the aggregate column
corresponds to the adjacent bin
in the current granularity
of the overview.
For example, in Figure~\ref{fig:customize}c,
the aggregate column displays
four aggregate statistics,
one per bin.
Users can issue several formulae simultaneously,
each giving rise to a new aggregate column.
However, adding an aggregate column
takes up screen space thus shrinking the spreadsheet view.
As a workaround, users can resize or
remove aggregate columns
to make more screen space available
to the spreadsheet (\textbf{DC3}).
When user issues a formula on the overview,
the spreadsheet column
corresponding to the aggregate column
is highlighted in grayish orange
(see Figure~\ref{fig:ux}c)---another example of coupled interaction.
For conditional formulae like \code{COUNTIF},
cells that satisfy the condition
are highlighted, \eg in Figure~\ref{fig:ux}c),
the cells with availability $\ge 60$ are colored in sky blue.
By viewing the colored cells,
users can find out which cells
are relevant to the aggregation operation\toappendix{ (\code{identify} in Table~\ref{tab:scope})}.

Creating an aggregate column on the overview mimics how users create pivot tables. Users
are not required to explicitly type formulae;
rather they simply select the formula from a drop-down menu,
and provide the necessary formula parameters to a web form.
This way, \noah maintains the semantics
of existing spreadsheet tools.
Currently, \noah supports statistical
and mathematical formulae that
operate over a range of data.
Therefore, creating an aggregate column
is equivalent to selecting a subset of data
on the sheet, \ie steering and then executing a formula,
helping users avoid cumbersome steering operations.
We have classified the formulae supported
into five categories: a) summary (\eg min, max, average),
b) frequency (\eg mode, large, small), c) conditional (\eg countif, sumif), d) spread (\eg var,stdev), and e) others (\eg sum, count).

Users can choose to view the result of formulae
either as raw values or charts and can toggle between the two.
Raw values are displayed
along with a colored bar, the \emph{value bar},
whose length is proportional to
the corresponding aggregate (see Figure~\ref{fig:ux}b). Users can use the lengths to visually compare across bins\toappendix{ (\code{compare} in Table~\ref{tab:scope})}.
The chart representation varies depending on the formula type.
Except for the \emph{others} category
all other categories can be represented by charts.
Figure~\ref{fig:agg} shows the
chart representation for these categories
along with different visual cues
that highlight formula results
as well as other information. We discuss these representations 
in detail in the supplementary material.
\toappendix{For example, for \emph{conditional} formulae
like \code{COUNTIF},
shading is used to de-emphasize data ranges
that do not satisfy the condition.
A similar shading technique is used for
the \emph{spread} category to highlight the standard deviation.
For the \emph{frequency} category,
coloring is used to identify the bin
with the result, \eg mode.
In the chart representation mode,
each entry of the aggregate column
contains one additional visual cue---a color bar
with shades of green on
the right of the chart (see Figure~\ref{fig:customize}).
The darker the color, the higher the value corresponding to that entry. Users can utilize the color intensity
to compare the results among different aggregate column entries.}

Finally, we note that the aggregate column
is kept in sync with the bins
as users zoom in and out,
eliminating repeated steering operations.
\noah does not maintain
any additional data structure for the aggregate column.
The histogram underlying the overview
records the result of the aggregate column entries
corresponding to the bins.
Next, we discuss how \noah maintains user's navigational context.

\begin{figure}[t]
%\vspace{-10pt}
{
\scriptsize
   \centering
\begin{tabular}{l|l}
   \includegraphics[width=0.20\textwidth]{images/summary.png} &
   \includegraphics[width=0.20\textwidth]{images/frequency.png} \\
   \textbf{Chart Type:} Summary &  \textbf{Chart Type:} Frequency \\
 \textbf{Representation:} Horizontal bar &  \textbf{Representation:} Histogram \\
 \textbf{Visual Cue:} Vertical line (result) & 
 \textbf{Visual Cue:} Color (highlight) \\ \hline
    & \\
    \includegraphics[width=0.20\textwidth]{images/conditional.png}  &
   \includegraphics[width=0.20\textwidth]{images/spread.png} \\
      \textbf{Chart Type:} Conditional  &  \textbf{Chart Type:} Spread \\
 \textbf{Representation:} Histogram &  \textbf{Representation:} Histogram \\
 \textbf{Visual Cue:} Vertical line (result), & \textbf{Visual Cue:} Vertical line (mean),\\
Shade (de-emphasize) &  Shade (variance)
\vspace{-5pt}
\end{tabular}
}
\caption{Formula types and their chart representations.}
 \vspace{-15pt}
   \label{fig:agg}

\end{figure}
\subsection{Context Bar}
The context bar consists of two components:
a) a breadcrumb, and b) a navigation history.
The breadcrumb~\cite{breadcrumb} displays the current navigation path (see Figure~\ref{fig:ux}e), thus maintaining the users' navigation context (\textbf{DC6}). Each component of the breadcrumb corresponds to a bin in the user's current navigation path. Therefore, users can visit any bins within the current navigation path by clicking on an appropriate component of the breadcrumb, without having to zoom in or zoom out. \noah also maintains a list of recently visited bins (\textbf{DC6})
(see Figure~\ref{fig:ux}d). \toappendix{If aggregate columns are displayed along with the overview, contents of those columns are updated accordingly (\textbf{DC2}), as users revisit a bin.}

\subsection{Implementation}
We have integrated \noah with {\scshape DataSpread}~\cite{datamodels}, a web-based spreadsheet. The {\scshape DataSpread} back-end maintains the histogram data structure and  supports the aggregate column computation via its built-in formula engine.\toappendix{ To enable seamless integration of the overview with the spreadsheet data, we leverage the hierarchical positional indexes used by {\scshape DataSpread} to access
the spreadsheet data. The index is essentially an order statistics tree~\cite{datamodels} built on the position (\eg row number) of the spreadsheet data. For any given navigation attribute, a new positional index is constructed first. \noah then leverages the positional mapping to access the underlying data corresponding to the navigation attribute and constructs the histogram depicting the overview. Therefore, \noah can be integrated to any spreadsheet and requires only access to the positional mapping structure of that spreadsheet.} The \noah front-end is built with HTML/CSS/JS technologies along with the D3 framework~\cite{d3} for generating charts.

### 2019-10-10 discussion.tex
Initial Overleaf Import

\section{Discussion and Conclusion}
\label{sec:discuss}
\saj{In this paper, we have introduced \noah, an in-situ navigation interface, designed as a spreadsheet plugin.
Using \noah, users can get a bird’s-eye view of the data, with the ability to scroll or seek
additional details on demand, using a multi-granularity overview, as well as
aggregate columns that eliminate cumbersome steering operations. Quantitatively, we find \noah to speed up navigation without compromising the accuracy of the tasks. Qualitatively, study participants identify it as positively impacting their experience while overviewing and navigating large datasets, and issuing formulae. 
However, participants identified unfamiliarity and lack of transparency of operations as some of the limitations of \noah. In this section, we discuss
how to address these limitations
and highlight other future enhancement opportunities for \noah.}

\stitle{Transparency and documentation.} \saj{Several operations in \noah 
are quite different from typical spreadsheet interactions, \eg zooming, bin customization. Moreover, participants found some of the terminologies quite different from typical spreadsheet terminologies (see Section~\ref{sec:rq2}). In addition, some participants complained about the lack of explanation surrounding the overview construction and aggregate column computation. In the future, these issues can be addressed by using more relatable terminologies and improved documentation.}

\stitle{Maintaining spreadsheet look and feel.}
\saj{In subsequent versions of \noah, we can further 
display the appropriate formula for each bin as users hover 
over the corresponding cell on the aggregate column. 
Moreover, participants ($N=5$) noted the fact that \noah currently 
does not support user defined formulae, another possible future enhancement. Bin customization in \noah is performed using a menu bar which adds an additional step. In Excel, the cell splitting and merging operations are direct and only requires a single click. Similar direct adjustment of data grouping strategies have been explored for visualization tools~\cite{sarvghad2018embedded} and can be introduced in a future version of \noah.}

\stitle{Binning for different data types.} \saj{The experience 
surrounding the construction
of the overview
can be further improved, specially for categorical data.
Currently, the bins of the overview
can be customized
only after the overview is constructed.
Providing the users
the capability to
select the representation
(similar to bin customization)
of the overview could have addressed this issue.
Understanding the impact of these
representation choices for the overview
can be an interesting future research.}

\stitle{Scope of overview-spreadsheet coordination.} 
\saj{Spreadsheet users may perform various edit operations, 
\eg updating values, adding/deleting rows/columns. 
However, \noah currently assumes the data to be read-only. 
In our next version, we can add support for
propagating the spreadsheet updates to the overview. 
Moreover, the charts displayed in an \emph{aggregate column} are non-interactive, 
\ie users cannot interact with the charts to visually look up relevant or 
interesting data points within the spreadsheet. 
In the future, we plan to extend \noah to support visual querying 
through the charts in an aggregate column, similar to multi-modal linked selections in Keshif~\cite{yalccin2018keshif}.}

\stitle{Enhancing the navigation experience.} \noah currently constructs the overview on a single attribute. We can add support for multi-attribute navigation (\eg explore the Airbnb data by city and neighborhood), and multi-level navigation (\eg explore the neighborhoods after zooming into a specific city in the Airbnb data). Furthermore, bin customization currently supports changing the bin boundaries only while maintaining the current order. Supporting user defined ordering to allow arbitrary reshuffling of the bins can be another enhancement.\toappendix{ However, such modification will require a redesign of the histogram based data structure since the ordering property will not hold anymore.Finally, a complete characterization of the spreadsheet operations that can be supported by \noah is another open question, for example, other operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting, can also be supported by \noah.}

\stitle{Expanding the scope of supported operations.} \saj{Spreadsheet operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting, can also be supported by \noah. Therefore, it is necessary to  completely characterize the scope of spreadsheet operations that can be supported by \noah. Other enhancement includes adding annotations, \eg visual cues, texts, to the overview and then exporting the customized overview---a required feature for information seeking tools~\cite{shneiderman2003eyes}.}

\stitle{Beyond Tabular Data.} \noah operates only on tabular data.
However, spreadsheets can be semi-structured---formulae 
and text can be interspersed with tabular data. 
We can extend \noah to act as a map highlighting heterogeneous regions on such complex spreadsheets which users can utilize to navigate the spreadsheet. 
We can leverage existing work on spreadsheet table detection~\cite{tablesense}, 
and property identification~\cite{chen2017spreadsheet} 
to construct the map.

\stitle{From Perceptual to Interactive Scalability.} 
The current version of \noah addresses the perceptual 
scalability challenges while navigating 
Excel-scale (one million rows) data. 
As modern spreadsheets continue to support 
increasingly larger datasets---{\scshape DataSpread}~\cite{datamodels} 
supports one billion rows---the interactions proposed in this paper may violate the interactive response time bound of 500 ms~\cite{liu2014effects}. This opens the door to a new set of research challenges that may range from approximate query processing to progressive data analytics.

\noah represents our 
first step towards a general purpose spreadsheet navigation plug-in that can make spreadsheets more effective
on large datasets that are increasingly the norm.

### 2019-10-10 introduction.tex
Initial Overleaf Import

%\IEEEraisesectionheading{\section{Introduction}}
\section{Introduction}
\label{sec:intro}
With a user base of more than one-tenth of the world's population,
spreadsheets are by far the most popular medium for ad-hoc exploration
and analysis of data by end-users~\cite{excel-users}. \saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions with more advanced analytical features~\cite{chan1996use,raden2005shedding}. One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.}
Spreadsheets enable users to view, structure, and present data
in an intuitive tabular layout, wherein users can map their data and tasks;
this tabular layout is essential to
the popularity of spreadsheets~\cite{nardi1990spreadsheet}.

Using this tabular layout effectively involves
navigation, \ie
{\em ``the process of viewing and manipulating the computer display to
show another portion of the information space''}~\cite{nav-study}.
Navigation is supported
via two unit operations, scrolling and steering.
\saj{Scrolling is the action of moving displayed text or graphics up, down, or across on a computer screen in order to view different parts of the spreadsheet.}
For example, when analyzing data, users may scroll
to compare data across different screens, or to get
a high-level view of the overall spreadsheet.
\saj{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
users may steer to select
the subset of the data
to be operated on as an argument within the formula.
Most frequently used spreadsheet formulae
require users to perform steering actions~\cite{bradbard2014spreadsheet, lawson2009comparison}.
Overall, both scrolling and steering are crucial as
users navigate spreadsheets to
identify, compare, and summarize data.

\begin{figure*}
     \centering %trim=left bottom right top
  %\includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/navigation.pdf}
  \includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/screenshot.pdf}
    \caption{\noah: navigation interface consisting of (a) a zoomable overview and (b) an aggregate column integrated with (c) a spreadsheet. A context bar consisting of (d) a navigation history displaying locations visited so far using the overview, and (e) a breadcrumb showing the current navigation path (\eg Home). (f) The user’s current focus in the spreadsheet is highlighted on the overview. (g) Columns corresponding to the navigation attribute (city) and aggregate column (availability) are highlighted on the spreadsheet.}
  \label{fig:ux}
\end{figure*}

However, navigating spreadsheet data using scrolling or steering
is often challenging, even for moderately sized spreadsheets that only span
a few screens~\cite{network-context,nardi1990spreadsheet}, due to multiple inter-related reasons:
\squishlist
\item {\em Loss of overview and context.}
When navigating
spreadsheets, users can easily lose the context
of where they are and where they should go next~\cite{network-context}.
The only navigational context provided by spreadsheets
is the built-in scrollbar that acts as a one-dimensional
overview and indicates the user's current location on the sheet.
However, since this overview does not capture the
layout and structure of the data,
users are forced to mentally assimilate
the layout and recall it on-demand, as they navigate a spreadsheet.

\item {\em Cognitive and mechanical burdens.}
The lack of contextual cues leads to severe
cognitive and mechanical burdens
for users~\cite{cockburn2009review}.
Users often end up taking their own drastic measures to
avoid getting lost; for example,
some users create personalized overviews
extrinsic to the spreadsheet,
by sketching maps
of spreadsheets on paper~\cite{network-context}.
Other users add their own landmarks such as headers or
colored cells, as a visual affordance to assist
in navigation~\cite{network-context}.
Steering via dragging the mouse pointer across multiple screens to select
a subset of data as input to a formula
can often be challenging as well:
the only remedy is for users to abandon steering entirely and
instead remember the range of
the subset of data of interest, and then correctly enter this range
as the argument to the formula, often giving rise to errors that are increasingly prevalent in spreadsheets~\cite{panko1998we}.

\item {\em Visual discontinuities.}
The limited viewport afforded to the user
introduces a visual discontinuity between the
information being displayed.
For example, comparing spatially separated subsets
of data within the spreadsheet
requires moving back and forth between multiple viewports,
which can be overwhelming~\cite{nardi1990spreadsheet,network-context}.
As an alternative, users tend to copy subsets of data side by side to
reduce the visual discontinuity~\cite{nardi1990spreadsheet,network-context},
which is cumbersome.
\squishend

Overall,  while navigating present-day spreadsheets, {\bf \em users often lose context,
get overwhelmed, and experience visual discontinuities.} 
Addressing these challenges requires considerable manual effort.
As we argue in Section~\ref{sec:related},
existing spreadsheet functionalities such as
pivot tables, named ranges, and subtotals,
while providing partial support for some of these challenges
(\eg pivot table generates a summary
while losing the mapping between the
raw data and the summary; named ranges requires users
to manually associate names with ranges of data),
do not eliminate them entirely.
With the ease of data generation, and with spreadsheets
now supporting data
of increasingly larger scales---\eg Google Sheets
now supports two million cells, a five-fold increase
from the previous limit of 400K cells---navigating
data within spreadsheets will become harder, not easier.

So, how do we support more effective navigation of data
within spreadsheets?
One approach would be to try to integrate
an overview of the overall structure of the data along
with the spreadsheet~\cite{grudin2001partitioning}.
The resulting integration would then be a classical
overview+detail interface
where the spreadsheet is the detailed view.
Overview+detail interfaces are used
to facilitate navigation in various domains such as 
text editors and maps~\cite{cockburn2009review},
and enable users to manipulate
the overview or detailed view,
to perform high-level or low-level operations,
respectively.
Overview+detail interfaces
have been shown to be effective in these domains,
reducing cognitive load for users by
providing the big picture first,
helping them quickly assimilate the information space~\cite{cockburn2009review}.
\saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}

However, while an overview plug-in for spreadsheets
does seem appealing and natural,
developing it leads to several challenges.

\squishlist
\item {\em Overview modality.}
\saj{One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software applications like Microsoft PowerPoint or text editors
like Sublime Text. 
Unfortunately, this approach
would not suffice for a spreadsheet---while it would help users make
somewhat more informed
scrolling decisions, the zoomed out presentation of the overview would be unreadable. An overview should provide a comprehensible big picture view of the spreadsheet.
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a global context of the user's current location
being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users sought more fine-grained or coarse-grained view of the overall structure of the data.}

\item {\em Construction of the overview.}
\saj{Given a spreadsheet
with many rows, 
one approach to constructing a dynamic overview is mapping rows of data to high level groups similar to online maps. In online maps, cities are grouped into states and states are grouped into countries forming a multi-granularity hierarchy. How do we automatically group spreadsheet rows together in a similar ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers? If the automatic grouping is not meaningful, how do we allow the users
to customize the overview
to construct meaningful groups? How do we facilitate interactions that enable users view the overview at multiple granularities?}

\item {\em Operations on the overview.}
\saj{Following the construction of a dynamic overview, the next challenge is to design simple interactions that achieve similar outcomes as scrolling and steering.
For example, an alternative to scrolling can be to leverage the groups of the overview to access the rows mapped to that group. How do we leverage the overview so that users can perform similar simple interaction to steer spreadsheet data and issue formula? As the granularity of the dynamic overview changes, how do we efficiently update the mapping from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless? How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch? How do we ensure such automatic recomputation scales with the size of the data?}

\item {\em Seamless integration as a plug-in.}
Finally, the design of the overview
should be generic so that it can be integrated
with any existing spreadsheet tool on any spreadsheet
dataset without impacting
existing functionalities or look-and-feel.
Any new interactions introduced should be consistent
with traditional spreadsheet semantics,
and complement existing spreadsheet interactions.
Interactions between the overview and the raw spreadsheet
should be coordinated to the extent possible.
\squishend

\begin{comment}
Since we want to allow users to control how the overview
is constructed, we can let them select the grouping
attribute(s). Following this, we need mechanisms to automatically group nearby rows
into meaningful bins.
This necessitates
{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
zooming (unlike text editors or presentation software),
on the overview view instead of the detailed view
(unlike map tools).
Several domains, \eg online maps and program visualization,
adopt semantic zooming
to ensure improved readability and
comprehension of the information displayed~\cite{summers2003experimental}.
However, to the best of our knowledge, no prior work
has ever developed semantic zooming for spreadsheet data.
Customizing these bins will also eliminate
visual discontinuities: users can simply arrange the bins
the way they want and keep them near each other
for easy comparison.
For this, we take inspiration from aggregation
in relational databases. In relational databases, each ``group'' is associated
with one or more aggregates. As long as the groups are
appropriately constructed (using the mechanisms alluded to above),
formulae can be issued once without requiring any steering, and
automatically maintained for all groups when
zooming in and zooming out. 
\end{comment}

\stitle{\noah.} We address the aforementioned challenges
in our tool \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
\saj{a multi-granularity overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a fine-grained or coarse grained perspective of the data distribution. The distribution at each granularity is captured by a histogram,
enabling users to assimilate the data. Each bin (group) of the histogram is mapped to a collection of rows in the spreadsheet.} Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface. Instead of steering to analyze the data,
users can issue formulae
on the overview with interactions similar to pivot table construction,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b). 
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

The primary contribution of our work is twofold:
\squishlist
\item We formalize the design of a general
navigation (overview+detail) interface
for exploration and analysis of large spreadsheets.
We realize this design in the form of \noah,
a plugin to an existing web-based spreadsheet tool,
ensuring that interactions supported by \noah
complement existing spreadsheet operations.
\item We conduct a user study to evaluate
the benefits and limitations of this plugin.
\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
tasks involving navigation 
correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
\squishend

\toappendix{The rest of the paper is organized as follows: In Section~\ref{sec:usage} we first present a case study of \noah being applied to a real-world scenario.  
In Section~\ref{sec:related}, we discuss related work. We then explain the design considerations for developing \noah in Section~\ref{sec:design}. In Section~\ref{sec:ui}, we present \noah’s user interface and supported operations. We then discuss the design and outcomes of a user study evaluating the effectiveness of \noah in Section~\ref{sec:study}. Finally, we discuss the limitations of \noah and possible future enhancements in Section~\ref{sec:discuss}.}

\section{\noah Use Cases}
\label{sec:usage}
\saj{As explained in Section~\ref{sec:intro}, users prefer spreadsheets over enterprise solutions to view, directly manipulate, and explore data~\cite{chan1996use,raden2005shedding}. To identify the scope of these tasks on spreadsheets (see Table~\ref{tab:scope}), we draw parallel to the typology of abstract data exploration tasks~\cite{brehmer2013multi}. The typology characterizes the exploration and manipulation tasks performed on visual representations of data. While all the tasks in Table~\ref{tab:scope} can be performed using spreadsheets, \noah enhances the experience for a subset of these tasks, indicated by a checkmark (\checkmark). We explain how \noah complements spreadsheets in accomplishing these tasks in Appendix.}

\begin{table}[!htb]\scriptsize
\caption{Uses cases of \noah.}
\label{tab:scope}
\centering
\begin{tabular}{c c}
\hline
Purpose & Use Cases   \\ \hline
\emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
\emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
\emph{Search}            & \code{browse} (\checkmark), \code{explore (\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
\emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
\end{tabular}
\end{table}

\saj{We now describe a usage scenario that captures the spreadsheet exploration tasks accommodated by \noah (Table~\ref{tab:scope}) while illustrating the benefits of integrating an overview with spreadsheets. Let's assume that} Rebecca, an investigative journalist
is exploring the Inside Airbnb dataset~\cite{web:airbnb},
a dataset of all the Airbnb listings across different US cities.
This dataset was created to investigate the long-standing
accusation that many listings in Airbnb are illegally
run as hotel businesses, while avoiding taxes;
any listing available for rent for more than $60$ days
a year is considered to be operated as a hotel~\cite{accusation}.

Without any prior knowledge of the dataset,
Rebecca starts by finding out information
about cities in the dataset and their corresponding listings;
for example, which cities are present,
and roughly how many listings does each city have \saj{(\code{browse} and \code{explore})}.
Without \noah, Rebecca may have used
a pivot table (discussed in Section~\ref{sec:related})
to construct
a summary---however, since this summary is disconnected
from the underlying data, it is hard for Rebecca to
map the summary statistics to
the raw data to obtain further details
about listings from any given city.
If she wanted to identify listings from a specific city \saj{(\code{lookup} and \code{locate} or \code{identify})},
Rebecca would have to either use search capabilities or
perform an explicit filter for this information,
and would have to switch back and forth between the
pivot table results and the raw listings,
present at disparate locations
on the spreadsheet.
Even at the first step of this exploration,
Rebecca would experience substantial cognitive burden,
loss of context,
and visual discontinuities,
with subsequent steps becoming progressively
more challenging.

Using \noah, she explores
the data by city,
with \noah providing a high-level overview of
cities (Figure~\ref{fig:ux}a).
The overview consists of sorted non-overlapping bins
containing one or more cities.
She can click on any bin and the corresponding
data will be displayed at the top of her screen.
For example, clicking on the {\em Ashville-Boston} bin
displays the Ashville listings (Figure~\ref{fig:ux}c).
She can also zoom into bins using the ``\textcolor{blue}{$\rangle$}'' arrows,
zoom out of bins using the ``\textcolor{blue}{$\langle$}'' arrows,
and pan by clicking on various bins at the same level.
We discuss the construction of the overview
and associated interactions in Section~\ref{sec:ui}.

Next, say Rebecca wants to analyze
one of the larger cities \saj{(\code{query}/\code{summarize}) to understand the renting pattern}.
She identifies these cities by
comparing the number of listings
for each city displayed on the overview \saj{(\code{compare})}.
She decides to focus on Boston, her hometown,
and wants to find out how many listings
in Boston violate
the ``rent availability $>60$ days'' condition.
In a typical spreadsheet, Rebecca needs to manually steer and then select
the Boston listings as input to a \code{COUNTIF} formula that counts the number of
rows that satisfy the above mentioned condition.
Using \noah, she can zoom into
the {\em Ashville-Boston} bin
(Figure~\ref{fig:concept}a and~\ref{fig:concept}b)
and then issues
a \code{COUNTIF} operation on the overview.
%\ie a \code{COUNTIF} formula\footnote{\code{COUNTIF} is a
%type of formula in spreadsheets that counts the number %of
%rows that satisfy some condition.}. 
The result is displayed as an {\em aggregate column}
alongside the overview (Figure~\ref{fig:ux}b).
Rebecca learns \saj{(\code{discover})} that more than half of the listings in Boston are 
effectively operating as hotels---a large number!

Based on this insight, Rebecca then wants to
understand availability statistics for an even
larger city, Chicago \saj{(\code{summarize})}.
As she uses the overview to navigate
to Chicago, \noah
automatically updates the aggregate column to
the \code{COUNTIF} formula results for Chicago,
without Rebecca needing to reissue it by performing a cumbersome steering operation as in traditional spreadsheets. 
Rebecca learns that Chicago exhibits 
a similar renting pattern as Boston, with more than half the listings operating as hotels. She can then check if this trend holds for all large cities \saj{(\code{compare})},
and whether the smaller cities have a different pattern \saj{(\code{discover})}. \saj{Note that, the rows that satisfy the ``rent availability $>60$ days'' condition, are listed
 in the spreadsheet adjacent
to the overview in sky blue (Figure~\ref{fig:ux}g)}. With the raw 
data presented side-by-side, she can also dive into other attributes
of the listings operating as hotels to see if there are any other identifying characteristics,
\eg if they are all managed by a small number of agencies acting as individual renters \saj{(\code{identify})}.

 Finally, as Rebecca navigates the data,
 her navigation history (Figure~\ref{fig:ux}d),
 \ie recently visited cities, and current navigation path
 (Figure~\ref{fig:ux}e) are kept up-to-date,
 allowing her to maintain context during navigation.
 She can revisit any previously visited cities
 by simply clicking on the relevant path
 in the navigation history.

Overall, with \noah, users can quickly
comprehend the data via the overview, access any region 
within the data without having to
scroll endlessly, and request additional details on demand without having to
steer across multiple screens. As users navigate and analyze the data, 
they can revisit previously accessed data via the navigation history, not losing context
of what they have explored.

### 2019-10-10 letter.tex
Initial Overleaf Import

\section*{Resubmission of the InfoVis 2019 submission titled ``Extending Spreadsheets to Support Seamless
Navigation at Scale'' to TVCG}

A previous version of this paper was submitted to InfoVis 2019. This version was rejected as the paper co-chairs deemed that revising the paper would require more time than allocated to meet reviewer expectations for the conference. The paper co-chairs mentioned the following: ``\textit{We particularly encourage such revisions where submissions were positively received by reviewers, but the revisions required were deemed to be beyond the scope of the conference review cycle. If you address the issues raised and subsequently submit to IEEE TVCG, please make reference to this IEEE InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}''

The present version of the paper addresses all of the issues mentioned by the reviewers. The entire set of reviews are attached. In particular, the reviewers asked for the following changes for a revision: \begin{itemize}
    \item More clearly articulate the merits over alternative tools, such as Keshif. 
     \item Convincingly explain the design choice around the binning mechanism or revise
    that approach.
    \item Justify the choice of comparing with Excel and note the limitations of that
    approach.
     \item Better justify the choice of tasks in the study. 
     \item Explain the choice of the four research questions.
     \item Provide more details about the study, including reporting intra-participant
    differences.
\end{itemize}

Here, we provide a response to the reviews. In our revised submission,
\textbf{(i) we discuss the merits of introducing a spreadsheet integration like \noah over alternative tools, \eg Keshif;
(ii) we explain our design choice surrounding the binning mechanism;
and 
(iii) we enrich the user study design section further by providing justification for comparing \noah and Excel, rephrasing our research questions reflecting the purpose of the user study, explaining the choice of quiz tasks,
iv) we reorganize the results section to  highlight key contributions of \noah while explaining participant experiences in using the tool. Moreover, we report the  intra-participant differences for the study}. 
The major changes to the paper include:

\begin{enumerate}
\leftmargin=25pt \rightmargin=0pt \labelsep=5pt \labelwidth=20pt \itemindent=0pt \listparindent=0pt \topsep=0pt plus 2pt minus 4pt \partopsep=0pt plus 1pt minus 1pt \parsep=0pt plus 1pt \itemsep=\parsep

\item \textit{Explanation of the binning mechanism.} We have explained our choice of the binning mechanism in the context the third design consideration (\emph{DC3}) and can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}. \saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge the limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss}.

\item \textit{Justification of Excel as a Baseline.}
We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets in interacting with data even for the users of enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and the TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to facilitate direct manipulation and arbitrary derivation of the data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which, far exceeds that of the TDA tools (see Section~\ref{sec:study_design}).

\item \textit{Merits over other tools.} \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation interface}, the concepts of which can be implemented and integrated as a plug-in to any spreadsheet. As spreadsheets have a massive user-base, enhancing the experience surrounding data exploration and formula computation while maintaining the spreadsheet look and feel as much as possible, has the potential to impact a lot of users. As we mentioned earlier, even users of enterprise solutions prefer Excel for direct data manipulation. Therefore, developing a solution to improve the experience with basic operations of such a tool has its merits.}

\item \textit{Enhancing the user study design.}
We have added additional explanations on the goal of the study while better articulating the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study. \saj{Specifically, we merged a number of research questions that explored how specific features of \noah affected participants' navigation experience into one broad question which explores their overall experience with \noah and its components. Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}, and b) explaining how these tasks can reveal how people may utilize \noah for spreadsheet navigation. Finally, we have identified two limitations of the study , resulting from sample size and the study design, which we justify based on the scope of the study.}

\item \textit{Reporting Intra-participant Differences.}
Finally, we have performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data. \saj{For the latter, we find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah. However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}. Therefore, we have included the results of the intra-participant analysis in appendix while briefly summarizing its outcomes in Section~\ref{sec:rq1}.}
\end{enumerate}

\saj{In addition to the aforementioned changes, we have reorganized Section~\ref{sec:results} to reflect the rephrased research questions while supporting our findings from the study by providing qualitative observations, such as participants' comments and interaction patterns with the \noah. We have explained how different features of \noah affected participants' spreadsheet navigation experience, either positively or negatively. Based on these observations, we have updated Section~\ref{sec:discuss} to include additional limitations of \noah while highlighting future research opportunities.}

### 2019-10-10 old-introduction.tex
Initial Overleaf Import

--- 2019-10-10 introduction.tex
@@
 helping them quickly assimilate the information space~\cite{cockburn2009review}.
-\saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}
+In this paper, our goal is to extend
+existing spreadsheets
+with an overview plug-in, giving an overview+detail interface.
 
@@
 developing it leads to several challenges.
-
 \squishlist
 \item {\em Overview modality.}
-\saj{One approach for an overview is to simply add a
+One approach for an overview is to simply add a
 zoomed out version of the entire sheet as a pane on the side,
@@
 somewhat more informed
-scrolling decisions, the zoomed out presentation of the overview would be unreadable. An overview should provide a comprehensible big picture view of the spreadsheet.
+scrolling decisions, the zoomed out presentation of the overview would be unreadable.
 Another approach, adopted by map tools like the
@@
 is to use the overview
-to provide a global context of the user's current location
-being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
-While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
-view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
+to provide a \saj{global} context of the user's current location
+\saj{being displayed} in the \saj{zoomed in} detailed view~\cite{cockburn2009review}.
+Users can perform zooming operations on the detailed
+view as the overview remains static.
+Since spreadsheets already d  isplay the raw data,
 zooming into and out of a detailed view consisting of
-this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users sought more fine-grained or coarse-grained view of the overall structure of the data.}
+this raw data is not meaningful.
+Instead, the goal of our spreadsheet
+overview is to provide a big picture view of the spreadsheet,
+capturing the
+data distribution and \saj{augmenting it with summarization},
+while enabling meaningful interactions
+for navigating the data.
+This necessitates
+{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
+zooming (unlike text editors or presentation software),
+on the overview view instead of the detailed view
+(unlike map tools).
+Several domains, \eg online maps and program visualization,
+adopt semantic zooming
+to ensure improved readability and
+comprehension of the information displayed~\cite{summers2003experimental}.
+However, to the best of our knowledge, no prior work
+has ever developed semantic zooming for spreadsheet data.
 
 \item {\em Construction of the overview.}
-\saj{Given a spreadsheet
+Given a spreadsheet
 with many rows,
-one approach to constructing a dynamic overview is mapping rows of data to high level groups similar to online maps.
-In online maps, cities are grouped into states and states are grouped into countries forming a multi-granularity hierarchy.
-How do we automatically group spreadsheet rows together in a similar ``meaningful''
+it is not clear how to employ semantic zooming
+to construct the overview.
+How do we group rows together in a ``meaningful''
 way
 such that this grouping applies to all 
-data types, including strings and numbers? If the automatic grouping is not meaningful, how do we allow the users
-to customize the overview
-to construct meaningful groups? How do we facilitate interactions that enable users view the overview at multiple granularities?}
+data types, including strings and numbers?
+Since we want to allow users to control how the overview
+is constructed, we can let them select the grouping
+attribute(s).
+Following this, we need mechanisms to automatically group nearby rows
+into meaningful bins.
+Further, the bins need to be constructed
+at various granularities,
+allowing users to zoom in and zoom out of the overview.
+We also need to allow the users
+to customize the bins
+to construct meaningful categories, while providing
+interpretable and intelligent defaults.
+Customizing these bins will also eliminate
+visual discontinuities: users can simply arrange the bins
+the way they want and keep them near each other
+for easy comparison.
 
 \item {\em Operations on the overview.}
-\saj{Following the construction of a dynamic overview, the next challenge is to design simple interactions that achieve similar outcomes as scrolling and steering.
-For example, an alternative to scrolling can be to leverage the groups of the overview to access the rows mapped to that group.
-How do we leverage the overview so that users can perform similar simple interaction to steer spreadsheet data and issue formula?
-As the granularity of the dynamic overview changes, how do we efficiently update the mapping from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless?
-How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch?
-How do we ensure such automatic recomputation scales with the size of the data?}
+While the scrolling issues are partially addressed
+by having the overview, the steering issues remain unaddressed.
+How do we leverage the overview to allow users
+to select regions of interest for issuing formulae
+without steering
+across multiple screens?
+As the granularity of the overview changes (during zoom in/out)
+how do we ensure that users don't have to reissue the formula from scratch?
+For this, we take inspiration from aggregation
+in relational databases. In relational databases, each ``group'' is associated
+with one or more aggregates. As long as the groups are
+appropriately constructed (using the mechanisms alluded to above),
+formulae can be issued once without requiring any steering, and
+automatically maintained for all groups when
+zooming in and zooming out.
 
@@
 should be coordinated to the extent possible.
+
 \squishend
 
-\begin{comment}
-Since we want to allow users to control how the overview
-is constructed, we can let them select the grouping
-attribute(s). Following this, we need mechanisms to automatically group nearby rows
-into meaningful bins.
-This necessitates
-{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
-zooming (unlike text editors or presentation software),
-on the overview view instead of the detailed view
-(unlike map tools).
-Several domains, \eg online maps and program visualization,
-adopt semantic zooming
-to ensure improved readability and
-comprehension of the information displayed~\cite{summers2003experimental}.
-However, to the best of our knowledge, no prior work
-has ever developed semantic zooming for spreadsheet data.
-Customizing these bins will also eliminate
-visual discontinuities: users can simply arrange the bins
-the way they want and keep them near each other
-for easy comparison.
-For this, we take inspiration from aggregation
-in relational databases. In relational databases, each ``group'' is associated
-with one or more aggregates. As long as the groups are
-appropriately constructed (using the mechanisms alluded to above),
-formulae can be issued once without requiring any steering, and
-automatically maintained for all groups when
-zooming in and zooming out. 
-\end{comment}
-
-\stitle{\noah.} We address the aforementioned challenges
+\noindent
+We address the aforementioned challenges
 in our tool \noah,
@@
 the data by a specific attribute,
-\saj{a multi-granularity overview is constructed and displayed
+an overview is constructed and displayed
 within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a).
-Users can zoom into or out of the overview to obtain a fine-grained or coarse grained perspective of the data distribution.
-The distribution at each granularity is captured by a histogram,
-enabling users to assimilate the data. Each bin (group) of the histogram is mapped to a collection of rows in the spreadsheet.} Cumbersome scrolling operations are eliminated in favor
-of a few clicks on the overview interface. Instead of steering to analyze the data,
-users can issue formulae
-on the overview with interactions similar to pivot table construction,
+The overview is a histogram
+and depicts the overall data distribution
+for the chosen attribute,
+enabling users to assimilate the data.
+\noah ensures that there is coordination between
+the overview and the spreadsheet:
+for example, panning and zooming on the overview
+are reflected on the spreadsheet by displaying
+the spreadsheet data corresponding to the bin
+currently in focus in the overview.
+Cumbersome scrolling operations are eliminated in favor
+of a few clicks on the overview interface.
+% \noah leverages {\scshape DataSpread}'s
+% underlying spatial mapping of data to ensure coordination between the overview and the spreadsheet so that users can relate the interactions on the overview with the raw data.
+Instead of steering to analyze the data,
+users can issue aggregation operations
+on the overview,
 and view results on a separate {\em aggregate column},
@@
 reducing visual discontinuity.
-\noah ensures that there is coordination between
-the overview and the spreadsheet:
-for example, panning and zooming on the overview
-are reflected on the spreadsheet by displaying
-the spreadsheet data corresponding to the bin
-currently in focus in the overview.
 Finally, \noah automatically creates

### 2019-10-10 old-result.tex
Initial Overleaf Import

\section{Results}
\label{sec:results}
In this section, we analyze the quantitative and 
qualitative data collected during the quiz and interview phases 
to address the research questions in Section~\ref{sec:study_design}. 
We first discuss the results of the quiz phase tasks. 
We then present participants’ experiences in using \noah and Excel 
based on the interview and survey responses.

\subsection{Analysis of the Quiz Phase}
We first compare the participants’ task completion time 
and accuracy for \noah with Excel.  
\begin{figure}[t]
   \centering
\begin{tabular}{c c}  %trim=left bottom right top
 \includegraphics[width=0.23\textwidth,trim={18 20 20 15},clip]{images/bird_box.pdf} &
   \includegraphics[width=0.23\textwidth,trim={18 20 20 15},clip]{images/airbnb_box.pdf} \\
   \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
\end{tabular}
\vspace{-13pt}
\caption{Submission times per category for each dataset. Median submission times are much smaller for \noah compared to Excel.}
\vspace{-20pt} 
\label{fig:timeBox}
\end{figure}

\subsubsection{Comparison of Task Completion Time and Accuracy (\emph{RQ1})}
\stitle{Distribution of Completion Times.}
In Figure~\ref{fig:timeBox}a and~\ref{fig:timeBox}b, 
we show the distribution of submission times of participants 
for the four task categories, for birdstrikes and airbnb respectively. 
For most categories, 
\emph{participants’ median submission time for \noah 
was less than the fastest submission time for Excel},
confirming that the new capabilities offered by \noah
makes spreadsheet navigation faster.

\stitle{Accuracy Distribution.} In Figure~\ref{fig:acc}a and~\ref{fig:acc}b, 
we show the percentage of correct submissions for the four task categories, 
for birdstrikes and airbnb respectively. 
For all the tasks except for \cmpA---for which the accuracy
was the same for both tools---participants 
attained slightly higher accuracy
with \noah compared to Excel. 
\toappendix{Analysis of screen capture for \cmpA 
revealed that between Florida---the correct answer---and California, 
three participants ($P12$, $P13$, and $P20$) out of 20 
chose the latter when using \noah.}

Since we conducted the study 
on a small population (20), 
we further evaluated the statistical significance of the results---see our supplementary material.
For all tasks except customize, the difference in completion times was statistically significant. 
However, the difference in accuracies was statistically significant 
for the steer tasks only.

\begin{figure}[!htbt]
\vspace{-7pt}
   \centering
\begin{tabular}{c c} %trim=left bottom right top
   \includegraphics[width=0.23\textwidth,trim={18 20 20 20},clip]{images/bird_accC.pdf} &
   \includegraphics[width=0.23\textwidth,trim={18 20 20 20},clip]{images/airbnb_accC.pdf} \\
   \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
\end{tabular}
\vspace{-13pt}
\caption{Per category accuracy for each dataset. Participants attained higher accuracy while completing tasks in NOAH compared to Excel.}
\vspace{-10pt}   
\label{fig:acc}
\end{figure}
\toappendix{
\stitle{Statistical Significance Testing.} Since we conducted the study on a small population (20 participants), we further evaluated the statistical significance of the results explained earlier. To measure the significance of the task completion times, we ran \emph{Mann-Whitney's U test} (as completion times did not follow a normal distribution). \emph{For all of the tasks except the customize task, we found a significant effect of the tools, \ie the response times for the tasks significantly differed by the choice of the tool} (see Table~\ref{tab:statTest}). On the other hand, we ran the \emph{Fisher's exact test} that measures statistical significance of categorical data (0/1 accuracy): \emph{only for steer tasks, the percentage of the accuracy of submissions significantly differed by the choice of the tool}. We now explain our observations regarding users navigational behavior as they performed different tasks, and how that affected their performance.

\begin{table}[!htb]{}
\vspace{-10pt}
\scriptsize
\caption{Statistical significance of submission time and accuracy comparisons between \noah and Excel. (*) indicates statistically significant.}
\label{tab:statTest}
\centering
\begin{tabular}{cccc}
\hline
Question & Category &  Time   &    Accuracy \\
& & ($p$ value) & ($p$ value) \\ \hline
$Q1$ & Steer & $\mathbf{0.0007}$ (*) &  $\mathbf{0.0033}$ (*)  \\
$Q2$ & Identify &   $\mathbf{2.49 \times 10^{-5}}$ (*) & $0.7475$ \\  
$Q3$ & Steer & $\mathbf{0.0043}$ (*) & $\mathbf{0.0202}$ (*) \\
$Q4$  & \cmpA & $\mathbf{0.0154}$ (*) & $1$ \\  
$Q5$ & \cmpB & $\mathbf{5.83 \times 10^{-6}}$ (*) & $0.48$\\
$Q6$ & Customize & $0.1207$ &    $0.0959$ \\\hline
\end{tabular}
\vspace{-10pt}
\end{table}
}
\subsubsection{Steer Results (\emph{RQ1})}
The steer tasks required participants 
to issue a \code{COUNTIF} formula on a data subset. 
However, in Excel, the completion time and accuracy for this
differed significantly from one dataset to another 
(see Figure~\ref{fig:timeBox} and~\ref{fig:acc}). 
We found that the data type of the columns 
on which steering was to be performed 
for the two datasets---binary (birdstrikes) 
as opposed to non-binary (airbnb)---contributed to the deviation.

\stitle{Data type impacts task performance in Excel.} Analysis of video recordings
revealed that for birdstrikes, several participants 
used the \emph{autosum}
feature to quickly count the number of 1s in the column---summing up 
binary values is equal to the number of $1$s in the collection. 
Other participants used the status bar at the 
bottom of the spreadsheet that displays the sum 
of the cells in a selected column. 
In both cases, participants avoided steering the data. 
On the other hand, for airbnb, participants 
could not use these shortcuts as the column was non-binary 
(it had 365 different values) 
and ended up steering the data, resulting in higher completion times. 
Therefore, the participants’ ability to avoid steering 
depended on the data type. 
Failure to avoid steering led to participants 
often selecting an incorrect range of data ($N=14$), resulting in incorrect
responses.The errors mostly occurred for airbnb ($N=11$) where participants
were not able to avoid steering.

\stitle{Data-independent navigation in \noah.} \emph{Using \noah, participants were able to avoid steering by issuing an aggregate formula 
on the overview, resulting in faster submission 
times as well as higher accuracy.} 
Irrespective of the data type, participants used the same strategy  
resulting in consistent performance---the submission time 
for both datasets with \noah is similar (see Figure~\ref{fig:timeBox}). 
\subsubsection{Compare Results (\emph{RQ1})}
Each comparison task in Excel required users to steer 
and then compare the results. 
For \cmpB tasks in Excel, participants 
had to perform $N$ comparisons, 
each time repeating similar tasks, \eg steering, or autosum.
As a result, \cmpB had higher submission times than \cmpA.

\stitle{Repetitive tasks avoided in \noah.} The completion times were faster in \noah
than in Excel, irrespective of the comparison tasks and datasets. \emph{\noah offers three additional
benefits apart from avoidance of steering that may have contributed to such improvement}:
a) participants used the navigation features to access
and compare different subsets of data quickly, 
b) participants did not need to reissue the
aggregation formula for any of the bins they navigated to, and 
c) participants used the value bars presented along with
the results in the aggregate column to visually compare different subsets.

\stitle{Visual discontinuity impacts comparisons.} 
For \cmpB tasks,
the number of subsets to be compared was 
higher in birdstikes (50) compared to airbnb (16). 
As a result,
participants exhibited lower accuracy for birdstrikes  
when using Excel.
In \noah, all the participants first split 
all the bins to create $N$ bins each corresponding to one state. 
Then participants panned across the overview 
to find the desired bin as they compared the values. 
Even in \noah, \emph{comparisons
between multiple values resulted in increased 
visual discontinuity leading to some ($N=4$ out of 20) incorrect submissions.}

\subsubsection{Identify Results (\emph{RQ1, RQ2})}
\vspace{-5pt}
\stitle{Coloring cues help accelerate identification in \noah.} 
For the \emph{identify} task, 
participants had to skim through all the cells 
in the current window in Excel, 
resulting in higher completion times. 
Again, participants had to skim over binary values when counting in 
birdstrikes versus non-binary values for airbnb, resulting in even higher completion times for the latter. 
In \noah, participants benefited from having visual cues in the form of colored cells, helping them relate the aggregate column with the raw data: \emph{they quickly identified the spreadsheet cells that did not satisfy the condition based on cell color}.
\subsubsection{Customize Results (\emph{RQ1, RQ3})}
\vspace{-5pt}
\stitle{Bin customization is powerful but time-consuming.} For airbnb in Excel, the customize task involved filtering out 26 values from the filter menu compared to filtering 451
values for the birdstrikes dataset. As a result, participants had to manually filter a large number of values and took more time
to submit their responses. Participants took much less time in \noah, as they were able to use bin customization.
However, the time taken for this task was higher than  other tasks in \noah, as
it required participants to restructure the overview
before any calculation could be performed. Unfamiliarity with customization
operations also contributed to higher task completion times (see Section~\ref{sec:discuss}).

\subsection{Analyzing Intra-participant Submission Times}
\saj{Figure~\ref{fig:ipdiff} shows the pivot table representation of the intra-participant submission time differences between \noah and Excel across the quiz tasks. The submission time differences are depicted by the circles---the larger the circle the higher the submission time differences. The color of the circle indicates whether the submission time for a task in \noah was faster (green) or slower (orange) than Excel. Figure~\ref{fig:ipdiff} confirms that majority of the submission times using \noah were faster compared to Excel---nineteen out of the twenty participants completed at least four tasks in less time using \noah compared to Excel. The difference in submission times is apparent for the first steer task which implies that when exploring an unknown dataset, \noah may enable users to navigate the data quickly compared to Excel. The submission time difference is even more apparent for the \cmpN task---participants having to repeat the same formula $N$ times in Excel contributed to such differences.}

\saj{However, for some of the tasks, less than a third of the participants' submission times were faster using Excel: \cmpA, customize, and the second steer task.
For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah, utilized the \emph{autosum} shortcut for the birdstrikes dataset. For the \cmpA tasks, three out of the five participants that required more time to submit answer using \noah, first used the bin customization feature to view all the unique bins which contributed to the higher submission time compared to Excel. On the other hand, as explained earlier, the unfamiliarity with the bin customization feature contributed to higher submission time using \noah for six participants.}

\begin{figure}
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 0 0 0,clip,width=\linewidth]{images/userinfo-ip.pdf}
\vspace{-20pt}
   \caption{Intra-participant submission time differences between \noah and Excel across the quiz tasks.}
\vspace{-15pt}
   \label{fig:ipdiff}
 \end{figure}

\subsection{Qualitative Observations}
\label{sec:interview_result}\label{sec:survey_result}
We now present a qualitative analysis of the participants' perceptions of both tools based on their interview responses.

\stitle{Data comprehension easier in \noah.} 
Without an overview, participants found it difficult to 
perform various tasks on Excel---\emph{Excel can get overwhelming 
if you have a lot of data in it and sometimes with that data 
finding things can be difficult ($P11$).} 
The binned representation of the data in \noah, 
on the other hand, helped users to comprehend 
the overall structure of the data better 
and prioritize the bin they want to visit next, enabling faster navigation. 
Out of 20 participants, 15 preferred \noah for tasks involving navigation. 
One participant ($P5$) commented: 
\emph{I think it was just a little bit easier to navigate 
and find where things were because you could already see what bins had what.} 
Another participant ($P1$) commented:  
\emph{I like \noah a lot better. It was a lot easier to look up different data 
and it was a lot quicker too}. 
Participants ($N=6$) additionally mentioned 
that they would prefer \noah over Excel when the dataset is large. 
One participant ($P2$) commented: 
\emph{I guess if I just had a large amount of data 
then I would prefer to use \noah because then you 
would be able to see all of it $\ldots$at once.} 
Finally, participants preferred the automatic data highlighting 
feature while performing the \emph{identify} task (16 out of 20). 
The colored cells focused their attention 
to relevant spreadsheet regions---\emph{It was a visual 
cue right there, made it very quick to count it up ($P17$).}
However, one participant ($P5$) mentioned the fact that 
some of the color choices would make \noah unusable 
for the colorblind population.

\toappendix{
\stitle{Overview-Spreadsheet Coordination.} Out of 20 participants, 16 participants preferred the automatic data highlighting feature of \noah while performing the \emph{identify} task. Moreover, the colored cells focused their attention to relevant regions within the spreadsheet---\emph{You didn't have to do any additional steps and it was a visual cue right there, made it very quick to count it up ($P17$).} However, one participant ($P5$) mentioned the fact that some of the color choices would make \noah unusable for the colorblind population.
}

\stitle{Scrolling and steering is cumbersome in Excel.} 
Participants found scrolling and steering in Excel 
to be cumbersome while issuing formulae 
and performing comparisons---\emph{The one thing 
with Excel is I always try to go to the bottom of the data 
and type in the formula, and with something really long like this, 
the scrolling is a little bit cumbersome ($P4$).} 
With \noah users can 
(a) avoid scrolling by using clicking or zooming operations, 
and (b) avoid steering by performing aggregate operations on the overview. 
One participant ($P3$) commented: 
\emph{And that creates convenience sort of because then you don't have to memorize anything and using the system becomes easier.} 
Another participant commented: \emph{With \noah, 
you don't have to highlight every number $\ldots$ 
versus Excel you actually have to select everything ($P12$).}

\stitle{Fine-grained control is powerful but sometimes difficult to use.}
The bin customization feature enables users to 
personalize the overview based on their specific needs. 
One participant ($P16$) commented: 
\emph{I did like the fact that it lets you take a data sheet and, in some way, containerize the stuff you care and the stuff you don't care about.} 
14 out of 20 
participants found the bin customization feature to be useful. 
Participants preferred the feature to Excel’s filtering feature 
when working with numeric data---\emph{That was so much easier in \noah 
than it was in Excel to be able to specify the range that you wanted it to go in ($P17$).} However, six participants found the feature to be difficult 
to use when working with textual data and would have 
preferred a pivot table-like single level overview---\emph{I would have expected them to completely be split, 
and then I can merge them if I want to ($P13$).}

\stitle{Participants preferred \noah to Excel on all subjective user satisfaction metrics ({\em RQ4})} Figure~\ref{fig:survey} displays the subjective participant ratings 
for various satisfaction metrics using a diverging stacked bar chart. 
For a given metric, each white ellipse denotes the average rating 
given by the participants for the corresponding tool. 
We can see from the figure that participants preferred \noah to Excel 
for all metrics. 
Notably, participants found \noah to be easier to use 
compared to Excel, while being faster in execution. 
Even though the features offered by \noah were new to participants 
they found it fairly easy to learn. 
\toappendix{We further conducted a statistical significance test---the \emph{Wilcoxon Signed-rank} test---on the survey responses which showed that for all the metrics, the ratings significantly differed by the choice of the tool, \ie \noah or Excel. 
The distribution of the ratings for none of the criteria followed a normal distribution.}

\begin{figure}
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 0 0 0,clip,width=\linewidth]{images/survey.pdf}
\vspace{-20pt}
   \caption{Participants found \noah to be easier to use compared to Excel while being faster in completing tasks involving navigation.}
\vspace{-15pt}
   \label{fig:survey}
 \end{figure}

\toappendix{
\begin{table}[!htb]{\scriptsize}
\caption{Survey results. (*) indicates statistical significance. }
\label{tab:survey}
\centering
\begin{tabular}{|l|l|l|l|}
\hline
Metric &    \noah   &    Excel   & $p$ value \\ \hline
Ease of Learning &  $\mu = 5.75$, &    $\mu = 4.22$, & $\mathbf{1.49 \times 10^{-7}}$ (*) \\
& $\sigma = 1.02$ & $\sigma = 1.41$ & \\\hline
Speed of Use &    $\mu = 6.03$, & $\mu = 4.22$, & $\mathbf{1.68 \times 10^{-7}}$ (*)\\
& $\sigma = 0.99$ & $\sigma = 1.65$ & \\\hline  
Ease of Use    & $\mu = 5.88$, & $\mu = 4.33$, & $\mathbf{7.85 \times 10^{-6}}$ (*) \\
& $\sigma = 0.90$ & $\sigma = 1.71$ & \\\hline
Confidence    & $\mu = 5.50$, & $\mu = 4.60$, & $\mathbf{0.0096}$ (*)\\
& $\sigma = 1.79$ & $\sigma = 1.50$ & \\\hline  
Comprehensibility & $\mu = 5.60$, & $\mu = 4.48$, & $\mathbf{0.0006}$ (*)\\
& $\sigma = 1.27$ & $\sigma = 1.65$ & \\\hline
Satisfaction & $\mu = 5.48$, &    $\mu = 4.52$, & $\mathbf{0.0018}$ (*)\\
& $\sigma = 1.16$ & $\sigma = 1.49$ & \\\hline
\end{tabular}
\end{table}
}
\section{Discussion}
\label{sec:discuss}

In this section, we discuss the shortcomings of our user study, 
the limitations of \noah as a spreadsheet plug-in for navigation, 
and possible enhancements.

\subsection{Limitations}
We first discuss the limitations of \noah.

\stitle{Limitations of Study Design.} Our evaluation compared a \noah-integrated spreadsheet with a traditional spreadsheet. However, it is certainly possible to evaluate intermediate versions of \noah that isolate the effects of the zoomable overview (visual clarity vs.~visual continuity), display layout (screen space trade-off), and 
contextual presentation of data (raw text vs.~chart representation of aggregate columns). Moreover, our participant pool demographics may not match the demographics of the general audience intended for this tool.

\stitle{Unfamiliarity with \noah.} Several interactions on \noah 
are quite different from typical spreadsheet interactions, \eg zooming, bin customization. 
This unfamiliarity led to some participants ($N=5$ out of 20) 
preferring Excel over \noah. \toappendix{One participant ($P11$) commented: \emph{Since I'm not used to spreadsheet data being presented that way, it took a little bit of getting used to.}} Participants found some of the terminologies 
used in the interface---\eg explore, bin---to be quite 
unfamiliar ($N=14$)  for an Excel audience. 
Moreover, two participants didn’t understand how the bins were constructed. 
These issues can be addressed by using more relatable terminologies and proper documentation. 
Two participants found the aggregation operations 
applied on the bins to be opaque compared to Excel 
where a user can directly manipulate the formula\toappendix{---\emph{I like the way that in Excel I can go look at my formula again if I wanted to make sure I'm doing everything I wanted to do correct ($P6$)}}. 
In subsequent versions of \noah, we can further 
display the appropriate formula for each bin as users hover 
over the corresponding cell on the aggregate column. 
Moreover, participants ($N=5$) noted the fact that \noah currently 
does not support user defined formulae.

\stitle{Binning Mechanism.} \saj{A number of participants ($N=6$)
did not prefer the current binning mechanism for categorical data. 
As explained in Section~\ref{sec:design},
the purpose of such binning
was to reduce visual discontinuity.
The outcomes of the \cmpB tasks in Section~\ref{sec:results}
further justifies this choice 
as participants ended up
making more mistakes
when the overview displayed all the unique bins.
However, the experience 
surrounding the construction
of the overview
can be further improved.
Currently, the bins of the overview
can be customized
only after the overview is constructed.
Providing the users
the capability to
select the representation
(similar to bin customization)
of the overview could have addressed this issue.
Understanding the impact of these
representation choices for the overview
can be an interesting future research.}

\subsection{Future Work}
We now discuss the to further enhance the capabilities of \noah.

\stitle{Beyond Tabular Data.} \noah operates only on tabular data. 
However, spreadsheets can be semi-structured---formulae 
and text can be interspersed with tabular data. 
We can extend \noah to act as a map highlighting heterogeneous regions on such complex spreadsheets which users can utilize to navigate the spreadsheet. 
We can leverage existing work on spreadsheet table detection~\cite{tablesense}, 
and property identification~\cite{chen2017spreadsheet} 
to construct the map.

\stitle{Scope of Overview-Spreadsheet Coordination.} \saj{can we generalize this to an unified framework.}
Spreadsheet users may perform various edit operations, 
\eg updating values, adding/deleting rows/columns. 
However, \noah currently assumes the data to be read-only. 
In our next version, we can add support for 
propagating the spreadsheet updates to the overview. 
Moreover, the charts displayed in an \emph{aggregate column} are non-interactive, 
\ie users cannot interact with the charts to visually look up relevant or 
interesting data points within the spreadsheet. 
In the future, we plan to extend \noah to support visual querying 
through the charts in an aggregate column.

\stitle{From Perceptual to Interactive Scalability.} 
The current version of \noah addresses the perceptual 
scalability challenges while navigating 
Excel-scale (one million rows) data. 
As modern spreadsheets continue to support 
even more large scale data---{\scshape DataSpread}~\cite{datamodels} 
supports one billion rows---the interactions proposed in this paper may violate the interactive response time bound of 500 ms~\cite{liu2014effects}. This opens the door to a new set of research challenges that may range from approximate query processing to progressive data analytics.

\stitle{Other Enhancements.} \noah currently constructs the overview on a single attribute. We can add support for multi-attribute navigation (\eg explore the Airbnb data by city and neighborhood), and multi-level navigation (\eg explore the neighborhoods after zooming into a specific city in the Airbnb data). \toappendix{Moreover, as shown in Section Table~\ref{tab:scope}, \noah does not allow users to add annotations, \eg visual cues, texts, to the overview. These operations can be implemented in the future as feature extensions.} Furthermore, bin customization currently supports changing the bin boundaries only while maintaining the current order. Supporting user defined ordering to allow arbitrary reshuffling of the bins can be another enhancement.\toappendix{ However, such modification will require a redesign of the histogram based data structure since the ordering property will not hold anymore.} Finally, a complete characterization of the spreadsheet operations that can be supported by \noah is another open question, \eg other operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting.

\section{Conclusions}
\label{sec:conclusion}
We presented \noah, an in-situ navigation interface, designed as a spreadsheet plugin.
Using \noah, users can get a bird’s eye view of the data, with the ability to scroll or seek
additional details on demand, using a multi-granularity overview, as well as
aggregate columns that eliminate cumbersome steering operations.
We conducted a user study that demonstrated that participants found \noah to be much more easy to use compared to Excel for navigating large spreadsheet datasets. Across different navigation tasks, participants were able to attain 
higher accuracy when using \noah compared to
Excel while being twice as fast. 
\noah represents our 
first step towards making spreadsheets more effective 
on large datasets that are increasingly the norm.

### 2019-10-10 related.tex
Initial Overleaf Import

\section{Related Work}
\label{sec:related}
In this section, we present tools and techniques
that partially address the limitations of navigating
spreadsheets as discussed in the introduction.

\subsection{Spreadsheet Tools and Prototypes}
\cut{Commercial spreadsheet tools include
Microsoft Excel, Google Sheets, and Airtable---of
these, Excel and Google Sheets support features
targeted at improving users' navigation experience.}

\cut{Excel allows users to maintain context during navigation
by manually creating named ranges~\cite{named-range},
that are essentially references
to a spreadsheet region,
and appear in a \emph{name box}
within the Excel menu bar.
Users can click on a named range
to navigate to the referred region.
However, the onus is on the user to create
named ranges for each region of interest.}
\stitle{Microsoft Excel.}
\saj{Excel enables users to manually create references
to a spreadsheet region using the named ranges~\cite{named-range} feature which is displayed within the Excel menu bar.
Users can click on a named range
to navigate to the referred region. However, the onus is on the user to create
named ranges for each region of interest. The Pivot table~\cite{pivot} feature provides a summary view, enabling users
to compare subsets of data without
having to navigate to various locations
within the spreadsheet.
This summary is placed
in a separate region in the spreadsheet,
preventing users from accessing
the data underlying the summary, thereby, impeding navigation. A similar overview feature, \code{SUBTOTAL}~\cite{subtotal},
adds a new row at the end of the
corresponding subset of data with summary
information. Users can expand the summary to view
the actual spreadsheet data.
However, for datasets with larger subsets (\eg for numeric data),
the size of the summary can become very
large spanning multiple screens and can cause
increased
visual discontinuity during navigation. Finally, NodeXL~\cite{hansen2010analyzing}
is an Excel plug-in that provides an overview
of a network
and enables users to perform navigational operations,
\eg zooming in/out, dynamic filtering, on the overview;
this plugin only supports network datasets,
such as biological or social networks.}

\stitle{Google Sheets Explore.}
Google Sheets Explore~\cite{web:explore} provides
an overview of the data by
auto-generating charts of data statistics.
Users can specify queries to the system
(similar to a web search)
asking for different summary statistics.
While Explore
is a convenient means to understand
high-level data characteristics,
it doesn't address the
navigational challenges
related to scrolling and steering.

%\subsection{Academic Spreadsheet Prototypes}
\cut{Existing academic work focuses
on adding a richer set of operations to spreadsheets,
or increasing their scalability.
We discuss work specifically targeting
navigation; other enhancements
such as managing hierarchical data~\cite{chang2016using}
or supporting joins~\cite{bakke2011spreadsheet},
are omitted.}

\stitle{Scalable Spreadsheet Summarization and Exploration.}
Smart-drill-down~\cite{joglekar2015smart}
generates an interactive summary of a large spreadsheet table as
a collection of rules; users can drill-down to
a specific rule to view more fine-grained rules.
Hillview~\cite{web:hillview} displays
the results of group-by queries
on large spreadsheet tables.
While these tools support summarization at scale,
providing an overview of the spreadsheet,
they don't preserve spreadsheet semantics,
nor do they make it easy to scroll or steer
through large spreadsheets.
ABC~\cite{Raman99scalablespreadsheets} and {\sc DataSpread}~\cite{datamodels} support interactive
exploration of very large spreadsheet datasets,
beyond main-memory limits, maintaining spreadsheet look-and-feel,
but do not provide any new spreadsheet capabilities to assist
with navigation.
We build \noah as a plugin
to {\sc DataSpread}, since it is open-source.

\stitle{Interactive Tables.}
TableLens~\cite{rao1994table}
is a focus+context view
for browsing numerical information
in tables, looking
much like a spreadsheet with embedded bar charts.
Cells out of focus display
graphical bars proportional in length
to the underlying values,
providing a visual overview of the data,
while cells within the user's current focus
are magnified and display the graphical bars and
the raw data.
Ideas similar to TableLens
have been adopted by DataLens~\cite{bederson2004datelens}
for visualizing digital calendars, and
by FOCUS~\cite{spenke1996focus}
and InfoZoom~\cite{spenke2003visualization}
for exploring database query results.
Like TableLens, \noah
embeds graphical bars, but within
the overview to depict the underlying data
distribution. \noah captures
the user's current focus by highlighting
the corresponding bin in the overview.
While TableLens provides an easy mechanism
to get a high-level view of the data
and spot outliers, it
suffers from the same disadvantages
that focus+context views have relative
to overview+detail ones.
Unlike \noah, which supports
multiple granularities via binning,
TableLens only supports
one granularity (zoomed in or
zoomed out): beyond
a certain size, navigating (scrolling or steering)
the zoomed out data
is still cumbersome for users. Moreover, TableLens does not maintain the spreadsheet look-and-feel or capabilities.
%TableLens does not support non-numeric data,
%nor does it maintain the spreadsheet look-and-feel or capabilities
%(maintained in \noah due
%to the detailed view).

\stitle{Visual Interactive Spreadsheets.}
VisSh~\cite{nunez2000vissh}, SI~\cite{levoy1994spreadsheets},
SSR~\cite{chi1997spreadsheet}, ASP~\cite{piersol1986object},
and PhotoSpread~\cite{kandel2008photospread}
extend the input/output capabilities of cells
within spreadsheets, to
display charts, animation, photos, or geometric objects, or
accept input via direct manipulation dialogs,
among others.
While these tools allow users to represent and manipulate
data in a more flexible manner,
which in turn could help users getting a
high-level sense of the data,
they do not necessarily help users navigate data
more effectively.

\subsection{Spreadsheet Alternatives}
%Finally, we discuss solutions to navigational and summarization
%challenges in other domains.

\stitle{Overview+Detail Interfaces.}
Cockburn et al.~\cite{cockburn2009review}
provides a detailed survey of overview+detail
and zooming interfaces.
To improve navigation within large documents,
overview+detail interfaces~\cite{cockburn2006faster,kratz2010semi} allow users to interact with an overview as they explore the document. Zooming interfaces~\cite{nekrasovski2006evaluation,plaisant2002spacetree} provide a multi-granularity overview of the data and support interactions like zoom in/out to navigate across various granularities.
We follow the same analogy of
providing an overview of the spreadsheet first,
allowing users to drill-down further.

\stitle{Multiple Coordinated Views.} Multiple coordinated views~\cite{roberts2007state}, \eg \emph{Snap}~\cite{north2000snap}, \emph{Elastic Documents}~\cite{badam2018elastic} connect multiple views, for example, an overview and a detailed data view while enabling coordination between these views through brushing and linking. Similarly, \noah connects tabular spreadsheets with an overview and updates the state of the spreadsheet as users interact with the overview and vice-versa.

\stitle{Tabular Data Analysis (TDA).}
Visualization tools such as Tableu~\cite{tableau}, Power BI~\cite{powerbi}, Keshif~\cite{yalccin2018keshif}, Voyager~\cite{wongsuphasawat2016voyager} and
analytical tools such as SPSS~\cite{noruvsis1986spss}, SAS~\cite{sas1985sas}, can all provide summaries of tabular data in various forms (visualizations,
aggregate statistics). 
\saj{However, these summaries 
are static overviews of the data---much like pivot tables, 
these summaries are not dynamically linked
to nor are co-located with
the underlying data. 
Naturally, 
users cannot navigate and directly manipulate the raw data
using these tools unlike spreadsheets. For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}). Even users of such tools want to inspect the underlying data to verify the summary statistics and visualizations~\cite{2017trust}.
%several wanted to be able to see the underlying data: although big data demands aggregations, analysts wanted to see individual records to spot-check their results, and to get a sense of what sat in a bucket. 
The goals of spreadsheets vary from the TDA tools in two different ways: a) facilitating direct manipulation of the data and b) enabling arbitrary derivation of new data and summaries
using various operations involving navigation,
\eg formulas or expressions.
\noah being a plugin to spreadsheets, facilitates a unified interface that
upholds both these goals
while enhancing
the navigational capabilities
for spreadsheet users.
}

### 2019-10-10 results.tex
Initial Overleaf Import

\section{Results}
\label{sec:results}
In this section, we analyze the quantitative and 
qualitative data collected during the quiz and interview phases 
to address our research questions.

\subsection{RQ1. Navigation performance of participants}
\label{sec:rq1}
To answer RQ1, we first compare our participants’ task completion times  
and accuracies in \noah and Excel while analyzing the survey response.  
\begin{figure}[t]
   \centering
\begin{tabular}{c c}  %trim=left bottom right top
 \includegraphics[width=0.23\textwidth,trim={18 37 20 15},clip]{images/bird_box.pdf} &
   \includegraphics[width=0.23\textwidth,trim={18 37 20 15},clip]{images/airbnb_box.pdf} \\
   \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
\end{tabular}
\caption{Submission times per category for each dataset. Median submission times are much smaller for \noah compared to Excel.}
\label{fig:timeBox}
\end{figure}

\subsubsection{Faster navigation without sacrificing accuracy}
In Figure~\ref{fig:timeBox}a and~\ref{fig:timeBox}b, 
we show the distribution of submission times of participants 
for the four task categories, for birdstrikes and Airbnb respectively. 
For most categories, 
\emph{participants' median submission times using \noah 
were less than the fastest submission times using Excel}. This suggests that
the new capabilities offered by \noah
made spreadsheet navigation faster for these tasks. We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations. Overall, majority of the submission times using \noah were faster compared to Excel---19 out of the 20 participants completed at least four tasks in less time using \noah compared to Excel. The submission time differences were more prominent for the steer, identify, and \cmpB tasks. Moreover, for all tasks except customize, the difference in submission times was statistically significant. Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.

\begin{figure}[!htbt]
   \centering
\begin{tabular}{c c} %trim=left bottom right top
   \includegraphics[width=0.23\textwidth,trim={18 35 20 20},clip]{images/bird_accC.pdf} &
   \includegraphics[width=0.23\textwidth,trim={18 35 20 10},clip]{images/airbnb_accC.pdf} \\
   \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
\end{tabular}
\caption{Per category accuracy for each dataset. Participants attained higher accuracy while completing tasks in NOAH compared to Excel.}
\label{fig:acc}
\end{figure}
%\subsubsection{Accurate submissions with \noah} 
In Figure~\ref{fig:acc}a and~\ref{fig:acc}b, 
we show the percentage of correct submissions for the four quiz task categories, 
for the birdstrikes and Airbnb datasets, respectively. 
For all the tasks except for the fourth task, \cmpA, for which the accuracy
was the same for both tools, participants 
attained slightly higher accuracy
with \noah compared to Excel. However, the difference in accuracies was statistically significant 
for the steer tasks only (see Appendix). 
\toappendix{Analysis of screen capture for \cmpA 
revealed that between Florida---the correct answer---and California, 
three participants ($P12$, $P13$, and $P20$) out of 20 
chose the latter when using \noah.}

\begin{figure}[!htbt]
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 0 0 0,clip,width=\linewidth]{images/survey.pdf}
   \caption{Participants found \noah to be easier to use compared to Excel while being faster in completing tasks involving navigation.}
   \label{fig:survey}
 \end{figure}

\subsubsection{Participants' preferences: \noah VS Excel} 
\saj{Figure~\ref{fig:survey} shows a diverging stacked bar chart representation of the survey results in which participants rated their experience with Excel and \noah. 
For each metric mentioned in Section~\ref{sec:study}, there are two stacked bar charts, one for Excel and one for \noah. Each component within a stacked bar represents the percentage of responses for the corresponding rating, where the ratings are on a scale of one one (strong disagreement) to seven (strong agreement). The average rating for each metric is represented with a white ellipse. Notably, \noah had a higher average rating than Excel for all the metrics. The aforementioned observation was further validated by a statistical significance test---the \emph{Wilcoxon Signed-rank Test} (see Appendix). In particular, participants felt that using \noah was faster and easier 
compared to Excel.}
%Even though the features offered by \noah were new to participants they found it fairly easy to learn. 
\toappendix{We further conducted a statistical significance test---the \emph{Wilcoxon Signed-rank} test---on the survey responses which showed that for all the metrics, the ratings significantly differed by the choice of the tool, \ie \noah or Excel. 
The distribution of the ratings for none of the criteria followed a normal distribution.}

\subsection{RQ2. Impact of \noah and its components on spreadsheet navigation}
\label{sec:rq2}
To answer RQ2, we assessed how \noah components impacted navigation in our study.
%We now discuss how \noah impacted participants' navigation experience compared to Excel.

\subsubsection{Binned Overview}
\stitle{Data navigation at scale.}
\saj{Without an overview, participants found it difficult to 
perform various tasks in Excel. One participant ($P11$) commented---\response{Excel can get overwhelming 
if you have a lot of data in it and sometimes with that data 
finding things can be difficult}. Participants ($N=6$) mentioned 
that they would prefer \noah over Excel when the dataset is large: 
\response{If I just had a large amount of data 
then I would prefer to use \noah because then you 
would be able to see all of it (bins) at once} ($P2$).
\noah's binned overview helped participants comprehend
the overall structure of the data better 
and prioritize the bin they want to visit next, enabling faster navigation. One participant ($P5$) commented: 
\response{I think it was just a little bit easier to navigate 
and find where things were because you could already see what bins had what.} Another participant ($P1$) said:  
\response{I like \noah a lot better. It was a lot easier to look up different data 
and it was a lot quicker too}.}
%Out of 20 participants, 15 preferred \noah for tasks involving navigation.

\stitle{Enhanced navigation with guided bin customization.} 
\saj{The bin customization feature enabled users to 
personalize the overview based on their specific needs. 
One participant ($P16$) commented: 
\response{I did like the fact that it lets you take a data sheet and, in some way, containerize the stuff you care and the stuff you don't care about.} 
14 out of 20 
participants found the bin customization feature to be useful. 
Participants preferred the feature to Excel's filtering feature 
when working with numeric data---\response{That was so much easier in \noah 
than it was in Excel to be able to specify the range that you wanted it to go in} ($P17$). Our analysis of the video recordings revealed that, for the birdstrikes dataset in Excel, the customize task involved filtering out the desired values from a total of 451 unique values. As a result, participants had to manually filter a large number of values and took more time
to submit their responses compared to \noah when they were able to use the bin customization feature. However, the time taken for this task was higher than  other tasks in \noah, as
it required participants to restructure the overview
before any calculation could be performed.
Unfamiliarity with customization
operations also contributed to higher task completion times.}

\stitle{Unfamiliar interactions.} \saj{Participants were unfamiliar with the bin customization feature.
This unfamiliarity led to some participants ($N=5$ out of 20) 
preferring Excel over \noah. One participant ($P11$) commented: \emph{Since I'm not used to spreadsheet data being presented that way, it took a little bit of getting used to.} Participants found some of the terminology 
used in the interface---\eg explore, bin---quite 
unfamiliar ($N=14$). 
Moreover, two participants didn’t understand how the bins were constructed.}

\stitle{Overview presentation across data types.}
\saj{Although participants appreciated the binned representation of the overview for numeric data, a number of participants ($N=6$) stated that they would've preferred a pivot table-like single level overview for categorical data where each bin corresponds to one item. One participant ($P13$) commented: \response{I would prefer it start with all the bins split, and then I can merge them as I want.} Another participant ($P4$) said---\response{When I started, it (\noah) had already grouped them, I think, alphabetically. So, that creates an extra step in that I then have to go split them and then re-merge them.}}

\subsubsection{Aggregate Column}
\stitle{Ease of issuing formulae.} 
\saj{The steer tasks required participants 
to issue a \code{COUNTIF} formula on a data subset. 
Participants found scrolling and steering in Excel 
to be cumbersome while issuing formulae---\response{The one thing
with Excel is I always try to go to the bottom of the data 
and type in the formula, and with something really long like this, 
the scrolling is a little bit cumbersome} ($P4$). 
With \noah, participants avoided
(a) scrolling by using clicking or zooming operations, 
and (b) steering by performing aggregate operations on the overview. Participants ($N=13$) found it easier to issue formulae using the aggregate column feature.
One participant ($P3$) commented: 
\response{And that creates convenience sort of because then you don't have to memorize anything and using the system becomes easier.} 
Another participant ($P13$) commented: \response{There were some formulas to calculate, that were definitely easier in \noah because the aggregate column did all the work and showed me the results.} However, two participants found the aggregation operations 
applied on the bins to be opaque compared to Excel 
where a user can directly manipulate the formula. 
}

\stitle{Higher efficiency in issuing formulae.} 
\saj{While the accuracies and submission times for the steer tasks in Excel varied significantly across datasets, using \noah, participants exhibited higher accuracies and faster submission times irrespective of the dataset (see Figure~\ref{fig:timeBox} and~\ref{fig:acc}). The automated and steering-free aggregate column feature of \noah contributed to high accuracies (100$\%$) for the steer tasks.  One participant ($P12$) commented: \response{With \noah, 
you don't have to highlight every number versus Excel where you actually have to select everything.} All of the 14
inaccurate submissions with Excel involved steering incorrect spreadsheets regions. 11 of the inaccurate submissions were with the Airbnb dataset. However, in \noah, participants were able to avoid steering by utilizing the aggregate column feature. Analysis of screen recordings of Excel usage
revealed that for birdstrikes dataset, several participants 
used the \emph{autosum}
feature to quickly count the number of $1$'s in a binary-valued column that was involved in the steering task. Summing up
binary values is equal to the number of $1$'s in the collection. 
Other participants used the status bar at the
bottom of the spreadsheet that displayed the sum 
of the cells in the selected column. 
In both cases, participants avoided steering the data. 
On the other hand, for Airbnb dataset, participants 
could not use these shortcuts as the column that was involved in the steering task was non-binary 
(it had 365 different values). Participants, at times, steered incorrect regions, resulting in inaccurate responses for the task.  
Therefore, the participants’ ability to avoid steering 
depended on the data type. 
Failure to avoid steering often led participants
to selecting an incorrect range of data ($N=14$), resulting in incorrect
responses.}

\subsubsection{Detailed View}
\stitle{Accelerated data inspection.} 
\saj{For the \emph{identify} task, 
participants had to skim through all the cells 
in the current window in Excel, 
resulting in higher completion times. 
%Again, participants had to skim over binary values when counting in birdstrikes versus non-binary values for Airbnb, resulting in even higher completion times for the latter. 
Even though Excel provides a conditional formatting feature, that adds one additional step when performing the \emph{idenitfy} task. 
In \noah, participants benefited from having visual cues in the form of colored cells, helping them relate the aggregate column with the raw data---\emph{You didn't have to do any additional steps and it was a visual cue right there, made it very quick to count it up ($P17$).} Another participant ($P9$) commented---\response{In Excel, you would have to do your own condition of formatting. But you have to build that every time  you need to ask a question. This one (\noah) at least something is pre-built in, and you can easily count.} However, one participant ($P3$) pointed out the fact that, when the data corresponding to the bin does not fit in the screen, they had to scroll through the data to identify relevant information.}
%Out of 20 participants, 16 participants preferred the automatic data highlighting feature of \noah while performing the \emph{identify} task. Moreover, the colored cells focused their attention to relevant regions within the spreadsheet Another participant ($P5$) mentioned the fact that some of the color choices would make \noah unusable for the colorblind population.}

\subsubsection{Context Bar}
\stitle{Utilizing history to avoid repeated execution.}
\saj{For the \cmpA task, participants utilized the context bar to navigate to the previously visited bin for the first steer task. As \noah automatically materialized the aggregate summaries of a previously visited bin, participants were able to view the aggregate column values instantly without having to repeat the same command. However, as Excel did not preserve any navigation history, participants had to re-execute the first steering operation. As a result, the submission times for \cmpA tasks were lower in \noah compared to Excel (see Figure~\ref{fig:timeBox}). One participant ($P16$) commented---\response{Once I got familiar with the interface, it was easy to just say, I want to see this state, and I like that fact that like automatically it goes into the bins on NOAH, gave me summary information.} Another participant ($P9$) said---\response{Noah was easy to find and compare and toggle in between.} }

\stitle{Visual discontinuity during navigation.} 
\saj{For \cmpB tasks, participants 
had to perform $N$ comparisons in \noah while issuing the aggregate column operation once. However, comparison among $N$ bins resulted in increased visual discontinuity. This lead to some ($N=4$ out of 20) incorrect submissions. In excel, the experience was worse, as the participants had to perform $N$ steering tasks. As a result, in Excel, \cmpB task submission times were very high compared to \cmpA tasks (see  Figure~\ref{fig:timeBox}). In addition, the accuracies of the \cmpB task in Excel were extremely low.}

%The completion times were faster in \noahthan in Excel, irrespective of the comparison tasks and datasets. \emph{\noah offers three additionalbenefits apart from avoidance of steering that may have contributed to such improvement}:
%a) participants used the navigation features to access and compare different subsets of data quickly, 
% b) participants did not need to reissue the aggregation formula for any of the bins they navigated to, and 
%c) participants used the value bars presented along with the results in the aggregate column to visually compare different subsets.
%For \cmpB tasks, the number of subsets to be compared was  higher in birdstikes (50) compared to Airbnb (16). As a result, participants exhibited lower accuracy for birdstrikes when using Excel. In \noah, all the participants first split all the bins to create $N$ bins each corresponding to one state. Then participants panned across the overview to find the desired bin as they compared the values. Even in \noah, comparisons between multiple values resulted in increased visual discontinuity leading to some ($N=4$ out of 20) incorrect submissions.

### 2019-10-10 study.tex
Initial Overleaf Import

\section{Evaluation Study Design}
\label{sec:study}

In this section, we present the design of a user study
to evaluate whether 
\noah helps address spreadsheet navigational challenges.
\subsection{Study Design and Participants}
\label{sec:study_design}
\saj{\noah was designed to address the navigational challenges in spreadsheets. As such, the goal of the study was to evaluate whether \noah enhances the navigational capabilities of traditional spreadsheets. Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah affects a spreadsheet user's navigation experience. In this study, we compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain. Similar domain specific evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or 
tree navigation~\cite{beard1990navigational}. As explained in Section~\ref{sec:related}, the goals of spreadsheets and the TDA tools are quite different. Therefore, we did not consider TDA tools for the comparative study.}
Our study was designed to answer the following questions:

\squishlist
\item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
\saj{How did the addition of an overview like \noah affect a spreadsheet users navigation performance?}
\item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
\saj{How did \noah and it's components impact spreadsheet users'
navigational experiences?}
%\item {\bf RQ3.} \sout{How does bin customization help users inject domain knowledge in addressing their specific exploration questions compared to the automated overview?} \sout{\saj{How do users utilize additional features of \noah, \eg bin customization, during navigation?}
%\item {\bf RQ4.} \sout{How is the user's subjective satisfaction with \noah compared to traditional spreadsheets?}}  \sout{\saj{How is the user's subjective satisfaction with \noah compared to the traditional spreadsheets when performing navigational tasks?}}
\squishend

\stitle{Study Design.} We conducted a $2 \times 2$ (2 datasets, 2 tools) mixed design within-subject study. The two tools used in the study were: Microsoft Excel, and \noah integrated within {\scshape DataSpread}~\cite{dataspread}. \saj{We chose Excel for our comparative study because it is the most popular spreadsheet in use today.}
The study consisted of three phases: (a) an introductory phase explaining
the essential features of \noah via a video tutorial, followed by a warm-up session where participants explored a flight dataset~\cite{web:flight} in \noah 
to familiarize themselves with its features,
(b) a quiz phase where the participants first used both the tools
to perform targeted tasks on two different datasets (described later) 
followed by a survey to provide feedback on their impressions about both Excel and \noah, and
(c) a semi-structured interview to collect qualitative data regarding the quiz phase.

\stitle{Datasets.} We used two datasets---the birdstrikes (used in Keshif~\cite{yalccin2018keshif} and Voyager~\cite{wongsuphasawat2016voyager}), and the Airbnb~\cite{web:airbnb} datasets. These datasets were chosen for their understandability to a general audience. The birdstrikes dataset records instances of birds hitting aeroplanes in different US states. The dataset has 10,868 records and 14 attributes (eight categorical, one spatial region, one temporal, four numeric). 
The Airbnb dataset was larger than the birdstrikes dataset. To ensure a fair comparison across tools, we created a sampled version of the original Airbnb dataset with 10,925 records, by uniformly sampling $10\%$ of the records from each US city. This dataset contained 15 attributes (six categorical, two spatial region, one temporal, six numeric).

\stitle{Participants.} We recruited 20 participants (11 female, 9 male) via flyers across the university and via a university email newsletter. 
The average age of the participants was $31.06$ years ($\sigma = 12.44$). 
The participants came from different backgrounds, 
\eg engineering (seven), business (five), administration (five), and natural science (three). 
During recruitment, 
prospective participants filled out an interest form
where they answered questions about their spreadsheet expertise, 
types of spreadsheet tasks performed, and usage of spreadsheet operations. 
Participants were asked to rate their expertise with
different spreadsheet software, \eg Excel and Google Sheets, 
and their frequency of using various spreadsheet 
tasks \eg data management, data analysis, statistical modeling, and what-if analysis.
We also asked participants about their familiarity 
with basic mathematical and statistical spreadsheet functions
as well as advanced operations, \eg pivot table,
 \code{SUBTOTAL}, and conditional formatting. 
To ensure that participants’ experience with spreadsheets 
didn’t affect their performance during the quiz phase, 
we only recruited participants who rated their experience 
with Excel to be greater than four on a scale of one (no expertise at all) to five (very experienced). 
The selected participants were familiar with performing 
various tasks on spreadsheets, \eg maintaining, tracking, and
analyzing, making predictions, and performing comparisons. 
All of the participants were familiar with the basic mathematical
and statistical functions supported by Excel. 
Each participant received \$10 per hour at the end of their session.

\subsection{Study Procedure}
\label{sec:procedure}
We now explain each of the phases of our study in more detail.

\stitle{Phase 1: Introduction to \noah.}
We began the study by showing a six-minute video tutorial explaining the features of \noah on a dataset of all the flights across the US for January 2018~\cite{web:flight}. The participants then explored the same dataset using \noah to familiarize themselves with the tool for about 10 minutes. The quiz phase began as soon as the participants finished their exploration. \toappendix{Note that we recruited only experienced Excel users for the study and therefore, we did not provide any introduction to Excel. }

\begin{table}[!htb]
\scriptsize
 \vspace{-10pt}
\caption{Quiz tasks for the birdstrikes dataset.}
\label{tab:questions}
\centering
\begin{tabular}{l l}
\hline
\multicolumn{1}{c}{Category} & \multicolumn{1}{c}{Question}    \\ \hline
steer                & Organize the data by State. How many flights that \\
                         & had damages (damage = 1) originated from Florida? \\ 
identify     & How many flights in the currently visible \\
                 &  spreadsheet window have damages?                                    \\ 
steer                & How many flights that had damages (damage = 1) \\
                         &   originated from California?                          \\ 
\cmpA & Which state between Florida and California \\
                         &   has a higher number of flights with damages?    \\ 
\cmpB & Find the state with the most birdstrike occurrences.  \\ 
customize       & Organize the data by \emph{altitude}. What is the average \\
                         &   cost of damages for altitude bin 0-450?          \\ \hline
\end{tabular}
\end{table}

\stitle{Phase 2: The Quiz Phase.}
\saj{The purpose of the quiz phase was to evaluate the effectiveness of the features supported by \noah in addressing spreadsheet limitations}. During the quiz phase, 
each participant performed specific tasks on the two datasets in two sessions, using Excel for one and \noah for the other. 
Each session was followed by a survey, described later. We
alternated the order of the datasets between consecutive participants. 
The order of the tools was alternated between every two participants. 
We developed an online JavaScript-based quiz system that recorded user responses and submission times. 
We also recorded the participants’ interactions
with both tools using screen capture software. 
Participants were informed that they can refer to the Internet for help as many times as they wanted. However, due to their familiarity with Excel, none of the participants required external help. 
For reference, we also provided a printed handout to the participants 
that contained screenshots with the features of \noah.

\emph{Quiz Tasks.} \saj{We designed six tasks grouped into four categories: 
steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} uses cases \code{summarize}, \code{compare}, \code{identify}, and \code{create}, respectively. All of the tasks except the customize task, are representative of the navigation interactions required for the most frequently issued spreadsheet formulae
~\cite{bradbard2014spreadsheet, lawson2009comparison}. The customize tasks was designed to mimic the scenario where a user would group data subsets of interest and required participants to utilize the bin customization feature. The tasks were presented in the same order as shown in Table~\ref{tab:questions} for the birdstrikes dataset. The questions for the Airbnb dataset were similar. The order of the questions mimics the spreadsheet navigation scenario presented in Section~\ref{sec:usage}.}
%All of these tasks require navigation of the spreadsheet data in the form of either scrolling or steering.

\saj{The steer tasks required participants to use the aggregate column feature in \noah as opposed to steering in Excel. The identify task required participants to interact with the detailed view and relate the aggregate result of the first steer task with the raw spreadsheet data. The \cmpA task asked the participants to compare the results of the first steer task (\eg Florida) with that of the second steer task (\eg California) which would require them to use the context bar to revisit a previously visited bin. The \cmpB task involved comparing statistics of all the data subsets. The purpose of the \cmpB task was to see how increasing the number of subsets would affect navigation. As explained earlier, the customize task was designed to evaluate the utility of bin customization feature.}

%As participants use both \noah and Excel to perform these tasks, by examining participants' task performance, we can understand a) how the addition of an overview affect participants' navigation experience (\emph{RQ1}), b) what are the impacts of the individual features of \noah in spreadsheet navigation (\emph{RQ2}), and c) how the participants utilize the features supported by \noah during navigation (\emph{RQ3})

\emph{Survey.} After each session, 
participants rated the corresponding tool used on six metrics: 
confidence, comprehensibility, level of satisfaction, 
ease and speed of use, and ease of learning for spreadsheet navigation, 
on a Likert scale from one (\eg strongly disagree) to seven (\eg strongly agree). 
The survey asked multiple questions related to these metrics, 
15 in total, to ensure reliability. 
Participants were also asked to mention the positive 
and negative aspects of both tools. 
The survey was designed to evaluate \emph{RQ4}.

\emph{Evaluation.} 
We evaluated the accuracy (either 0 or 1) and
completion time for each of the six tasks. 
\saj{We combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be explained with multiple information sources.}
For example, 
we analyzed the video recordings of participants' interaction with the tools during the quiz phase. 
We further analyzed the survey responses to quantify the usability of both the tools.

\stitle{Phase 3: Interview Phase.} 
Following the survey, we conducted a semi-structured interview to 
identify participants’ preferred tools for different tasks 
and to understand the reasoning behind their choices. 
We also asked participants to comment on the usefulness of 
different features provided by \noah and Excel.

\subsection{Study Limitations.}
\kgk{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies. One limitation arises from our participant demographics.  We conducted the study with $20$ participants from four different backgrounds. Our participant pool demographics partially represent the demographics of the general audience intended for \noah.
%is in sample size and participant demographics: we conducted the study with $20$ participants from four different backgrounds.Therefore, our participant pool demographics partially represents the demographics of the general audience intended for \noah.
A larger sample with more participants that better represented the spreadsheet user population would have provided more ecological validity to generalize our findings.
}
%While a larger sample with more diverse backgrounds would have allowed us to perform more definitive quantitative analysis, we combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be explained with multiple information sources.

\saj{Another limitation of the study is that we only compared the performance of a \noah-integrated spreadsheet with a traditional spreadsheet. We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} due to their limitations in the types of interactions supported (see Section~\ref{sec:related}). Moreover, we wanted the participants to utilize the typical spreadsheet operations that they are comfortable with to complete those tasks and see how introduction of \noah affected their navigation experience.}

\saj{Finally, we did not isolate the effects of the individual features of \noah to better understand the implications of those. For example, the effect of binned overview (visual clarity versus visual continuity), display layout (screen space trade-off), and
contextual presentation of data (raw text versus chart representation of aggregate columns). However, the goal of the study was to understand participants' navigation experience in the presence and absence of \noah. A more fine-grained study can be conducted in the future, to inspect the contribution of individual components of \noah in further detail.}

\begin{comment}
\infovis{
\begin{itemize}
    \item Justify the choice of comparing to Excel and note the limitations of that approach (see section 6.1 intro and the {\bf study design paragraph})
    \item Better justify the choice of tasks in the study (see section 6.2 \emph{quiz tasks})
    \item Explain the choice of the four research questions. (see section 6.1 intro)
    \item Provide more details about the study, including reporting intra-participant
    differences (see section 7.2)
\end{itemize}
}
\end{comment}

### 2019-10-10 system.tex
Initial Overleaf Import

--- 2019-10-10 appendix.tex
@@
 and statistical significance results from our user study.
-\section*{\noah Use Cases}
+\section{\noah Use Cases}
 As mentioned in Section~\ref{sec:design}, the goal of designing 
@@
 Table~\ref{tab:scope} captures all three levels and highlights where \noah can be used.
+
+\begin{table}[!htb]{\scriptsize}
+\caption{Uses case of \noah.}
+\label{tab:scope}
+\centering
+\begin{tabular}{c c}
+\hline
+Purpose & Use Cases   \\ \hline
+\emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
+\emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
+\emph{Search}            & \code{browse} (\checkmark), \code{explore (\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
+\emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
+\end{tabular}
+\end{table}
 
@@
 
-\section*{Explaining the Chart Representations}
+\section{Explaining the Chart Representations}
 
@@
 
-\section*{Implementation and  Architecture}
+\section{Implementation and  Architecture}
 \label{sec:sys}
@@
 
-\subsection*{System Architecture}
+\subsection{System Architecture}
 
@@
 
-\section*{Results}
+\section{Results}
 In this section, we evaluate the statistical significance of the task performance results and survey responses presented in Section~\ref{sec:results}. 
-\begin{figure}
-    \centering
-    %\vspace{-10pt}
-    \includegraphics[trim=0 57 0 145,clip,width=\linewidth]{images/userinfo-ip-1.pdf}
-   \caption{Intra-participant submission time differences between \noah and Excel across the quiz tasks.}
-   \label{fig:ipdiff}
- \end{figure}
-
-\subsection{Intra-participant differences.}
-Figure~\ref{fig:ipdiff} depicts the intra-participant submission time differences between \noah and Excel, across all the quiz tasks.
-For a given task and a participant, the corresponding circle denotes by what percentage (between $0.30\%$ to $94.33\%$) a system is faster than the competing system.
-The color of the circle denotes which system was faster (green: \noah faster, yellow: Excel faster).
-The larger the circle the faster the corresponding system is.
-For example, participant $P9$'s submission time was $91.78\%$ faster with \noah for the \cmpB task and $72.62\%$ faster with Excel for the \cmpA task.
-Figure~\ref{fig:ipdiff} confirms that majority of the submission times using \noah were faster compared to Excel---nineteen out of the twenty participants completed at least four tasks in less time using \noah compared to Excel.
-The submission time difference is even more apparent for the steer, identify, and \cmpB tasks.
-
-For \cmpA, customize, and the second steer task, less than a third of the participants' submission times were faster using Excel. 
-For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah, utilized the \emph{autosum} shortcut for the birdstrikes dataset.
-For the \cmpA tasks, a number of participants ($N=3$) that required more time to submit answer using \noah, first used the bin customization feature to view all the unique bins which contributed to the higher submission time compared to Excel.
-On the other hand, as explained earlier, the unfamiliarity with the bin customization feature contributed to higher submission time using \noah for six participants (yellow circles).
-
 \subsection{Statistical Significance: Task Performance}

### 2019-10-10 letter.tex
small changes to the letter, testing the overleaf github sync

@@
-\section*{Resubmission of the InfoVis 2019 submission titled ``Extending Spreadsheets to Support Seamless
+%!TEX root = main.tex
+
+\section*{Resubmission of InfoVis 2019 submission ``Extending Spreadsheets to Support Seamless
 Navigation at Scale'' to TVCG}
@@
 A previous version of this paper was submitted to InfoVis 2019.
-This version was rejected as the paper co-chairs deemed that revising the paper would require more time than allocated to meet reviewer expectations for the conference.
+This version was rejected as the paper co-chairs deemed that revising the paper would require more time than allocated to meet reviewer expectations.
 The paper co-chairs mentioned the following: ``\textit{We particularly encourage such revisions where submissions were positively received by reviewers, but the revisions required were deemed to be beyond the scope of the conference review cycle.
-If you address the issues raised and subsequently submit to IEEE TVCG, please make reference to this IEEE InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}''
+If you address the issues raised and subsequently submit to TVCG, please make reference to this  InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}''
 
@@
 The entire set of reviews are attached.
-In particular, the reviewers asked for the following changes for a revision: \begin{itemize}
+In particular, the reviewers asked for the following changes for a revision:
+\begin{itemize}
     \item More clearly articulate the merits over alternative tools, such as Keshif. 
@@
 \end{itemize} 
-
-Here, we provide a response to the reviews. In our revised submission,
-\textbf{(i) we discuss the merits of introducing a spreadsheet integration like \noah over alternative tools, \eg Keshif;
-(ii) we explain our design choice surrounding the binning mechanism;
+We have spent the last four months 
+preparing a revised version of this paper.
+\agp{One thing that is confusing is that the bulletpoints above
+don't neatly map to the bulletpoints below, or the enumerated list after that.}
+Here, we provide a response to the reviews. 
+In our revised submission,
+(i) we discuss the merits of having an interface like \noah
+integrated within an existing spreadsheet, as opposed
+to other general-purpose exploratory tools such as  Keshif;
+(ii) we justify our design choice surrounding the binning mechanism;
 and 
-(iii) we enrich the user study design section further by providing justification for comparing \noah and Excel, rephrasing our research questions reflecting the purpose of the user study, explaining the choice of quiz tasks,
-iv) we reorganize the results section to  highlight key contributions of \noah while explaining participant experiences in using the tool.
-Moreover, we report the  intra-participant differences for the study}.
+(iii) we enrich the user study design section further 
+by providing justifications for comparing \noah and Excel, 
+rephrasing our research questions reflecting the purpose of the user study, explaining the choice of quiz tasks,
+(iv) we reorganize the results section to  highlight key contributions of \noah while explaining participant experiences in using the tool.
+Moreover, we report the  intra-participant differences for the study.
 The major changes to the paper include:

### 2019-10-10 introduction.tex
midway through intro and letter

@@
+%!TEX root = main.tex
+
 %\IEEEraisesectionheading{\section{Introduction}}
@@
 and analysis of data by end-users~\cite{excel-users}. 
-\saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions with more advanced analytical features~\cite{chan1996use,raden2005shedding}.
+\saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
+with more advanced analytical features~\cite{chan1996use,raden2005shedding}. 
 One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.}
@@
 via two unit operations, scrolling and steering.
-\saj{Scrolling is the action of moving displayed text or graphics up, down, or across on a computer screen in order to view different parts of the spreadsheet.}
+\saj{Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.}
 For example, when analyzing data, users may scroll
@@
 
-However, navigating spreadsheet data using scrolling or steering
-is often challenging, even for moderately sized spreadsheets that only span
-a few screens~\cite{network-context,nardi1990spreadsheet}, due to multiple inter-related reasons:
+\new{
+However, navigating spreadsheets using scrolling or steering
+is challenging, since spreadsheet data span multiple screens,
+making it hard to synthesize, analyze, makes sense of, or
+operate on it~\cite{network-context,nardi1990spreadsheet}.
+With the ease of data generation, and with spreadsheets
+now supporting increasingly larger datasets, \eg Google Sheets
+now supports two million cells~\cite{}, a five-fold increase
+from the previous limit of 400K cells, 
+{\em navigating
+data within spreadsheets is only becoming even harder},
+thanks to
+multiple inter-related reasons: 
+}
 \squishlist
@@
 
-Overall,  while navigating present-day spreadsheets, {\bf \em users often lose context,
+\noindent
+Overall,  while navigating present-day spreadsheets, 
+{\bf \em users often lose context,
 get overwhelmed, and experience visual discontinuities.} 
 Addressing these challenges requires considerable manual effort.
-As we argue in Section~\ref{sec:related},
-existing spreadsheet functionalities such as
+As we \new{will} argue in Section~\ref{sec:related},
+existing spreadsheet features such as
 pivot tables, named ranges, and subtotals,
-while providing partial support for some of these challenges
-(\eg pivot table generates a summary
-while losing the mapping between the
-raw data and the summary; named ranges requires users
-to manually associate names with ranges of data),
-do not eliminate them entirely.
-With the ease of data generation, and with spreadsheets
-now supporting data
-of increasingly larger scales---\eg Google Sheets
-now supports two million cells, a five-fold increase
-from the previous limit of 400K cells---navigating
-data within spreadsheets will become harder, not easier.
+\new{partially alleviate 
+some of the aforementioned 
+challenges but do not eliminate them entirely.
+For example, pivot tables generate a summary
+while losing the correspondence between the
+raw data and the summary, while named ranges require users
+to manually associate names with ranges of data.}
 
@@
 an overview of the overall structure of the data along
-with the spreadsheet~\cite{grudin2001partitioning}.
-The resulting integration would then be a classical
+with the spreadsheet~\cite{grudin2001partitioning}
+\new{resulting in} a classical
 overview+detail interface
@@
 to facilitate navigation in various domains such as 
-text editors and maps~\cite{cockburn2009review},
-and enable users to manipulate
+text editors and maps~\cite{cockburn2009review}.
+\new{Users can manipulate
 the overview or detailed view,
 to perform high-level or low-level operations,
-respectively.
+respectively.}
 Overview+detail interfaces
@@
 reducing cognitive load for users by
-providing the big picture first,
-helping them quickly assimilate the information space~\cite{cockburn2009review}.
+providing \new{them} the big picture first,
+helping them quickly assimilate 
+the information space~\cite{cockburn2009review}.
+\agp{got until here}
 \saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}

### 2019-10-10 letter.tex
midway through intro and letter

@@
 \agp{One thing that is confusing is that the bulletpoints above
-don't neatly map to the bulletpoints below, or the enumerated list after that.}
+don't neatly map to the bulletpoints below, or the enumerated list after that. Stopped reading after this}
 Here, we provide a response to the reviews. 

### 2019-10-10 introduction.tex
some more intro done

@@
 \label{sec:intro}
-With a user base of more than one-tenth of the world's population,
-spreadsheets are by far the most popular medium for ad-hoc exploration
-and analysis of data by end-users~\cite{excel-users}. 
+\new{With a user base of more than  one-tenth of the world's population,
+{\bf \em spreadsheets are by far the most popular medium for ad-hoc exploration
+and analysis of data}~\cite{excel-users}.} 
 \saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
@@
 the information space~\cite{cockburn2009review}.
-\agp{got until here}
-\saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}
+\new{Our goal is to {\em integrate an overview plug-in 
+with spreadsheets} that captures
+the overall structure of the data,
+while supporting interactions that address
+the difficulties in typical navigational 
+operations like scrolling and steering.
+{\bf \em It is essential
+that our interface is a plug-in that 
+enhances the capabilities of spreadsheets
+that so many users
+are used to and reliant on,
+as opposed to a potentially jarring or confusing 
+replacement for spreadsheets.}
+}
+% \saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}
 
@@
 \item {\em Overview modality.}
-\saj{One approach for an overview is to simply add a
+\saj{
+One approach for an overview is to simply add a
 zoomed out version of the entire sheet as a pane on the side,
-as in popular presentation software applications like Microsoft PowerPoint or text editors
+as in popular presentation software like 
+Microsoft PowerPoint or text editors
 like Sublime Text. 
 Unfortunately, this approach
-would not suffice for a spreadsheet---while it would help users make
-somewhat more informed
-scrolling decisions, the zoomed out presentation of the overview would be unreadable. An overview should provide a comprehensible big picture view of the spreadsheet.
+would not suffice for a spreadsheet.
+\new{An overview should provide a
+comprehensible big picture view:
+for a spreadsheet, when zooming out beyond a point, 
+the overview would be effectively 
+unreadable.}
 Another approach, adopted by map tools like the
@@
 zooming into and out of a detailed view consisting of
-this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users sought more fine-grained or coarse-grained view of the overall structure of the data.}
+this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users seek a more fine-grained or coarse-grained view of the overall structure of the data.}
 

### 2019-10-10 introduction.tex
more:x

@@
 zooming into and out of a detailed view consisting of
-this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users seek a more fine-grained or coarse-grained view of the overall structure of the data.}
+this raw data is not meaningful. 
+\new{How do we design an overview to dynamically 
+change as users 
+seek a more fine-grained or coarse-grained view 
+of the overall structure of the data?} 
+}
 
@@
 with many rows, 
-one approach to constructing a dynamic overview is mapping rows of data to high level groups similar to online maps.
-In online maps, cities are grouped into states and states are grouped into countries forming a multi-granularity hierarchy.
-How do we automatically group spreadsheet rows together in a similar ``meaningful''
+one approach to constructing a dynamic overview 
+is mapping rows of data to high level groups, 
+similar to online maps. 
+In online maps, cities are grouped into states 
+and states are grouped into countries, 
+forming a multi-granularity hierarchy. 
+How do we automatically group spreadsheet 
+rows together in a similar ``meaningful''
 way
 such that this grouping applies to all 
-data types, including strings and numbers? If the automatic grouping is not meaningful, how do we allow the users
-to customize the overview
-to construct meaningful groups? How do we facilitate interactions that enable users view the overview at multiple granularities?}
+data types, including strings and numbers? 
+If the automatically generated grouping is 
+not semantically meaningful, how do we allow the users
+to customize the grouping modality? 
+How do we facilitate interactions that enable 
+users to view the overview at multiple granularities?}
 
 \item {\em Operations on the overview.}
-\saj{Following the construction of a dynamic overview, the next challenge is to design simple interactions that achieve similar outcomes as scrolling and steering.
-For example, an alternative to scrolling can be to leverage the groups of the overview to access the rows mapped to that group.
-How do we leverage the overview so that users can perform similar simple interaction to steer spreadsheet data and issue formula?
-As the granularity of the dynamic overview changes, how do we efficiently update the mapping from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless?
+\saj{Following the construction of a dynamic overview, 
+the next challenge is to design simple interactions 
+that achieve similar outcomes as scrolling and steering.
+For example, an alternative to scrolling 
+can be to leverage the groups of the overview 
+to access the rows mapped to that group. 
+Similarly, how do we leverage
+the overview to support simple interactions
+to steer spreadsheet data (i.e., select a range of data)
+for formula computation?
+As the granularity of the dynamic overview 
+changes, how do we efficiently update the mapping 
+from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless? 
 How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch? 
-How do we ensure such automatic recomputation scales with the size of the data?}
+How do we ensure such automatic recomputation 
+scales with the size of the data?\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}}
 
@@
 should be generic so that it can be integrated
-with any existing spreadsheet tool on any spreadsheet
+with any existing spreadsheet tool on any 
 dataset without impacting
@@
 \squishend
+\agp{A general comment: can we ensure that either we use
+questions for all of the items above or none of them for uniformity?}
 
@@
 
-\stitle{\noah.} We address the aforementioned challenges
-in our tool \noah,
+\stitle{\noah: \new{an navigation plug-in for spreadsheets}.} We address the aforementioned challenges
+in \noah,
 an in-situ
@@
 \saj{a multi-granularity overview is constructed and displayed
-within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a).
-Users can zoom into or out of the overview to obtain a fine-grained or coarse grained perspective of the data distribution.
-The distribution at each granularity is captured by a histogram,
-enabling users to assimilate the data. Each bin (group) of the histogram is mapped to a collection of rows in the spreadsheet.} Cumbersome scrolling operations are eliminated in favor
+within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a \new{fine or coarse-}grained 
+perspective of the data distribution. 
+The distribution at each granularity 
+is captured by a histogram,
+enabling users to assimilate the data \new{via summary
+statistics}. 
+Each bin (group) of the histogram is 
+mapped to a collection of rows in the spreadsheet.} 
+Cumbersome scrolling operations are eliminated in favor
 of a few clicks on the overview interface. Instead of steering to analyze the data,
@@
 tasks involving navigation 
-correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
+correctly and quickly in \noah.
+Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} \agp{mention how these tasks were representative or general}
 \squishend

### 2019-10-10 introduction.tex
split into usecases

@@
 Finally, we discuss the limitations of \noah and possible future enhancements in Section~\ref{sec:discuss}.}
-
-\section{\noah Use Cases}
-\label{sec:usage}
-\saj{As explained in Section~\ref{sec:intro}, users prefer spreadsheets over enterprise solutions to view, directly manipulate, and explore data~\cite{chan1996use,raden2005shedding}.
-To identify the scope of these tasks on spreadsheets (see Table~\ref{tab:scope}), we draw parallel to the typology of abstract data exploration tasks~\cite{brehmer2013multi}.
-The typology characterizes the exploration and manipulation tasks performed on visual representations of data.
-While all the tasks in Table~\ref{tab:scope} can be performed using spreadsheets, \noah enhances the experience for a subset of these tasks, indicated by a checkmark (\checkmark).
-We explain how \noah complements spreadsheets in accomplishing these tasks in Appendix.}
-
-\begin{table}[!htb]\scriptsize
-\caption{Uses cases of \noah.}
-\label{tab:scope}
-\centering
-\begin{tabular}{c c}
-\hline
-Purpose & Use Cases   \\ \hline
-\emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
-\emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
-\emph{Search}            & \code{browse} (\checkmark), \code{explore (\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
-\emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
-\end{tabular}
-\end{table}
-
-\saj{We now describe a usage scenario that captures the spreadsheet exploration tasks accommodated by \noah (Table~\ref{tab:scope}) while illustrating the benefits of integrating an overview with spreadsheets.
-Let's assume that} Rebecca, an investigative journalist
-is exploring the Inside Airbnb dataset~\cite{web:airbnb},
-a dataset of all the Airbnb listings across different US cities.
-This dataset was created to investigate the long-standing
-accusation that many listings in Airbnb are illegally
-run as hotel businesses, while avoiding taxes;
-any listing available for rent for more than $60$ days
-a year is considered to be operated as a hotel~\cite{accusation}.
-
-Without any prior knowledge of the dataset,
-Rebecca starts by finding out information
-about cities in the dataset and their corresponding listings;
-for example, which cities are present,
-and roughly how many listings does each city have \saj{(\code{browse} and \code{explore})}.
-Without \noah, Rebecca may have used
-a pivot table (discussed in Section~\ref{sec:related})
-to construct
-a summary---however, since this summary is disconnected
-from the underlying data, it is hard for Rebecca to
-map the summary statistics to
-the raw data to obtain further details
-about listings from any given city.
-If she wanted to identify listings from a specific city \saj{(\code{lookup} and \code{locate} or \code{identify})},
-Rebecca would have to either use search capabilities or
-perform an explicit filter for this information,
-and would have to switch back and forth between the
-pivot table results and the raw listings,
-present at disparate locations
-on the spreadsheet.
-Even at the first step of this exploration,
-Rebecca would experience substantial cognitive burden,
-loss of context,
-and visual discontinuities,
-with subsequent steps becoming progressively
-more challenging.
-
-Using \noah, she explores
-the data by city,
-with \noah providing a high-level overview of
-cities (Figure~\ref{fig:ux}a).
-The overview consists of sorted non-overlapping bins
-containing one or more cities.
-She can click on any bin and the corresponding
-data will be displayed at the top of her screen.
-For example, clicking on the {\em Ashville-Boston} bin
-displays the Ashville listings (Figure~\ref{fig:ux}c).
-She can also zoom into bins using the ``\textcolor{blue}{$\rangle$}'' arrows,
-zoom out of bins using the ``\textcolor{blue}{$\langle$}'' arrows,
-and pan by clicking on various bins at the same level.
-We discuss the construction of the overview
-and associated interactions in Section~\ref{sec:ui}.
-
-Next, say Rebecca wants to analyze
-one of the larger cities \saj{(\code{query}/\code{summarize}) to understand the renting pattern}.
-She identifies these cities by
-comparing the number of listings
-for each city displayed on the overview \saj{(\code{compare})}.
-She decides to focus on Boston, her hometown,
-and wants to find out how many listings
-in Boston violate
-the ``rent availability $>60$ days'' condition.
-In a typical spreadsheet, Rebecca needs to manually steer and then select
-the Boston listings as input to a \code{COUNTIF} formula that counts the number of
-rows that satisfy the above mentioned condition.
-Using \noah, she can zoom into
-the {\em Ashville-Boston} bin
-(Figure~\ref{fig:concept}a and~\ref{fig:concept}b)
-and then issues
-a \code{COUNTIF} operation on the overview.
-%\ie a \code{COUNTIF} formula\footnote{\code{COUNTIF} is a
-%type of formula in spreadsheets that counts the number %of
-%rows that satisfy some condition.}. 
-The result is displayed as an {\em aggregate column}
-alongside the overview (Figure~\ref{fig:ux}b).
-Rebecca learns \saj{(\code{discover})} that more than half of the listings in Boston are 
-effectively operating as hotels---a large number!
-
-Based on this insight, Rebecca then wants to
-understand availability statistics for an even
-larger city, Chicago \saj{(\code{summarize})}.
-As she uses the overview to navigate
-to Chicago, \noah
-automatically updates the aggregate column to
-the \code{COUNTIF} formula results for Chicago,
-without Rebecca needing to reissue it by performing a cumbersome steering operation as in traditional spreadsheets. 
-Rebecca learns that Chicago exhibits 
-a similar renting pattern as Boston, with more than half the listings operating as hotels. She can then check if this trend holds for all large cities \saj{(\code{compare})},
-and whether the smaller cities have a different pattern \saj{(\code{discover})}. \saj{Note that, the rows that satisfy the ``rent availability $>60$ days'' condition, are listed
- in the spreadsheet adjacent
-to the overview in sky blue (Figure~\ref{fig:ux}g)}. With the raw 
-data presented side-by-side, she can also dive into other attributes
-of the listings operating as hotels to see if there are any other identifying characteristics,
-\eg if they are all managed by a small number of agencies acting as individual renters \saj{(\code{identify})}.
-
- Finally, as Rebecca navigates the data,
- her navigation history (Figure~\ref{fig:ux}d),
- \ie recently visited cities, and current navigation path
- (Figure~\ref{fig:ux}e) are kept up-to-date,
- allowing her to maintain context during navigation.
- She can revisit any previously visited cities
- by simply clicking on the relevant path
- in the navigation history.
-
-Overall, with \noah, users can quickly
-comprehend the data via the overview, access any region 
-within the data without having to
-scroll endlessly, and request additional details on demand without having to
-steer across multiple screens. As users navigate and analyze the data, 
-they can revisit previously accessed data via the navigation history, not losing context
-of what they have explored.

### 2019-10-10 usecases.tex
split into usecases

--- 2019-10-10 old-introduction.tex
@@
-%\IEEEraisesectionheading{\section{Introduction}}
-\section{Introduction}
-\label{sec:intro}
-With a user base of more than one-tenth of the world's population,
-spreadsheets are by far the most popular medium for ad-hoc exploration
-and analysis of data by end-users~\cite{excel-users}.
-\saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions with more advanced analytical features~\cite{chan1996use,raden2005shedding}.
-One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.}
-Spreadsheets enable users to view, structure, and present data
-in an intuitive tabular layout, wherein users can map their data and tasks;
-this tabular layout is essential to
-the popularity of spreadsheets~\cite{nardi1990spreadsheet}.
-
-Using this tabular layout effectively involves
-navigation, \ie
-{\em ``the process of viewing and manipulating the computer display to
-show another portion of the information space''}~\cite{nav-study}.
-Navigation is supported
-via two unit operations, scrolling and steering.
-\saj{Scrolling is the action of moving displayed text or graphics up, down, or across on a computer screen in order to view different parts of the spreadsheet.}
-For example, when analyzing data, users may scroll
-to compare data across different screens, or to get
-a high-level view of the overall spreadsheet.
-\saj{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
-users may steer to select
-the subset of the data
-to be operated on as an argument within the formula.
-Most frequently used spreadsheet formulae
-require users to perform steering actions~\cite{bradbard2014spreadsheet, lawson2009comparison}.
-Overall, both scrolling and steering are crucial as
-users navigate spreadsheets to
-identify, compare, and summarize data.
-
-\begin{figure*}
-     \centering %trim=left bottom right top
-  %\includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/navigation.pdf}
-  \includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/screenshot.pdf}
-    \caption{\noah: navigation interface consisting of (a) a zoomable overview and (b) an aggregate column integrated with (c) a spreadsheet.
-A context bar consisting of (d) a navigation history displaying locations visited so far using the overview, and (e) a breadcrumb showing the current navigation path (\eg Home).
-(f) The user’s current focus in the spreadsheet is highlighted on the overview.
-(g) Columns corresponding to the navigation attribute (city) and aggregate column (availability) are highlighted on the spreadsheet.}
-  \label{fig:ux}
-\end{figure*}
-
-However, navigating spreadsheet data using scrolling or steering
-is often challenging, even for moderately sized spreadsheets that only span
-a few screens~\cite{network-context,nardi1990spreadsheet}, due to multiple inter-related reasons:
-\squishlist
-\item {\em Loss of overview and context.}
-When navigating
-spreadsheets, users can easily lose the context
-of where they are and where they should go next~\cite{network-context}.
-The only navigational context provided by spreadsheets
-is the built-in scrollbar that acts as a one-dimensional
-overview and indicates the user's current location on the sheet.
-However, since this overview does not capture the
-layout and structure of the data,
-users are forced to mentally assimilate
-the layout and recall it on-demand, as they navigate a spreadsheet.
-
-\item {\em Cognitive and mechanical burdens.}
-The lack of contextual cues leads to severe
-cognitive and mechanical burdens
-for users~\cite{cockburn2009review}.
-Users often end up taking their own drastic measures to
-avoid getting lost; for example,
-some users create personalized overviews
-extrinsic to the spreadsheet,
-by sketching maps
-of spreadsheets on paper~\cite{network-context}.
-Other users add their own landmarks such as headers or
-colored cells, as a visual affordance to assist
-in navigation~\cite{network-context}.
-Steering via dragging the mouse pointer across multiple screens to select
-a subset of data as input to a formula
-can often be challenging as well:
-the only remedy is for users to abandon steering entirely and
-instead remember the range of
-the subset of data of interest, and then correctly enter this range
-as the argument to the formula, often giving rise to errors that are increasingly prevalent in spreadsheets~\cite{panko1998we}.
-
-\item {\em Visual discontinuities.}
-The limited viewport afforded to the user
-introduces a visual discontinuity between the
-information being displayed.
-For example, comparing spatially separated subsets
-of data within the spreadsheet
-requires moving back and forth between multiple viewports,
-which can be overwhelming~\cite{nardi1990spreadsheet,network-context}.
-As an alternative, users tend to copy subsets of data side by side to
-reduce the visual discontinuity~\cite{nardi1990spreadsheet,network-context},
-which is cumbersome.
-\squishend
-
-Overall,  while navigating present-day spreadsheets, {\bf \em users often lose context,
-get overwhelmed, and experience visual discontinuities.} 
-Addressing these challenges requires considerable manual effort.
-As we argue in Section~\ref{sec:related},
-existing spreadsheet functionalities such as
-pivot tables, named ranges, and subtotals,
-while providing partial support for some of these challenges
-(\eg pivot table generates a summary
-while losing the mapping between the
-raw data and the summary; named ranges requires users
-to manually associate names with ranges of data),
-do not eliminate them entirely.
-With the ease of data generation, and with spreadsheets
-now supporting data
-of increasingly larger scales---\eg Google Sheets
-now supports two million cells, a five-fold increase
-from the previous limit of 400K cells---navigating
-data within spreadsheets will become harder, not easier.
-
-So, how do we support more effective navigation of data
-within spreadsheets?
-One approach would be to try to integrate
-an overview of the overall structure of the data along
-with the spreadsheet~\cite{grudin2001partitioning}.
-The resulting integration would then be a classical
-overview+detail interface
-where the spreadsheet is the detailed view.
-Overview+detail interfaces are used
-to facilitate navigation in various domains such as 
-text editors and maps~\cite{cockburn2009review},
-and enable users to manipulate
-the overview or detailed view,
-to perform high-level or low-level operations,
-respectively.
-Overview+detail interfaces
-have been shown to be effective in these domains,
-reducing cognitive load for users by
-providing the big picture first,
-helping them quickly assimilate the information space~\cite{cockburn2009review}.
-In this paper, our goal is to extend
-existing spreadsheets
-with an overview plug-in, giving an overview+detail interface.
-
-However, while an overview plug-in for spreadsheets
-does seem appealing and natural,
-developing it leads to several challenges.
-\squishlist
-\item {\em Overview modality.}
-One approach for an overview is to simply add a
-zoomed out version of the entire sheet as a pane on the side,
-as in popular presentation software applications like Microsoft PowerPoint or text editors
-like Sublime Text. 
-Unfortunately, this approach
-would not suffice for a spreadsheet---while it would help users make
-somewhat more informed
-scrolling decisions, the zoomed out presentation of the overview would be unreadable.
-Another approach, adopted by map tools like the
-early versions of Google maps,
-is to use the overview
-to provide a \saj{global} context of the user's current location
-\saj{being displayed} in the \saj{zoomed in} detailed view~\cite{cockburn2009review}.
-Users can perform zooming operations on the detailed
-view as the overview remains static.
-Since spreadsheets already d  isplay the raw data,
-zooming into and out of a detailed view consisting of
-this raw data is not meaningful.
-Instead, the goal of our spreadsheet
-overview is to provide a big picture view of the spreadsheet,
-capturing the
-data distribution and \saj{augmenting it with summarization},
-while enabling meaningful interactions
-for navigating the data.
-This necessitates
-{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
-zooming (unlike text editors or presentation software),
-on the overview view instead of the detailed view
-(unlike map tools).
-Several domains, \eg online maps and program visualization,
-adopt semantic zooming
-to ensure improved readability and
-comprehension of the information displayed~\cite{summers2003experimental}.
-However, to the best of our knowledge, no prior work
-has ever developed semantic zooming for spreadsheet data.
-
-\item {\em Construction of the overview.}
-Given a spreadsheet
-with many rows,
-it is not clear how to employ semantic zooming
-to construct the overview.
-How do we group rows together in a ``meaningful''
-way
-such that this grouping applies to all 
-data types, including strings and numbers?
-Since we want to allow users to control how the overview
-is constructed, we can let them select the grouping
-attribute(s).
-Following this, we need mechanisms to automatically group nearby rows
-into meaningful bins.
-Further, the bins need to be constructed
-at various granularities,
-allowing users to zoom in and zoom out of the overview.
-We also need to allow the users
-to customize the bins
-to construct meaningful categories, while providing
-interpretable and intelligent defaults.
-Customizing these bins will also eliminate
-visual discontinuities: users can simply arrange the bins
-the way they want and keep them near each other
-for easy comparison.
-
-\item {\em Operations on the overview.}
-While the scrolling issues are partially addressed
-by having the overview, the steering issues remain unaddressed.
-How do we leverage the overview to allow users
-to select regions of interest for issuing formulae
-without steering
-across multiple screens?
-As the granularity of the overview changes (during zoom in/out)
-how do we ensure that users don't have to reissue the formula from scratch?
-For this, we take inspiration from aggregation
-in relational databases. In relational databases, each ``group'' is associated
-with one or more aggregates. As long as the groups are
-appropriately constructed (using the mechanisms alluded to above),
-formulae can be issued once without requiring any steering, and
-automatically maintained for all groups when
-zooming in and zooming out.
-
-\item {\em Seamless integration as a plug-in.}
-Finally, the design of the overview
-should be generic so that it can be integrated
-with any existing spreadsheet tool on any spreadsheet
-dataset without impacting
-existing functionalities or look-and-feel.
-Any new interactions introduced should be consistent
-with traditional spreadsheet semantics,
-and complement existing spreadsheet interactions.
-Interactions between the overview and the raw spreadsheet
-should be coordinated to the extent possible.
-
-\squishend
-
-\noindent
-We address the aforementioned challenges
-in our tool \noah,
-an in-situ
-\textbf{\underline{n}}avigation interface
-for \textbf{\underline{o}}verviewing
-and \textbf{\underline{a}}nalyzing
-spreadsheet data \textbf{\underline{h}}olistically.
-\noah is constructed as a plugin to
-an existing spreadsheet tool,
-{\scshape DataSpread}~\cite{dataspread},
-an open-source scalable web-based spreadsheet.
-While \noah's design is not tied to {\sc DataSpread},
-we opted not to use other popular spreadsheet
-tools like Google Sheets and Microsoft Excel because they
-are closed source.
-Figure~\ref{fig:ux} shows a snapshot of \noah.
-When the user chooses to explore
-the data by a specific attribute,
-an overview is constructed and displayed
-within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a).
-The overview is a histogram
-and depicts the overall data distribution
-for the chosen attribute,
-enabling users to assimilate the data.
-\noah ensures that there is coordination between
-the overview and the spreadsheet:
-for example, panning and zooming on the overview
-are reflected on the spreadsheet by displaying
-the spreadsheet data corresponding to the bin
-currently in focus in the overview.
-Cumbersome scrolling operations are eliminated in favor
-of a few clicks on the overview interface.
-% \noah leverages {\scshape DataSpread}'s
-% underlying spatial mapping of data to ensure coordination between the overview and the spreadsheet so that users can relate the interactions on the overview with the raw data.
-Instead of steering to analyze the data,
-users can issue aggregation operations
-on the overview,
-and view results on a separate {\em aggregate column},
-alongside the overview (Figure~\ref{fig:ux}b).
-In this manner,
-users can issue formulae on different
-subsets of the data while remaining on the same screen,
-reducing visual discontinuity.
-Finally, \noah automatically creates
-contextual and historical information
-(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
-while displaying visual cues
-(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
-so that users don't lose context during navigation.
-
-The primary contribution of our work is twofold:
-\squishlist
-\item We formalize the design of a general
-navigation (overview+detail) interface
-for exploration and analysis of large spreadsheets.
-We realize this design in the form of \noah,
-a plugin to an existing web-based spreadsheet tool,
-ensuring that interactions supported by \noah
-complement existing spreadsheet operations.
-\item We conduct a user study to evaluate
-the benefits and limitations of this plugin.
-\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
-tasks involving navigation 
-correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
-\squishend
-
-\toappendix{The rest of the paper is organized as follows: In Section~\ref{sec:usage} we first present a case study of \noah being applied to a real-world scenario.  
-In Section~\ref{sec:related}, we discuss related work.
-We then explain the design considerations for developing \noah in Section~\ref{sec:design}.
-In Section~\ref{sec:ui}, we present \noah’s user interface and supported operations.
-We then discuss the design and outcomes of a user study evaluating the effectiveness of \noah in Section~\ref{sec:study}.
-Finally, we discuss the limitations of \noah and possible future enhancements in Section~\ref{sec:discuss}.}
+%!TEX root = main.tex
 

### 2019-10-11 introduction.tex
intro Aditya comments addressed. ref added for Google sheet

@@
 and analysis of data}~\cite{excel-users}.} 
-\saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
+Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
 with more advanced analytical features~\cite{chan1996use,raden2005shedding}. 
-One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.}
+One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.
 Spreadsheets enable users to view, structure, and present data
@@
 via two unit operations, scrolling and steering.
-\saj{Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.}
+Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.
 For example, when analyzing data, users may scroll
@@
 a high-level view of the overall spreadsheet.
-\saj{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
+Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region. For example, to issue a formula,
 users may steer to select
@@
 now supporting increasingly larger datasets, \eg Google Sheets
-now supports two million cells~\cite{}, a five-fold increase
+now supports five million cells~\cite{gs}, a $12.5X$ increase
 from the previous limit of 400K cells, 
@@
 \item {\em Overview modality.}
-\saj{
 One approach for an overview is to simply add a
@@
 like Sublime Text.
+\newsaj{The zoomed out overview
+displays the entire data 
+at a lower magnification scale.} 
 Unfortunately, this approach
@@
 comprehensible big picture view:
-for a spreadsheet, when zooming out beyond a point, 
-the overview would be effectively 
-unreadable.}
+for a \newsaj{spreadsheet of numbers, text, or formulae}, 
+when zooming out beyond a point, 
+\newsaj{an overview displayed at 
+such low magnification scale} 
+would be unreadable.}
 Another approach, adopted by map tools like the
@@
 of the overall structure of the data?}
-}
 
 \item {\em Construction of the overview.}
-\saj{Given a spreadsheet
+Given a spreadsheet
 with many rows, 
@@
 How do we facilitate interactions that enable 
-users to view the overview at multiple granularities?}
+users to view the overview at multiple granularities?
 
 \item {\em Operations on the overview.}
-\saj{Following the construction of a dynamic overview, 
+Following the construction of a dynamic overview, 
 the next challenge is to design simple interactions 
@@
 to access the rows mapped to that group. 
+\newsaj{As the granularity of the dynamic overview 
+changes, how do we efficiently update the mapping 
+from spreadsheet rows to the 
+finer or coarser groups 
+so that scrolling remains seamless? 
 Similarly, how do we leverage
-the overview to support simple interactions
-to steer spreadsheet data (i.e., select a range of data)
-for formula computation?
-As the granularity of the dynamic overview 
-changes, how do we efficiently update the mapping 
-from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless? 
-How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch? 
-How do we ensure such automatic recomputation 
-scales with the size of the data?\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}}
+the overview to steer spreadsheet data (i.e., select a range of data)
+for formula computation? How do we present the results of
+the formula within the overview? One approach can be to adopt
+the pivot table-like presentation of results. Within their summary view, 
+pivot tables display 
+aggregate formula results (\eg \code{SUM}, \code{COUNT}) 
+alongside each group. 
+However, unlike pivot tables, users can explore the overview at
+multiple-granularities. As the granularity changes,
+the grouping of rows also changes; 
+making the previous formula results 
+inconsistent with the new groups.
+How do we 
+recompute the results of a formula 
+as the granularity changes 
+without requiring the users to 
+reissue the formula from scratch?}
+\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}
 
 \item {\em Seamless integration as a plug-in.}
-Finally, the design of the overview
-should be generic so that it can be integrated
+\newsaj{Finally, how can we 
+design an overview with a generic set of features,
+that can be integrated
 with any existing spreadsheet tool on any 
 dataset without impacting
-existing functionalities or look-and-feel.
-Any new interactions introduced should be consistent
-with traditional spreadsheet semantics,
-and complement existing spreadsheet interactions.
-Interactions between the overview and the raw spreadsheet
-should be coordinated to the extent possible.
+existing functionalities or look-and-feel?
+How do we ensure that the interactions 
+supported by the overview 
+are 
+consistent with
+traditional spreadsheet semantics,
+and complement existing spreadsheet interactions?
+How to we enable coordinated interactions across both views, \ie 
+the overview and the raw spreadsheet, such that they
+remain consistent at any given time?}
 \squishend
@@
 
-\begin{comment}
-Since we want to allow users to control how the overview
-is constructed, we can let them select the grouping
-attribute(s). Following this, we need mechanisms to automatically group nearby rows
-into meaningful bins.
-This necessitates
-{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
-zooming (unlike text editors or presentation software),
-on the overview view instead of the detailed view
-(unlike map tools).
-Several domains, \eg online maps and program visualization,
-adopt semantic zooming
-to ensure improved readability and
-comprehension of the information displayed~\cite{summers2003experimental}.
-However, to the best of our knowledge, no prior work
-has ever developed semantic zooming for spreadsheet data.
-Customizing these bins will also eliminate
-visual discontinuities: users can simply arrange the bins
-the way they want and keep them near each other
-for easy comparison.
-For this, we take inspiration from aggregation
-in relational databases. In relational databases, each ``group'' is associated
-with one or more aggregates. As long as the groups are
-appropriately constructed (using the mechanisms alluded to above),
-formulae can be issued once without requiring any steering, and
-automatically maintained for all groups when
-zooming in and zooming out. 
-\end{comment}
-
-\stitle{\noah: \new{an navigation plug-in for spreadsheets}.} We address the aforementioned challenges
+\stitle{\noah: \new{a navigation plug-in for spreadsheets}.} We address the aforementioned challenges
 in \noah,
@@
 the data by a specific attribute,
-\saj{a multi-granularity overview is constructed and displayed
+a multi-granularity overview is constructed and displayed
 within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a \new{fine or coarse-}grained 
@@
 Each bin (group) of the histogram is 
-mapped to a collection of rows in the spreadsheet.} 
+mapped to a collection of rows in the spreadsheet. 
 Cumbersome scrolling operations are eliminated in favor
@@
 the benefits and limitations of this plugin.
-\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
+\newsaj{The study required users to
+perform tasks that were representative of
+popular spreadsheet operations,
+using both Microsoft Excel and \noah.}
+The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
 tasks involving navigation 
 correctly and quickly in \noah.
-Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} \agp{mention how these tasks were representative or general}
+Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.
+\agp{mention how these tasks were representative or general}
 \squishend

### 2019-10-11 introduction.tex
teal color added back

--- 2019-10-10 introduction.tex
@@
 to access the rows mapped to that group. 
+\newsaj{As the granularity of the dynamic overview 
+changes, how do we efficiently update the mapping 
+from spreadsheet rows to the 
+finer or coarser groups 
+so that scrolling remains seamless? 
 Similarly, how do we leverage
-the overview to support simple interactions
-to steer spreadsheet data (i.e., select a range of data)
-for formula computation?
-As the granularity of the dynamic overview 
-changes, how do we efficiently update the mapping 
-from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless? 
-How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch? 
-How do we ensure such automatic recomputation 
-scales with the size of the data?\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}}
+the overview to steer spreadsheet data (i.e., select a range of data)
+for formula computation? How do we present the results of
+the formula within the overview? One approach can be to adopt
+the pivot table-like presentation of results. Within their summary view, 
+pivot tables display 
+aggregate formula results (\eg \code{SUM}, \code{COUNT}) 
+alongside each group. 
+However, unlike pivot tables, users can explore the overview at
+multiple-granularities. As the granularity changes,
+the grouping of rows also changes; 
+making the previous formula results 
+inconsistent with the new groups.
+How do we 
+recompute the results of a formula 
+as the granularity changes 
+without requiring the users to 
+reissue the formula from scratch?}}
+\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}
 
 \item {\em Seamless integration as a plug-in.}
-Finally, the design of the overview
-should be generic so that it can be integrated
+\newsaj{Finally, how can we 
+design an overview with a generic set of features,
+that can be integrated
 with any existing spreadsheet tool on any 
 dataset without impacting
-existing functionalities or look-and-feel.
-Any new interactions introduced should be consistent
-with traditional spreadsheet semantics,
-and complement existing spreadsheet interactions.
-Interactions between the overview and the raw spreadsheet
-should be coordinated to the extent possible.
+existing functionalities or look-and-feel?
+How do we ensure that the interactions 
+supported by the overview 
+are 
+consistent with
+traditional spreadsheet semantics,
+and complement existing spreadsheet interactions?
+How to we enable coordinated interactions across both views, \ie 
+the overview and the raw spreadsheet, such that they
+remain consistent at any given time?}
 \squishend
@@
 the benefits and limitations of this plugin.
+\newsaj{The study required users to
+perform tasks that were representative of
+popular spreadsheet operations,
+using both Microsoft Excel and \noah.}
 \saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet

### 2019-10-11 letter.tex
letter reorganized

@@
 \begin{itemize}
-    \item More clearly articulate the merits over alternative tools, such as Keshif. 
-     \item Convincingly explain the design choice around the binning mechanism or revise
+    \item[\textbf{C1:}] More clearly articulate the merits over alternative tools, such as Keshif. 
+     \item[\textbf{C2:}] Convincingly explain the design choice around the binning mechanism or revise
     that approach.
-    \item Justify the choice of comparing with Excel and note the limitations of that
+    \item[\textbf{C3:}] Justify the choice of comparing with Excel and note the limitations of that
     approach.
-     \item Better justify the choice of tasks in the study. 
-     \item Explain the choice of the four research questions.
-     \item Provide more details about the study, including reporting intra-participant
+     \item[\textbf{C4:}] Better justify the choice of tasks in the study. 
+     \item[\textbf{C5:}] Explain the choice of the four research questions.
+     \item[\textbf{C6:}] Provide more details about the study, including reporting intra-participant
     differences.
@@
 integrated within an existing spreadsheet, as opposed
-to other general-purpose exploratory tools such as  Keshif;
-(ii) we justify our design choice surrounding the binning mechanism;
-and 
+to other general-purpose exploratory tools such as  Keshif (\textbf{C1});
+(ii) we justify our design choice surrounding the binning mechanism
+while discussing its limitations (\textbf{C2}); 
 (iii) we enrich the user study design section further 
-by providing justifications for comparing \noah and Excel, 
-rephrasing our research questions reflecting the purpose of the user study, explaining the choice of quiz tasks,
-(iv) we reorganize the results section to  highlight key contributions of \noah while explaining participant experiences in using the tool.
-Moreover, we report the  intra-participant differences for the study.
+by providing justifications for comparing \noah and Excel (\textbf{C3}), 
+explaining the choice of quiz tasks (\textbf{C4}), rephrasing our research questions reflecting the purpose of the user study (\textbf{C5}), and reporting the intra-participant differences for the study (\textbf{C6}).
+Moreover, to answer the rephrased research questions, we 
+have reorganized the results section to highlight key contributions of \noah while explaining participant experiences in using the tool.
 The major changes to the paper include:
@@
 
-\item \textit{Explanation of the binning mechanism.} We have explained our choice of the binning mechanism in the context the third design consideration (\emph{DC3}) and can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}.
-\saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge the limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss}.
-
-\item \textit{Justification of Excel as a Baseline.}
-We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets in interacting with data even for the users of enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and the TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to facilitate direct manipulation and arbitrary derivation of the data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which, far exceeds that of the TDA tools (see Section~\ref{sec:study_design}).
-
-\item \textit{Merits over other tools.} \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation interface}, the concepts of which can be implemented and integrated as a plug-in to any spreadsheet.
+\item \textit{Merits over other tools} (\textbf{C1}): \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation interface}, the concepts of which can be implemented and integrated as a plug-in to any spreadsheet.
 As spreadsheets have a massive user-base, enhancing the experience surrounding data exploration and formula computation while maintaining the spreadsheet look and feel as much as possible, has the potential to impact a lot of users.
@@
 
-\item \textit{Enhancing the user study design.}
+\item \textit{Explanation of the binning mechanism} (\textbf{C2}): We have explained our choice of the binning mechanism in the context the third design consideration (\emph{DC3}) and can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}.
+\saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge the limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss}.
+
+\item \textit{Justification of Excel as a Baseline} (\textbf{C2}):
+We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets in interacting with data even for the users of enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and the TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to facilitate direct manipulation and arbitrary derivation of the data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which, far exceeds that of the TDA tools (see Section~\ref{sec:study_design}).
+
+\item \textit{Enhancing the user study design} (\textbf{C4, C5}):
 We have added additional explanations on the goal of the study while better articulating the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
@@
 
-\item \textit{Reporting Intra-participant Differences.}
+\item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
 Finally, we have performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data.

### 2019-10-11 usecases.tex
partial pass on usecases

@@
 \label{sec:usage}
-\saj{As explained in Section~\ref{sec:intro}, users prefer spreadsheets over enterprise solutions to view, directly manipulate, and explore data~\cite{chan1996use,raden2005shedding}.
-To identify the scope of these tasks on spreadsheets (see Table~\ref{tab:scope}), we draw parallel to the typology of abstract data exploration tasks~\cite{brehmer2013multi}.
-The typology characterizes the exploration and manipulation tasks performed on visual representations of data.
-While all the tasks in Table~\ref{tab:scope} can be performed using spreadsheets, \noah enhances the experience for a subset of these tasks, indicated by a checkmark (\checkmark).
-We explain how \noah complements spreadsheets in accomplishing these tasks in Appendix.}
+\new{Users prefer spreadsheets 
+over enterprise solutions 
+to view, explore, and analyze data~\cite{chan1996use,raden2005shedding}. 
+To understand the scope of typical user 
+tasks on spreadsheets, 
+we make use of the typology 
+of abstract data exploration tasks~\cite{brehmer2013multi}---see 
+Table~\ref{tab:scope}. 
+This typology characterizes the range of domain-independent
+tasks performed on visual representations of data,
+developed after analyzing task classification 
+systems in over two dozen papers, 
+and has been applied to a variety of 
+scenarios, including XXX\agp{argue for
+ utility with citations.}
+While all the tasks in Table~\ref{tab:scope} 
+can be performed using spreadsheets, 
+\noah enhances the experience 
+for many of these tasks, 
+indicated by a checkmark (\checkmark). 
+We describe these tasks in the context
+of a real usage scenario for \noah below.
+Additional details regarding how \noah 
+complements spreadsheets 
+in accomplishing these tasks can be found in 
+the Appendix.}
+
+% \begin{table}[!htb]\scriptsize
+% \vspace{-10pt}
+% \caption{Use cases of \noah, employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
+% \label{tab:scope}
+% \vspace{-10pt}
+% \centering
+% \begin{tabular}{c p{2.8in}}
+% \hline
+% Purpose & Use Cases   \\ \hline
+% \emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
+% \emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
+% \emph{Search}            & \code{browse} (\checkmark), \code{explore(\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
+% \emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
+% \end{tabular}
+% \vspace{-5pt}
+% \end{table}
 
 \begin{table}[!htb]\scriptsize
-\caption{Uses cases of \noah.}
+\new{
+\vspace{-10pt}
+\caption{Use cases of \noah, employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
 \label{tab:scope}
+\vspace{-10pt}
 \centering
-\begin{tabular}{c c}
+\begin{tabular}{c p{2.8in}}
 \hline
 Purpose & Use Cases   \\ \hline
-\emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
-\emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
-\emph{Search}            & \code{browse} (\checkmark), \code{explore (\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
-\emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
+\emph{Consume}         & 
+\code{discover} (\checkmark: {\em generation of hypotheses}, e.g., Rebecca finds a trend in larger cities and wants to check if it is present in smaller cities), 
+\code{present} (\checkmark: {\em communication of information}, e.g., Rebecca takes a screenshot of \noah to share with her coworkers of the overall availability trends in the context of raw listings),
+\code{enjoy} (\checkmark: {\em casual encounters with visualization}, e.g., Rebecca uses the overview ``at a glance'' 
+to understand which cities are present in the dataset, 
+and how many listings are present per city)\\
+\emph{Search}            & 
+\code{explore} (\checkmark: {\em searching based on characteristics where location is unknown}, e.g.,), 
+\code{browse} (\checkmark: {\em searching based on characteristics where location is known}, e.g.,), 
+\code{locate} (\checkmark: {\em searching based on entities where location is unknown}, e.g., Rebecca wants to find all entries corresponding to a given city like Chicago), \code{lookup} (\checkmark: {\em searching based on entities where location is known}, e.g.,) \\
+\emph{Query}    		&  
+\code{identify} (\checkmark: {\em returning the characteristics of entity found during search}, e.g., Rebecca wants to examine Chicago listings and the typical availabilities of listings in Chicago), 
+\code{compare} (\checkmark: {\em returning characteristics of multiple entities}, e.g., Rebecca wants to compare listing patterns in Boston to that of Chicago), \code{summarize} (\checkmark: {\em returning characteristics of several entities}, e.g., Rebecca wants to gain an understanding of overall rental patterns across cities)  \\
+\emph{Produce}         & \code{export/save/generate} ($\times$), \code{record} (\checkmark:)\\
+\hline
 \end{tabular}
+}
+\vspace{-5pt}
 \end{table}
 
-\saj{We now describe a usage scenario that captures the spreadsheet exploration tasks accommodated by \noah (Table~\ref{tab:scope}) while illustrating the benefits of integrating an overview with spreadsheets.
-Let's assume that} Rebecca, an investigative journalist
-is exploring the Inside Airbnb dataset~\cite{web:airbnb},
-a dataset of all the Airbnb listings across different US cities.
+\new{We now describe a usage scenario 
+that illustrates the benefits of integrating \noah
+into typical spreadsheets. 
+Let's assume that Rebecca, a journalist,
+is exploring the {\em Inside Airbnb} dataset~\cite{web:airbnb},
+a dataset of all the Airbnb listings across different US cities.}
 This dataset was created to investigate the long-standing
@@
 
-Without any prior knowledge of the dataset,
-Rebecca starts by finding out information
-about cities in the dataset and their corresponding listings;
-for example, which cities are present,
-and roughly how many listings does each city have \saj{(\code{browse} and \code{explore})}.
-Without \noah, Rebecca may have used
+\new{Given that this is the first time she's examining
+this dataset,
+Rebecca wants to first gain a bird's eye view of the data. 
+Without \noah, Rebecca 
+would have had to use} 
 a pivot table (discussed in Section~\ref{sec:related})
@@
 about listings from any given city.
-If she wanted to identify listings from a specific city \saj{(\code{lookup} and \code{locate} or \code{identify})},
+If she wanted to examine listings from a 
+specific city,
 Rebecca would have to either use search capabilities or
@@
 on the spreadsheet.
-Even at the first step of this exploration,
-Rebecca would experience substantial cognitive burden,
+\new{Even at the first step of exploration,
+Rebecca would experience {\em substantial cognitive burdens,
 loss of context,
-and visual discontinuities,
+and visual discontinuities},
 with subsequent steps becoming progressively
-more challenging.
+more challenging.}
 
-Using \noah, she explores
-the data by city,
-with \noah providing a high-level overview of
-cities (Figure~\ref{fig:ux}a).
+\new{Using \noah, 
+she organizes the overview by city
+and starts casually exploring the dataset, understanding
+which cities are present, and roughly how many listings
+does each city have---with \noah providing a high-level overview of
+cities (Figure~\ref{fig:ux}a) (\code{enjoy}).}
 The overview consists of sorted non-overlapping bins
@@
 data will be displayed at the top of her screen.
-For example, clicking on the {\em Ashville-Boston} bin
-displays the Ashville listings (Figure~\ref{fig:ux}c).
+\new{For example, clicking on the {\em Ashville-Boston} bin
+displays the Ashville listings (Figure~\ref{fig:ux}c);
+she can similarly find and examine properties of 
+the Chicago listings by clicking on the {\em Chicago-Denver} bin (\code{locate} followed by \code{identify}).}
 She can also zoom into bins using the ``\textcolor{blue}{$\rangle$}'' arrows,
@@
 
-Next, say Rebecca wants to analyze
-one of the larger cities \saj{(\code{query}/\code{summarize}) to understand the renting pattern}.
-She identifies these cities by
-comparing the number of listings
-for each city displayed on the overview \saj{(\code{compare})}.
+\new{Next, say Rebecca wants to analyze
+one of the larger cities to understand the 
+overall renting pattern (\code{summarize}).
+She studies a few cities at a time, examining and comparing
+the number of listings for each city, as displayed on the overview (\code{compare})}.
 She decides to focus on Boston, her hometown,

### 2019-10-11 usecases.tex
finished pass on usecases

@@
 \vspace{-10pt}
-\caption{Use cases of \noah, employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
+\caption{Example use cases of \noah, employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
 \label{tab:scope}
@@
 \code{discover} (\checkmark: {\em generation of hypotheses}, e.g., Rebecca finds a trend in larger cities and wants to check if it is present in smaller cities), 
-\code{present} (\checkmark: {\em communication of information}, e.g., Rebecca takes a screenshot of \noah to share with her coworkers of the overall availability trends in the context of raw listings),
+\code{present} (\checkmark: {\em communication of information}, e.g., Rebecca sees the overall availability trends in the context of raw listings, and can present this view to her co-workers), 
 \code{enjoy} (\checkmark: {\em casual encounters with visualization}, e.g., Rebecca uses the overview ``at a glance'' 
@@
 \emph{Search}            & 
-\code{explore} (\checkmark: {\em searching based on characteristics where location is unknown}, e.g.,), 
-\code{browse} (\checkmark: {\em searching based on characteristics where location is known}, e.g.,), 
-\code{locate} (\checkmark: {\em searching based on entities where location is unknown}, e.g., Rebecca wants to find all entries corresponding to a given city like Chicago), \code{lookup} (\checkmark: {\em searching based on entities where location is known}, e.g.,) \\
+\code{explore/browse} (\checkmark: {\em searching based on characteristics where location is unknown/known}, e.g., Rebecca tries to find Chicago listings with availability greater than 60 days), 
+\code{locate/lookup} (\checkmark: {\em searching based on entities where location is unknown/known}, e.g., Rebecca wants to find all entries corresponding to a given city like Chicago) \\
 \emph{Query}    		&  
-\code{identify} (\checkmark: {\em returning the characteristics of entity found during search}, e.g., Rebecca wants to examine Chicago listings and the typical availabilities of listings in Chicago), 
+\code{identify} (\checkmark: {\em returning the characteristics of entity found during search}, e.g., Rebecca wants to examine Chicago listings to assess typical availabilities of listings in Chicago),
 \code{compare} (\checkmark: {\em returning characteristics of multiple entities}, e.g., Rebecca wants to compare listing patterns in Boston to that of Chicago), \code{summarize} (\checkmark: {\em returning characteristics of several entities}, e.g., Rebecca wants to gain an understanding of overall rental patterns across cities)  \\
-\emph{Produce}         & \code{export/save/generate} ($\times$), \code{record} (\checkmark:)\\
+\emph{Produce}         & \code{export/save} ($\times$), \code{generate/record} (\checkmark: {\em generation or recording of new information}, e.g., Rebecca issues an aggregate formula to generate summary availability statistics across cities)\\
 \hline
@@
 in Boston violate
-the ``rent availability $>60$ days'' condition.
+the ``rent availability $>60$ days'' condition \new{(\code{identify})}.
 In a typical spreadsheet, Rebecca needs to manually steer and then select
-the Boston listings as input to a \code{COUNTIF} formula that counts the number of
+the Boston listings as input to a \code{COUNTIF} 
+formula that counts the number of
 rows that satisfy the above mentioned condition.
@@
 and then issues
-a \code{COUNTIF} operation on the overview.
+a \code{COUNTIF} operation on the overview  \new{(\code{generate})}.
 %\ie a \code{COUNTIF} formula\footnote{\code{COUNTIF} is a
@@
 alongside the overview (Figure~\ref{fig:ux}b).
-Rebecca learns \saj{(\code{discover})} that more than half of the listings in Boston are 
-effectively operating as hotels---a large number!
+Rebecca learns that 
+more than half of the listings in Boston are 
+effectively operating as hotels \new{(\code{discover})}---a large number!
 
@@
 understand availability statistics for an even
-larger city, Chicago \saj{(\code{summarize})}.
+larger city, Chicago \new{(\code{compare})}.
 As she uses the overview to navigate
@@
 automatically updates the aggregate column to
-the \code{COUNTIF} formula results for Chicago,
-without Rebecca needing to reissue it by performing a cumbersome steering operation as in traditional spreadsheets. 
+the \code{COUNTIF} formula results for Chicago \new{(\code{identify})},
+without Rebecca needing to reissue it by performing a 
+cumbersome steering operation as in traditional spreadsheets. 
 Rebecca learns that Chicago exhibits 
-a similar renting pattern as Boston, with more than half the listings operating as hotels. She can then check if this trend holds for all large cities \saj{(\code{compare})},
-and whether the smaller cities have a different pattern \saj{(\code{discover})}. \saj{Note that, the rows that satisfy the ``rent availability $>60$ days'' condition, are listed
+a similar renting pattern as Boston, with more than half the listings operating as hotels. 
+\new{She can then hypothesizes that this trend may hold for all large cities,
+and can check whether the smaller cities have a different pattern (\code{discover}). Note that, the rows that satisfy the 
+``rent availability $>60$ days'' condition, are listed
  in the spreadsheet adjacent
-to the overview in sky blue (Figure~\ref{fig:ux}g)}. With the raw 
+to the overview in sky blue (Figure~\ref{fig:ux}g) (\code{explore})}. 
+With the raw 
 data presented side-by-side, she can also dive into other attributes
 of the listings operating as hotels to see if there are any other identifying characteristics,
-\eg if they are all managed by a small number of agencies acting as individual renters \saj{(\code{identify})}.
+\eg if they are all managed by a small number of agencies acting as individual renters \new{(\code{identify})}.
 
@@
  (Figure~\ref{fig:ux}e) are kept up-to-date,
- allowing her to maintain context during navigation.
- She can revisit any previously visited cities
+ allowing her to maintain context during navigation \new{(\code{record})}.
+ She can revisit any previously visited cities \new{(\code{lookup})}
  by simply clicking on the relevant path

### 2019-10-11 introduction.tex
partial intro

@@
 {\bf \em spreadsheets are by far the most popular medium for ad-hoc exploration
-and analysis of data}~\cite{excel-users}.} 
-\saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
+and analysis of data}~\cite{excel-users}.
+Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
 with more advanced analytical features~\cite{chan1996use,raden2005shedding}. 
@@
 via two unit operations, scrolling and steering.
-\saj{Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.}
+\new{Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.}
 For example, when analyzing data, users may scroll
@@
 a high-level view of the overall spreadsheet.
-\saj{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
+\new{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
 users may steer to select
@@
 now supporting increasingly larger datasets, \eg Google Sheets
-now supports five million cells~\cite{gs}, a $12.5X$ increase
+now supports five million cells~\cite{gs}, a $12.5\times$ increase
 from the previous limit of 400K cells, 

### 2019-10-11 introduction.tex
intro

@@
 \item {\em Overview modality.}
-\saj{
-One approach for an overview is to simply add a
-zoomed out version of the entire sheet as a pane on the side,
+\new{
+One could simply add a
+zoomed out version of the entire sheet as a pane on the side
+to create a spreadsheet overview,
 as in popular presentation software like 
-Microsoft PowerPoint or text editors
+Microsoft PowerPoint, or text editors
 like Sublime Text.
-\newsaj{The zoomed out overview
-displays the entire data 
-at a lower magnification scale.} 
+The zoomed out overview
+would 
+display the data 
+at a lower magnification. 
 Unfortunately, this approach
 would not suffice for a spreadsheet.
-\new{An overview should provide a
-comprehensible big picture view:
-for a \newsaj{spreadsheet of numbers, text, or formulae}, 
+An overview should provide a
+comprehensible big picture view;
+for a spreadsheet of numbers, text, or formulae, 
 when zooming out beyond a point, 
-\newsaj{an overview displayed at 
-such low magnification scale} 
-would be unreadable.}
+an overview displayed at 
+such low magnification 
+would be unreadable.
 Another approach, adopted by map tools like the
@@
 to provide a global context of the user's current location
-being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
+currently displayed in the zoomed-in detailed view~\cite{cockburn2009review}.
 While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
-view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
+view which allows objects to be represented differently at different scales. 
+Since spreadsheets already display the raw data,
 zooming into and out of a detailed view consisting of
 this raw data is not meaningful. 
-\new{How do we design an overview to dynamically 
+How do we design an overview to dynamically 
 change as users 
 seek a more fine-grained or coarse-grained view 
-of the overall structure of the data?} 
+of the overall structure of the data?
 }
@@
 \item {\em Construction of the overview.}
-\saj{Given a spreadsheet
+\new{Given a spreadsheet
 with many rows, 
@@
 \item {\em Operations on the overview.}
-\saj{Following the construction of a dynamic overview, 
+\new{Following the construction of a dynamic overview, 
 the next challenge is to design simple interactions 
@@
 to access the rows mapped to that group. 
-\newsaj{As the granularity of the dynamic overview 
+As the granularity of the dynamic overview 
 changes, how do we efficiently update the mapping 
@@
 the formula within the overview? One approach can be to adopt
-the pivot table-like presentation of results. Within their summary view, 
+the pivot table-like presentation of results. 
+Within their summary view, 
 pivot tables display 
@@
 However, unlike pivot tables, users can explore the overview at
-multiple-granularities. As the granularity changes,
+multiple granularities. As the granularity changes,
 the grouping of rows also changes; 
@@
 recompute the results of a formula 
+in a convenient manner
 as the granularity changes 
 without requiring the users to 
-reissue the formula from scratch?}}
-\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}
+reissue the formula from scratch?}
 
 \item {\em Seamless integration as a plug-in.}
-\newsaj{Finally, how can we 
+\new{Finally, how can we 
 design an overview with a generic set of features,
 that can be integrated
-with any existing spreadsheet tool on any 
-dataset without impacting
+with any existing spreadsheet tool,
+operating on any 
+dataset, without impacting
 existing functionalities or look-and-feel?
@@
 \squishend
-\agp{A general comment: can we ensure that either we use
-questions for all of the items above or none of them for uniformity?}
 
@@
 the data by a specific attribute,
-\saj{a multi-granularity overview is constructed and displayed
-within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a \new{fine or coarse-}grained 
+\new{a multi-granularity overview is constructed and displayed
+within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a fine or coarse-grained 
 perspective of the data distribution. 
@@
 We realize this design in the form of \noah,
-a plugin to an existing web-based spreadsheet tool,
+a plugin to a spreadsheet tool,
 ensuring that interactions supported by \noah
@@
 the benefits and limitations of this plugin.
-\newsaj{The study required users to
+\new{The study required users to
 perform tasks that were representative of
 popular spreadsheet operations,
-using both Microsoft Excel and \noah.}
-\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
-tasks involving navigation 
-correctly and quickly in \noah.
-Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} \agp{mention how these tasks were representative or general}
+using both Microsoft Excel and \noah.
+The study shows that compared to Microsoft Excel, 
+participants were able to complete spreadsheet
+navigation tasks 
+correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
 \squishend

### 2019-10-11 related.tex
related work commit

@@
+%!TEX root = main.tex
+
 \section{Related Work}
 \label{sec:related}
-In this section, we present tools and techniques
+We now discuss tools and techniques
 that partially address the limitations of navigating
-spreadsheets as discussed in the introduction.
+spreadsheets.
 
@@
 targeted at improving users' navigation experience.}
+\agp{Abrupt, add at least one line to say that existing spreadsheet tools recognize challenges in nav or something}
 
@@
 \stitle{Microsoft Excel.}
-\saj{Excel enables users to manually create references
-to a spreadsheet region using the named ranges~\cite{named-range} feature which is displayed within the Excel menu bar.
+\new{Excel enables users to manually create references
+to a spreadsheet region using the named 
+ranges~\cite{named-range} feature,
+accessible from the menu bar.
 Users can click on a named range
-to navigate to the referred region. However, the onus is on the user to create
-named ranges for each region of interest. The Pivot table~\cite{pivot} feature provides a summary view, enabling users
+to navigate to the referred region. 
+However, the onus is on the user to create
+named ranges for each region of interest. 
+The pivot table~\cite{pivot} feature 
+allows users to create a summary view to 
+compare subsets of data without having to 
+provides a summary view, enabling users
 to compare subsets of data without
 having to navigate to various locations
-within the spreadsheet.
+within the sheet.
 This summary is placed
-in a separate region in the spreadsheet,
+in a separate region of the spreadsheet,
 preventing users from accessing
-the data underlying the summary, thereby, impeding navigation. A similar overview feature, \code{SUBTOTAL}~\cite{subtotal},
-adds a new row at the end of the
-corresponding subset of data with summary
+the data underlying the summary, impeding navigation. 
+A similar overview feature, \code{SUBTOTAL}~\cite{subtotal},
+adds a new row at the end of each distinct
+subset of data with summary
 information. Users can expand the summary to view
 the actual spreadsheet data.
-However, for datasets with larger subsets (\eg for numeric data),
-the size of the summary can become very
-large spanning multiple screens and can cause
+However, for datasets with many subsets (\eg for numeric data),
+the number of new lines inserted (i.e., the summary) 
+can itself become very
+large, spanning multiple screens, and can cause
 increased
 visual discontinuity during navigation. Finally, NodeXL~\cite{hansen2010analyzing}
-is an Excel plug-in that provides an overview
-of a network
-and enables users to perform navigational operations,
+is a plug-in that provides a spreadsheet network overview
+and supports navigational operations,
 \eg zooming in/out, dynamic filtering, on the overview;
-this plugin only supports network datasets,
+this plug-in only supports network datasets,
 such as biological or social networks.}
@@
 Hillview~\cite{web:hillview} displays
-the results of group-by queries
+the approximate results of group-by queries
 on large spreadsheet tables.
@@
 aggregate statistics). 
-\saj{However, these summaries 
+\new{However, these summaries 
 are static overviews of the data---much like pivot tables, 
@@
 to nor are co-located with
-the underlying data. 
-Naturally, 
-users cannot navigate and directly manipulate the raw data
-using these tools unlike spreadsheets.
+the underlying raw data. 
+As such, the spreadsheet look-and-feel is lost,
+and as a result, users
+cannot navigate and directly manipulate the raw data
+in the way they do in spreadsheets. 
 For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}).
+\agp{previous line problematic: they can...
+because all these tools support to varying extents the typology}
 Even users of such tools want to inspect the underlying data to verify the summary statistics and visualizations~\cite{2017trust}.
+\agp{this line is problematic too...
+because the citation is to an AQP tool rather than some general purpose study}
 %several wanted to be able to see the underlying data: although big data demands aggregations, analysts wanted to see individual records to spot-check their results, and to get a sense of what sat in a bucket.
-The goals of spreadsheets vary from the TDA tools in two different ways: a) facilitating direct manipulation of the data and b) enabling arbitrary derivation of new data and summaries
+The goals of spreadsheets differ from TDA 
+tools in two ways: a) facilitating direct manipulation of raw 
+data and b) enabling arbitrary derivation of new data and summaries
 using various operations involving navigation,
 \eg formulas or expressions.
-\noah being a plugin to spreadsheets, facilitates a unified interface that
+\noah being a plug-in to spreadsheets, 
+facilitates a unified interface that
 upholds both these goals
 while enhancing
-the navigational capabilities
+navigational capabilities
 for spreadsheet users.

### 2019-10-30 design.tex
aditya comments addressed

@@
 However, maintaining the overview in a separate
-location from the data can be cumbersome;
+location from the data can \newsaj{lead to loss of context};
 instead, having it co-located with the data
@@
 Users often need to access subsets of data,
-and study their properties in detail (\eg via steering);
-navigating back and forth \saj{between different subsets of data}
+and study their properties in detail, \eg via steering.
+Navigating back and forth \saj{between different subsets of data}
 can lead to increased visual discontinuity.
@@
 \ie an interaction on one should be reflected on
-the other interface~\cite{wang2000guidelines}.
+the other~\cite{wang2000guidelines}.
 For example, as a user scrolls through
@@
 
-\para{DC6. Display contextual and historical information regarding users' navigation}
+\para{DC6. Display contextual and historical navigation information}
 The interface should record navigation history,
@@
 on the navigation attribute ``city''
-has two levels of granularities.
-The highest level of the granularity
+has granularity levels.
+The highest (coarsest) granularity level
 consists of four bins.
 Figure~\ref{fig:concept}a
-depicts \noah showing the first four bins,
+depicts the first four bins,
 the first of which is {\em Ashville-Boston}.
@@
 it spans, \eg starting row and
-ending row number,
+ending row number, and 
 the total number of rows the region spans.
@@
 one for each bin in the next granularity.
-On the other hand,
-since the third bin from the top
+Since the third bin from the top
 (\emph{LA}) spans only one city,
@@
 Multi-granularity interfaces
-provide an alternative to a single-granularity
-spatial partition of the data space,
-by varying the scale at which the data is displayed,
-enabling multiple views at varying granularities~\cite{cockburn2009review}.
+provide an alternative to \newsaj{the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
 Users can resize the overview to control the amount of spreadsheet data that remains visible.
@@
 space between overview and detail.}
-Moreover, users can also hide the overview if required.}
+Users can also hide the overview if required.}
 
-\saj{As explained earlier,
-we divide each granularity
-of the overview into bins.
-The highest (coarsest) level of the granularity
-provides a global view with less visual clarity
-and reduced visual discontinuity,
-As the bins become more fine-grained, visual clarity improves while visual discontinuity increases.
+\newsaj{The data structure underlying the overview
+is an equi-depth histogram constructed
+on the values
+in the navigation attribute column.
+Equi-depth histograms are commonly
+used for summarizing the statistical properties
+of data, with applications such as
+database query optimizers~\cite{chaudhuri1998random}.
 The binned representation of the overview
@@
 by reducing visual discontinuity. 
-The bins of the histogram
-act as landmarks
+The bins act as landmarks
 in the overview, enabling 
 users to skip irrelevant bins
-and quickly navigate to the desired subset of data.
-The equi-depth histogram 
+and quickly navigate to the desired subset of data. We now discuss how the overview is constructed.}
+
+\stitle{Overview Construction.}
+\newsaj{The equi-depth histogram 
 can be constructed on 
@@
 in the Airbnb dataset.}
-
-\stitle{Overview Construction.}
-The data structure underlying the overview
-is an equi-depth histogram constructed
-on the values
-in the navigation attribute column.
-Equi-depth histograms are commonly
-used for summarizing the statistical properties
-of data, with applications such as
-database query optimizers~\cite{chaudhuri1998random}.
 Each bin in the equi-depth histogram
@@
 that the ranges represented by each bin have no overlap.
-\toappendix{\agp{We don't have time but for the future, this is
-something that would have been great to describe the algorithm on
-for technical depth.}}
+\agp{Aditya: would be valuable to describe the algorithm to increase
+technical depth.}
+\saj{Saj: can we add a pseudocode in appendix and refer to that?}
 
@@
 corresponding spreadsheet data (\textbf{DC4}),
-\ie interaction on the overview
+\ie interactions on the overview
 may be reflected on the spreadsheet and vice-versa.
-Indicating the navigation attribute on the spreadsheet
+\newsaj{One example of this coordinations is,} indicating the navigation attribute on the spreadsheet
 using color (see the lime green column in Figure~\ref{fig:ux}c)
-as user constructs the overview, is an example of such coordination. However, not all overview interactions are coupled
+as user constructs the overview. However, not all overview interactions are coupled
 with the spreadsheet and vice versa.
@@
 the spreadsheet view does not change.
-For example, in Figure~\ref{fig:concept}a
+For example, in Figure~\ref{fig:concept}a,
 the user zooms into the \emph{Ashville-Boston} bin;
@@
 there's no need to update the spreadsheet view.
-Similarly, operations like panning on the overview without clicking
+Similarly, operations like panning on the overview without clicking,
 and customizing the overview
-do not change user's current focus and are decoupled.
+do not change user's current focus and are therefore decoupled.
 Online maps also adopt similar
@@
 However, adding an aggregate column
-takes up screen space thus shrinking the spreadsheet view.
+takes up screen space, shrinking the spreadsheet view.
 As a workaround, users can resize or
-remove aggregate columns
-to make more screen space available
-to the spreadsheet (\textbf{DC3}).
-When user issues a formula on the overview,
+remove aggregate columns if required (\textbf{DC3}).
+When the user issues a formula on the overview,
 the spreadsheet column
@@
 cells that satisfy the condition
-are highlighted, \eg in Figure~\ref{fig:ux}c),
+are highlighted, \eg in Figure~\ref{fig:ux}c,
 the cells with availability $\ge 60$ are colored in sky blue.
-By viewing the colored cells,
-users can find out which cells
+In this manner,
+users can quickly determine which cells
 are relevant to the aggregation operation\toappendix{ (\code{identify} in Table~\ref{tab:scope})}.
@@
 rather they simply select the formula from a drop-down menu,
-and provide the necessary formula parameters to a web form.
-This way, \noah maintains the semantics
-of existing spreadsheet tools.
-Currently, \noah supports statistical
+and provide the necessary formula parameters to a form.
+\newsaj{The aggregate column can employ any} statistical
 and mathematical formulae that
 operate over a range of data.
-Therefore, creating an aggregate column
-is equivalent to selecting a subset of data
-on the sheet, \ie steering and then executing a formula,
+\newsaj{Therefore, creating an aggregate column
+is equivalent to selecting subsets of data
+on the sheet, \ie steering, and then executing a formula on this subset},
 helping users avoid cumbersome steering operations.
@@
 
-Users can choose to view the result of formulae
-either as raw values or charts and can toggle between the two.
+Users can view the results
+either as raw values or as charts, and can toggle between the two.
 Raw values are displayed
@@
 The chart representation varies depending on the formula type.
-Except for the \emph{others} category
-all other categories can be represented by charts.
+All other categories except for the \emph{others} category
+can be represented by charts.
 Figure~\ref{fig:agg} shows the
@@
 as well as other information. We discuss these representations 
-in detail in the supplementary material.
+in detail in the Appendix.
 \toappendix{For example, for \emph{conditional} formulae

### 2019-10-30 related.tex
aditya comments addressed

@@
 \agp{Abrupt, add at least one line to say that existing spreadsheet tools recognize challenges in nav or something}
+\newsaj{Existing spreadsheet tools---both commercial software and academic prototypes---recognize some of these navigation challenges while adopting different measures to address those.}
 
@@
 aggregate statistics). 
-\new{However, these summaries 
+\new{These summaries 
 are static overviews of the data---much like pivot tables, 
@@
 the underlying raw data.
-As such, the spreadsheet look-and-feel is lost,
+\newsaj{For example, Keshif~\cite{yalccin2018keshif} can display all the unique values corresponding to an attribute of interest, \eg cities of the Airbnb data~\cite{web:airbnb}.
+However, users cannot view or inspect the raw data corresponding to each city.
+With TDA tools, the spreadsheet look-and-feel is lost,
 and as a result, users
-cannot navigate and directly manipulate the raw data
-in the way they do in spreadsheets. 
-For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}).
-\agp{previous line problematic: they can...
-because all these tools support to varying extents the typology}
+loose the capability to directly manipulate raw data, derive new data, and issue formulae for free-form analysis.}  
+\cut{For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}).
+Aditya: previous line problematic: they can...
+because all these tools support to varying extents the typology.
 Even users of such tools want to inspect the underlying data to verify the summary statistics and visualizations~\cite{2017trust}.
-\agp{this line is problematic too...
+Aditya:this line is problematic too...
 because the citation is to an AQP tool rather than some general purpose study}
-%several wanted to be able to see the underlying data: although big data demands aggregations, analysts wanted to see individual records to spot-check their results, and to get a sense of what sat in a bucket.
-The goals of spreadsheets differ from TDA 
+\newsaj{Therefore, the goals of spreadsheets differ from TDA 
 tools in two ways: a) facilitating direct manipulation of raw 
@@
 using various operations involving navigation,
-\eg formulas or expressions.
+\eg formulas or expressions.}
 \noah being a plug-in to spreadsheets, 

### 2019-10-30 study.tex
aditya comments addressed

@@
 \label{sec:study_design}
-\saj{\noah was designed to address the navigational challenges in spreadsheets.
-As such, the goal of the study was to evaluate whether \noah enhances the navigational capabilities of traditional spreadsheets.
-Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah affects a spreadsheet user's navigation experience.
-In this study, we compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain.
-Similar domain specific evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or
+\newsaj{The goal of our study is to evaluate whether \noah addresses the navigational challenges in traditional spreadsheets.
+Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah impacts a spreadsheet user's navigation experience as they explore and make sense of data.
+Therefore, we decided to compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain.} \saj{Similar domain specific-evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or
 tree navigation~\cite{beard1990navigational}.
-As explained in Section~\ref{sec:related}, the goals of spreadsheets and the TDA tools are quite different.
+As explained in Section~\ref{sec:related}, the goals \newsaj{and user populations} of spreadsheets and TDA tools are quite different.
 Therefore, we did not consider TDA tools for the comparative study.}
 Our study was designed to answer the following questions:
-
+\agp{The RQs feel repetitive. What did Karrie think.}
+\newsaj{She was fine with it.
+I think there are subtle differences.
+the RQ1 tries corresponds to quantifiable performance metric like accuracy, latency.
+RQ2 is qualitative feedback of user experience.}
 \squishlist
 \item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
-\saj{How did the addition of an overview like \noah affect a spreadsheet users navigation performance?}
+\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of speed and accuracy?}}
 \item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
-\saj{How did \noah and it's components impact spreadsheet users'
+\saj{How did \noah and it's components affect spreadsheet users'
 navigational experiences?}
@@
  \vspace{-10pt}
-\caption{Quiz tasks for the birdstrikes dataset.}
+\caption{\newsaj{Quiz tasks for the birdstrikes dataset. The use cases correspond to the task typology discussed in Section~\ref{sec:usage}}}
 \label{tab:questions}
 \centering
-\begin{tabular}{l l}
+\begin{tabular}{l l l}
 \hline
-\multicolumn{1}{c}{Category} & \multicolumn{1}{c}{Question}    \\ \hline
-steer                & Organize the data by State. How many flights that \\
-                         & had damages (damage = 1) originated from Florida? \\ 
-identify     & How many flights in the currently visible \\
-                 &  spreadsheet window have damages?                                    \\ 
-steer                & How many flights that had damages (damage = 1) \\
-                         &   originated from California?                          \\ 
-\cmpA & Which state between Florida and California \\
-                         &   has a higher number of flights with damages?    \\ 
-\cmpB & Find the state with the most birdstrike occurrences.  \\ 
-customize       & Organize the data by \emph{altitude}. What is the average \\
-                         &   cost of damages for altitude bin 0-450?          \\ \hline
+Category & Question & Use case    \\ \hline
+steer                & Organize the data by State. How many& \\
+                     & flights that had damages (damage = 1)& \code{summarize}\\
+                     & originated from Florida? \\ 
+identify             & How many flights in the currently &\\
+                     & visible spreadsheet window  &  \code{identify}\\ 
+                     & have damages? & \\
+steer                & How many flights that had damages & \code{summarize}\\
+                     & originated from California?                         & \\ 
+\cmpA                & Which state between Florida and &\\
+                     & California  has a higher number of & \code{compare}\\ 
+                     & flights with damages?    &\\
+\cmpB                & Find the state with the most & \code{compare}\\
+                     & birdstrike occurrences.  &\\ 
+customize            & Organize the data by \emph{altitude}. &\\
+                     & What is the average cost of damages & \code{generate}\\
+                     &    for altitude bin 0-450?         & \\ \hline
 \end{tabular}
@@
 \emph{Quiz Tasks.} \saj{We designed six tasks grouped into four categories: 
-steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} uses cases \code{summarize}, \code{compare}, \code{identify}, and \code{create}, respectively.
+steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} uses cases \code{summarize}, \code{compare}, \code{identify}, and \code{generate}, respectively.
 All of the tasks except the customize task, are representative of the navigation interactions required for the most frequently issued spreadsheet formulae
@@
 \subsection{Study Limitations.}
-\kgk{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies.
+\saj{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies.
 One limitation arises from our participant demographics.
@@
 We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} due to their limitations in the types of interactions supported (see Section~\ref{sec:related}).
-Moreover, we wanted the participants to utilize the typical spreadsheet operations that they are comfortable with to complete those tasks and see how introduction of \noah affected their navigation experience.}
+Moreover, instead of forcing the participants use unfamiliar spreadsheet operations, we wanted them to utilize the typical spreadsheet operations that they are comfortable with, to complete the tasks and see how introduction of \noah affected their navigation experience.}
 

### 2019-10-30 usecases.tex
aditya comments addressed

@@
 systems in over two dozen papers, 
-and has been applied to a variety of 
-scenarios, including XXX\agp{argue for
+and can be applied to a variety of 
+scenarios, including \newsaj{analytical reasoning during exploration~\cite{shrinivasan2008supporting}, multi-dimensional data analysis~\cite{valiati2006taxonomy}, graph comprehension~\cite{friel2001making}, cartography~\cite{roth2012empirically}, among others.}\agp{argue for
  utility with citations.}

### 2019-10-31 letter.tex
aditya comments addressed

@@
 The paper co-chairs mentioned the following: ``\textit{We particularly encourage such revisions where submissions were positively received by reviewers, but the revisions required were deemed to be beyond the scope of the conference review cycle.
-If you address the issues raised and subsequently submit to TVCG, please make reference to this  InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}''
-
-The present version of the paper addresses all of the issues mentioned by the reviewers.
-The entire set of reviews are attached.
+If you address the issues raised and subsequently submit to TVCG, please make reference to this  InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}'' The entire set of reviews are attached.
 In particular, the reviewers asked for the following changes for a revision:
@@
 We have spent the last four months 
-preparing a revised version of this paper.
-\agp{One thing that is confusing is that the bulletpoints above
-don't neatly map to the bulletpoints below, or the enumerated list after that. Stopped reading after this}
-Here, we provide a response to the reviews. 
-In our revised submission,
-(i) we discuss the merits of having an interface like \noah
-integrated within an existing spreadsheet, as opposed
-to other general-purpose exploratory tools such as  Keshif (\textbf{C1});
-(ii) we justify our design choice surrounding the binning mechanism
-while discussing its limitations (\textbf{C2}); 
-(iii) we enrich the user study design section further 
-by providing justifications for comparing \noah and Excel (\textbf{C3}), 
-explaining the choice of quiz tasks (\textbf{C4}), rephrasing our research questions reflecting the purpose of the user study (\textbf{C5}), and reporting the intra-participant differences for the study (\textbf{C6}).
-Moreover, to answer the rephrased research questions, we 
-have reorganized the results section to highlight key contributions of \noah while explaining participant experiences in using the tool.
-The major changes to the paper include:
+preparing a revised version of this paper. We thank the reviewers and meta-reviewer for their detailed,
+constructive feedback. 
+We have taken the feedback very seriously in preparing
+the revised version, and we believe the paper is substantially
+stronger as a result.
+\newsaj{The revision necessitated a number of updates to the content of each section and are highlighted in teal color.
+In Section 1, we cited existing work that highlight the widespread adoption of spreadsheets even among users of advanced enterprise solutions, adding further merit to our goal of addressing navigation challenges within spreadsheets.
+Recent discourse within the visualization community also echoes our sentiment\footnote{https://twitter.com/FILWD/status/1187411664611749888}---instead of designing a new tool that may cater to a small population, we aim to improve user experience with spreadsheet tools boasting more than $700$ million users.
+Moreover, we augmented the definitions of spreadsheet interactions like scrolling and steering and clearly articulated the challenges of designing a general purpose plug-in for spreadsheets.
+In Section 2, we defined the scope of tasks supported by spreadsheets guided by existing taxonomy of abstract visualization tasks.
+We further enhanced our usage scenario by highlighting how different use cases of \noah maps to the scope defined for spreadsheets.
+In Section 5, we added additional discussions to clearly articulate the difference in the goals of spreadsheets spreadsheets and TDA tools which further justifies choice of Excel, the most widely used spreadsheet tool, as our baseline in the evaluation study.
+In Section 4, we added further clarification to the third design consideration (\textbf{DC3}).
+In Section 5, we provided justifications for developing a multi-granularity overview while explaining why we opted for histogram as an overview representation.
+In Section 6, we further clarified the goals of the evaluation study while rephrasing the research questions to better reflect those goals.
+We didn't alter our previous research questions, rather grouped a few research questions into a more high level and less ambiguous question.
+Moreover, we explained our choice of the quiz tasks and discussed three limitations of our study in detail.
+We revamped Section 7 completely to better reflect our research questions while adding more qualitative observations regarding the user experience with \noah.
+We also included an additional analysis on the intra-participant differences.
+In Section 8, we first added a summary of our takeaways from the evaluation study while discussing additional limitations of \noah and future enhancement opportunities.
+We now discuss how the updates made to all of these sections addressed the changes suggested by the reviewers.}
 
@@
 
-\item \textit{Merits over other tools} (\textbf{C1}): \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation interface}, the concepts of which can be implemented and integrated as a plug-in to any spreadsheet.
-As spreadsheets have a massive user-base, enhancing the experience surrounding data exploration and formula computation while maintaining the spreadsheet look and feel as much as possible, has the potential to impact a lot of users.
-As we mentioned earlier, even users of enterprise solutions prefer Excel for direct data manipulation.
-Therefore, developing a solution to improve the experience with basic operations of such a tool has its merits.}
+\item \textit{Merits over other tools} (\textbf{C1}): \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation plug-in} to any existing spreadsheet.
+As spreadsheets have a massive user-base, enhancing exploration and formula computation on large datasets while maintaining the spreadsheet look and feel as much as possible, has the potential to impact a lot of users.
+Most of these users employ spreadsheets as their primary data management and analysis tool while shunning enterprise solutions with more advanced features.
+Therefore, our goal is to develop a solution to improve navigation within spreadsheets.}
 
-\item \textit{Explanation of the binning mechanism} (\textbf{C2}): We have explained our choice of the binning mechanism in the context the third design consideration (\emph{DC3}) and can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}.
-\saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge the limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss}.
+\item \textit{Explanation of the binning mechanism} (\textbf{C2}): We have explained our choice of the binning mechanism in the context of the third design consideration (\emph{DC3}).
+This explanation can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}.
+\saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss} in the context of our user study findings.
 
 \item \textit{Justification of Excel as a Baseline} (\textbf{C2}):
-We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets in interacting with data even for the users of enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and the TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to facilitate direct manipulation and arbitrary derivation of the data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which, far exceeds that of the TDA tools (see Section~\ref{sec:study_design}).
+We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets among users who shun more advanced enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to present the raw data as is, amenable to editing, formula computation, and comparison of derived data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which far exceeds that of TDA tools (see Section~\ref{sec:study_design}).
 
 \item \textit{Enhancing the user study design} (\textbf{C4, C5}):
-We have added additional explanations on the goal of the study while better articulating the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
+We have added additional explanations regarding the goal of the study while better articulating the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
 \saj{Specifically, we merged a number of research questions that explored how specific features of \noah affected participants' navigation experience into one broad question which explores their overall experience with \noah and its components.
-Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}, and b) explaining how these tasks can reveal how people may utilize \noah for spreadsheet navigation.
-Finally, we have identified two limitations of the study , resulting from sample size and the study design, which we justify based on the scope of the study.}
+Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), and b) explaining how these tasks can reveal how people may use \noah for spreadsheet navigation.
+Finally, we have identified two limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.}
 
@@
 
-\saj{In addition to the aforementioned changes, we have reorganized Section~\ref{sec:results} to reflect the rephrased research questions while supporting our findings from the study by providing qualitative observations, such as participants' comments and interaction patterns with the \noah.
+\saj{In addition to the aforementioned changes, we have reorganized Section~\ref{sec:results} to reflect the rephrased research questions.
+Furthermore, we discuss our findings from the study in the context of the research questions by providing qualitative observations, such as participants' comments and interaction patterns with the \noah.
 We have explained how different features of \noah affected participants' spreadsheet navigation experience, either positively or negatively.

### 2019-10-31 letter.tex
aditya comments addressed

@@
 \end{enumerate}
-
-\saj{In addition to the aforementioned changes, we have reorganized Section~\ref{sec:results} to reflect the rephrased research questions.
-Furthermore, we discuss our findings from the study in the context of the research questions by providing qualitative observations, such as participants' comments and interaction patterns with the \noah.
-We have explained how different features of \noah affected participants' spreadsheet navigation experience, either positively or negatively.
-Based on these observations, we have updated Section~\ref{sec:discuss} to include additional limitations of \noah while highlighting future research opportunities.}

### 2019-10-31 letter.tex
substantial changes to letter

@@
 \end{itemize} 
-We have spent the last four months 
-preparing a revised version of this paper. We thank the reviewers and meta-reviewer for their detailed,
+\new{We thank the reviewers and meta-reviewer for their detailed,
 constructive feedback. 
-We have taken the feedback very seriously in preparing
-the revised version, and we believe the paper is substantially
+Taking this feedback into account, 
+we have spent the last four months 
+preparing a revised version of this paper.
+We believe the paper is substantially
 stronger as a result. 
-\newsaj{The revision necessitated a number of updates to the content of each section and are highlighted in teal color.
-In Section 1, we cited existing work that highlight the widespread adoption of spreadsheets even among users of advanced enterprise solutions, adding further merit to our goal of addressing navigation challenges within spreadsheets.
-Recent discourse within the visualization community also echoes our sentiment\footnote{https://twitter.com/FILWD/status/1187411664611749888}---instead of designing a new tool that may cater to a small population, we aim to improve user experience with spreadsheet tools boasting more than $700$ million users.
-Moreover, we augmented the definitions of spreadsheet interactions like scrolling and steering and clearly articulated the challenges of designing a general purpose plug-in for spreadsheets.
-In Section 2, we defined the scope of tasks supported by spreadsheets guided by existing taxonomy of abstract visualization tasks.
-We further enhanced our usage scenario by highlighting how different use cases of \noah maps to the scope defined for spreadsheets.
-In Section 5, we added additional discussions to clearly articulate the difference in the goals of spreadsheets spreadsheets and TDA tools which further justifies choice of Excel, the most widely used spreadsheet tool, as our baseline in the evaluation study.
-In Section 4, we added further clarification to the third design consideration (\textbf{DC3}).
-In Section 5, we provided justifications for developing a multi-granularity overview while explaining why we opted for histogram as an overview representation.
-In Section 6, we further clarified the goals of the evaluation study while rephrasing the research questions to better reflect those goals.
-We didn't alter our previous research questions, rather grouped a few research questions into a more high level and less ambiguous question.
-Moreover, we explained our choice of the quiz tasks and discussed three limitations of our study in detail.
-We revamped Section 7 completely to better reflect our research questions while adding more qualitative observations regarding the user experience with \noah.
-We also included an additional analysis on the intra-participant differences.
-In Section 8, we first added a summary of our takeaways from the evaluation study while discussing additional limitations of \noah and future enhancement opportunities.
-We now discuss how the updates made to all of these sections addressed the changes suggested by the reviewers.}
+Every section in the paper has been revised;
+the changes are highlighted in blue.
+We briefly describe the changes to each section
+before returning to the changes required in the revision 
+({\bf C1}--{\bf C6}).
+\subsection*{Changes Organized by Section}
+\begin{itemize}
+	\leftmargin=25pt \rightmargin=0pt \labelsep=5pt \labelwidth=20pt \itemindent=0pt \listparindent=0pt \topsep=0pt plus 2pt minus 4pt \partopsep=0pt plus 1pt minus 1pt \parsep=0pt plus 1pt \itemsep=\parsep
+\item 
+In Section~\ref{sec:intro}, we now cite
+work that highlights the widespread adoption
+of spreadsheets even among users of 
+advanced enterprise solutions, further justifying 
+why we focused on addressing navigation challenges within
+spreadsheets. In fact, recent 
+debate within the visualization community following
+VIS 2019 also echoes this view\footnote{https://twitter.com/FILWD/status/1187411664611749888}---instead of designing a new sophisticated tool from scratch that 
+may cater to a small population, 
+we aim to enhance the user experience for
+existing spreadsheet tools
+with nearly a billion users.
+Furthermore, in Section~\ref{sec:intro},  
+the definitions of spreadsheet 
+interactions like scrolling and steering have been augmented, and 
+we now clearly articulate the challenges of 
+designing a general-purpose plug-in for spreadsheets.
+\item 
+In Section~\ref{sec:usage}, 
+our usage scenario has been substantially revised 
+using Brehmer and Munzner's typology~\cite{brehmer2013multi} to
+clarify the scope of tasks supported by 
+a \noah-enhanced spreadsheet; our new Table~\ref{tab:scope}
+provides use-cases supported within \noah
+for each category.
+\item
+In Section~\ref{sec:related},
+we now clearly
+articulate the differences in goals
+of spreadsheets and tabular data analysis tools (TDA),
+further justifying our choice of Excel,
+the most widely used spreadsheet tool,
+as our baseline in the evaluation study.
+\item 
+In Section~\ref{sec:design},
+we now explain the  
+third design consideration 
+for \noah more clearly
+(\textbf{DC3})---focusing on motivating the binning
+mechanism. 
+\item In Section~\ref{sec:ui}, 
+we now justify why we are developing a 
+multi-granularity overview,
+and explain why 
+we opted for histograms as an overview representation. 
+\item 
+In Section~\ref{sec:study}, 
+we further clarified the goals of our study
+and reframed our research questions to 
+better reflect those goals. 
+Our research questions haven't been substantially altered;
+rather, they have been grouped together and made less
+ambiguous. In this version,
+we now explain our choice
+for the quiz tasks and discuss 
+three limitations of our study in detail. 
+\item Section~\ref{sec:results} has been completely
+revamped to better reflect our research questions,
+while adding more qualitative observations 
+regarding the user experience with \noah. 
+We now
+include additional analysis results on intra-participant differences. 
+\item In Section~\ref{sec:discuss}, we have added a summary of our takeaways from the evaluation study, while discussing additional limitations of \noah and future enhancement opportunities. 
+\end{itemize}
+\subsection*{Changes Organized by Reviewer Concerns (C1--C6)}
+}
 
@@
 
-\item \textit{Merits over other tools} (\textbf{C1}): \saj{In the paper, we have identified the major contribution of \noah as its \emph{design as a general purpose navigation plug-in} to any existing spreadsheet.
-As spreadsheets have a massive user-base, enhancing exploration and formula computation on large datasets while maintaining the spreadsheet look and feel as much as possible, has the potential to impact a lot of users.
-Most of these users employ spreadsheets as their primary data management and analysis tool while shunning enterprise solutions with more advanced features.
-Therefore, our goal is to develop a solution to improve navigation within spreadsheets.}
+\item \textit{Merits over other tools} (\textbf{C1}): 
+\new{Throughout the paper (and especially in 
+Sections~\ref{sec:intro}, \ref{sec:related},
+and \ref{sec:study}), we have emphasized
+that the main contribution of \noah
+is its \emph{design as a general-purpose navigation plug-in} 
+to any existing spreadsheet system. 
+As spreadsheets have a massive user-base, 
+enhancing exploration and formula computation 
+on large datasets while maintaining 
+the spreadsheet look-and-feel as much as possible, 
+has the potential to impact hundreds of millions 
+of spreadsheet users. 
+Most of these users employ spreadsheets 
+as their primary data management and analysis 
+tool while shunning enterprise solutions 
+with more advanced features. 
+Therefore, our goal is to develop a 
+solution to improve navigation within spreadsheets.
+Indeed, we could have tried to enhance navigation
+in other tools, such as Keshif, or Tableau, like
+the reviewers suggest---but we would be forcing
+spreadsheet users to adopt an entirely new tool, something
+that they are clearly loath to do.}
 
-\item \textit{Explanation of the binning mechanism} (\textbf{C2}): We have explained our choice of the binning mechanism in the context of the third design consideration (\emph{DC3}).
-This explanation can be found under the title ``Why a Multi-granularity Binned Overview'' in Section~\ref{sec:design}.
-\saj{In brief, we opted for binning to provide a clear and concise representation of the overall data distribution while minimizing user's back and forth movement across multiple screens during navigation.} However, we acknowledge limitation of this choice for binning categorical data and discuss possible solutions in Section~\ref{sec:discuss} in the context of our user study findings.
+\item \textit{Explanation of the binning mechanism} (\textbf{C2}): 
+\new{We now explain our choice
+of the binning mechanism in 
+the context of the third design consideration (\emph{DC3}), 
+as detailed in Section~\ref{sec:design}.
+We further expand on this in the blurb titled 
+``Why a Multi-granularity Binned Overview'' 
+in Section~\ref{sec:ui}.
+In brief, we opted for binning to 
+provide a clear and concise 
+representation of the overall data 
+distribution while minimizing user's back 
+and forth movement across multiple screens 
+during navigation. 
+At some level, we do need to limit the number
+of values displayed so that the overview
+fits on the screen: binning is a natural solution
+for that issue. 
+\agp{Possibly argue using the immens paper?}
+However, we acknowledge limitations 
+for this choice for binning categorical data 
+and discuss possible solutions 
+in Section~\ref{sec:discuss} 
+in the context of our user study findings. 
+}
 
 \item \textit{Justification of Excel as a Baseline} (\textbf{C2}):
-We have provided justification for using Excel over tabular data analysis (TDA) tools a) by highlighting the appeal of spreadsheets among users who shun more advanced enterprise solutions (see Section~\ref{sec:intro}), b) by identifying the scope of tasks supported by spreadsheets and TDA tools (see Section~\ref{sec:usage} and~\ref{sec:related}), c) by explaining that the goal of spreadsheets is to present the raw data as is, amenable to editing, formula computation, and comparison of derived data unlike TDA tools that hide the data while providing summarized statistics (see Section~\ref{sec:related}), and d) by highlighting the importance of addressing the shortcomings of a tool, the user base of which far exceeds that of TDA tools (see Section~\ref{sec:study_design}).
+\new{Since our goal is to improve spreadsheet
+user experience---and there are very good reasons for doing so,
+as outlined in item 1) above---the natural comparison point
+is Excel. 
+That said, beyond popularity, 
+we have provided a thorough justification
+for what spreadsheet systems like Excel offer
+relative to tabular data analysis tools (TDA),
+making it a more appropriate point of comparison.
+Our justification for using Excel over 
+tabular data analysis (TDA) tools involves 
+a) highlighting the appeal of spreadsheets 
+among users who shun more advanced enterprise 
+tools (Section~\ref{sec:intro}), 
+b) identifying the scope of tasks 
+supported by spreadsheets and 
+TDA tools (Section~\ref{sec:usage} and~\ref{sec:related}), 
+c) explaining that the goal of spreadsheets 
+is to present the raw data as is, 
+amenable to editing, formula computation, and comparison of derived data, all done in-situ,
+unlike TDA tools that hide the data 
+while providing summarized statistics 
+(Section~\ref{sec:related}), and 
+d) highlighting the importance of 
+addressing the shortcomings of a tool, 
+the user base of which far exceeds 
+that of TDA tools (Section~\ref{sec:study}). 
+}
 
-\item \textit{Enhancing the user study design} (\textbf{C4, C5}):
-We have added additional explanations regarding the goal of the study while better articulating the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
+\item \textit{Enhancing the user study design} 
+(\textbf{C4, C5}):
+\agp{Revise to clarify where these changes are made
+and HOW these changes solve the issue. Just saying
+we merged doesn't necessarily address the issue.}
+We have added additional 
+explanations regarding the goal 
+of the study while better articulating 
+the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
 \saj{Specifically, we merged a number of research questions that explored how specific features of \noah affected participants' navigation experience into one broad question which explores their overall experience with \noah and its components.
@@
 \item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
-Finally, we have performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data.
-\saj{For the latter, we find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah.
+\new{We performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data: this new analysis can be found in XXX\agp{pls add; insights on trends sounds vague let's be more clear}.} \saj{For the latter, we find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah.
 However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}.

### 2019-10-31 introduction.tex
changes until related

@@
 and complement existing spreadsheet interactions?
-How to we enable coordinated interactions across both views, \ie 
+How do we enable coordinated interactions across both views, \ie 
 the overview and the raw spreadsheet, such that they

### 2019-10-31 related.tex
changes until related

@@
 \label{sec:related}
+\new{
 We now discuss tools and techniques
-that partially address the limitations of navigating
-spreadsheets.
+that partially address the limitations 
+of navigating
+data, both inside and outside spreadsheets.
+}
 
@@
 targeted at improving users' navigation experience.}
-\agp{Abrupt, add at least one line to say that existing spreadsheet tools recognize challenges in nav or something}
-\newsaj{Existing spreadsheet tools---both commercial software and academic prototypes---recognize some of these navigation challenges while adopting different measures to address those.}
+% \agp{Abrupt, add at least one line to say that existing spreadsheet tools recognize challenges in nav or something}
+\new{Both commercial spreadsheet tools
+as well as academic prototypes provide
+partial solutions to navigational challenges.}
 
@@
 %challenges in other domains.
+\new{We draw from work on navigation interfaces
+in non-spreadsheet interfaces as well.}
 
@@
 \stitle{Multiple Coordinated Views.} Multiple coordinated views~\cite{roberts2007state}, \eg \emph{Snap}~\cite{north2000snap}, \emph{Elastic Documents}~\cite{badam2018elastic} connect multiple views, for example, an overview and a detailed data view while enabling coordination between these views through brushing and linking.
-Similarly, \noah connects tabular spreadsheets with an overview and updates the state of the spreadsheet as users interact with the overview and vice-versa.
+Similarly, \noah connects spreadsheets with an overview and updates the spreadsheet as users interact with the overview and vice-versa.
 
 \stitle{Tabular Data Analysis (TDA).}
-Visualization tools such as Tableu~\cite{tableau}, Power BI~\cite{powerbi}, Keshif~\cite{yalccin2018keshif}, Voyager~\cite{wongsuphasawat2016voyager} and
+Visualization tools such as Tableau~\cite{tableau}, Power BI~\cite{powerbi}, Keshif~\cite{yalccin2018keshif}, Voyager~\cite{wongsuphasawat2016voyager} and
 analytical tools such as SPSS~\cite{noruvsis1986spss}, SAS~\cite{sas1985sas}, can all provide summaries of tabular data in various forms (visualizations,
@@
 the underlying raw data. 
-\newsaj{For example, Keshif~\cite{yalccin2018keshif} can display all the unique values corresponding to an attribute of interest, \eg cities of the Airbnb data~\cite{web:airbnb}.
-However, users cannot view or inspect the raw data corresponding to each city.
+For example, Keshif~\cite{yalccin2018keshif} can display all the unique values corresponding to an attribute of interest, \eg cities of the Airbnb data~\cite{web:airbnb}.
+However, users cannot view or inspect the raw data corresponding to each city
+in a spreadsheet-like tabular setting, while being able to
+edit this raw data at will. 
 With TDA tools, the spreadsheet look-and-feel is lost,
 and as a result, users
-loose the capability to directly manipulate raw data, derive new data, and issue formulae for free-form analysis.}  
+lose the ability to directly manipulate raw data, derive new data, and issue formulae for free-form analysis.  
 \cut{For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}).
@@
 because the citation is to an AQP tool rather than some general purpose study}
-\newsaj{Therefore, the goals of spreadsheets differ from TDA 
+Therefore, the goals of spreadsheets differ from TDA 
 tools in two ways: a) facilitating direct manipulation of raw 
-data and b) enabling arbitrary derivation of new data and summaries
+data in-situ and b) enabling arbitrary derivation of new data and summaries
 using various operations involving navigation,
-\eg formulas or expressions.}
+\eg formulae or expressions\agp{expressions?}.
 \noah being a plug-in to spreadsheets, 
-facilitates a unified interface that
+provides a unified interface that
 upholds both these goals

### 2019-10-31 usecases.tex
changes until related

--- 2019-10-11 usecases.tex
@@
 and has been applied to a variety of 
-scenarios, including XXX\agp{argue for
- utility with citations.}
+scenarios, including \newsaj{analytical reasoning during exploration~\cite{shrinivasan2008supporting}, multi-dimensional data analysis~\cite{valiati2006taxonomy}, graph comprehension~\cite{friel2001making}, cartography~\cite{roth2012empirically}, among others.}\agp{argue for
+ utility with citations.} \agp{This doesn't cut it because i wanted to argue that it HAS been used, not that it can be used. So can we find citations post 2013?.}
 While all the tasks in Table~\ref{tab:scope} 
@@
 \vspace{-10pt}
-\caption{Example use cases of \noah, employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
+\caption{Example use cases where \noah provides benefits beyond spreadsheets (labeled by \checkmark if improved; $\times$ if it
+remains the same), employing Brehmer and Munzner's typology~\cite{brehmer2013multi}.}
 \label{tab:scope}

### 2019-11-01 design.tex
s1-3 and letter comments

@@
 that abstracts the data at varying levels of detail.
-Multi-granularity interfaces
-provide an alternative to \newsaj{the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
+\newsaj{Multi-granularity representations 
+have shown to
+scale better to 
+larger datasets----presenting information 
+at multiple granularities
+makes visual representations
+more perceptually scalable
+and less cluttered~\cite{elmqvist2009hierarchical}.
+The multi-granularity overview of \noah
+provides an alternative to the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
 Users can resize the overview to control the amount of spreadsheet data that remains visible.
@@
 database query optimizers~\cite{chaudhuri1998random}.
-The binned representation of the overview
+The binned representation 
+of the overview
 using equi-depth histograms 
-enables faster navigation
-by reducing visual discontinuity. 
-The bins act as landmarks
+aggregates data 
+and visualizes density 
+by counting 
+the number of data points 
+falling within each bin. 
+Binned aggregation 
+highlights both high level (\eg densities) and low level (e.g., outliers)
+details of large datasets, 
+while enabling
+multi-granularity representation of data
+via the choice of bin size~\cite{liu2013immens}. 
+Moreover, binning reduces 
+visual discontinuity during navigation
+as users are able to 
+view an overview 
+that fits in the computer screen.
+As a result, users can 
+quickly navigate the data---the bins act as landmarks
 in the overview, enabling 

### 2019-11-01 letter.tex
changes until related
s1-3 and letter comments

@@
 using Brehmer and Munzner's typology~\cite{brehmer2013multi} to
-clarify the scope of tasks supported by 
-a \noah-enhanced spreadsheet; our new Table~\ref{tab:scope}
+clarify the scope of tasks supported by spreadsheets and enhanced
+by a spreadsheet with a \noah plug-in; our new Table~\ref{tab:scope}
 provides use-cases supported within \noah
@@
 \agp{Possibly argue using the immens paper?}
+\newsaj{We cite related work on 
+multi-scale aggregation~\cite{elmqvist2009hierarchical} 
+to motivate the choice of 
+a multi-granularity overview
+and show that binned aggregation via histogram~\cite{liu2013immens}
+is a suitable representation of such an overview.}
 However, we acknowledge limitations 
@@
 
-\item \textit{Justification of Excel as a Baseline} (\textbf{C2}):
+\item \textit{Justification of Excel as a Baseline} (\textbf{C3}):
 \new{Since our goal is to improve spreadsheet
@@
 the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
-\saj{Specifically, we merged a number of research questions that explored how specific features of \noah affected participants' navigation experience into one broad question which explores their overall experience with \noah and its components.
+\newsaj{Specifically, we merged our previous \emph{RQ1} (evaluating the quiz task performance) with \emph{RQ4} (summarizing user feedback on performance via survey) into a unified research question (new \emph{RQ1}) on participants' navigation performance.
+Moreover, we merged a previous \emph{RQ2} and \emph{RQ3} that explored how specific features of \noah affected participants' navigation experience into one broad question (new \emph{RQ2}) which explores their overall experience with \noah and its components}.
 Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), and b) explaining how these tasks can reveal how people may use \noah for spreadsheet navigation.
-Finally, we have identified two limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.}
+Finally, we have identified three limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.
 
 \item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
-\new{We performed additional analysis of the study data to report intra-participant submission time differences while providing insights on the trends that emerged from analyzing the data: this new analysis can be found in XXX\agp{pls add; insights on trends sounds vague let's be more clear}.} \saj{For the latter, we find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah.
-However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}.
-Therefore, we have included the results of the intra-participant analysis in appendix while briefly summarizing its outcomes in Section~\ref{sec:rq1}.}
+\newsaj{We performed additional analysis of the study data to report intra-participant submission time differences for the quiz phase tasks.
+While we summarize the results of the analysis in Section~\ref{sec:time_acc}, the detailed result can be found in Appendix}\agp{pls add; insights on trends sounds vague let's be more clear}.
+\newsaj{We find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah---all but one participants completed at least four tasks in less time using \noah compared to Excel.} However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}.
+Therefore, we have included the results of the intra-participant analysis in Appendix.
 \end{enumerate}

### 2019-11-01 results.tex
s1-3 and letter comments

@@
 \subsubsection{Faster navigation without sacrificing accuracy}
+\label{sec:time_acc}
 In Figure~\ref{fig:timeBox}a and~\ref{fig:timeBox}b, 
@@
 made spreadsheet navigation faster for these tasks.
-We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations.
+\saj{We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations.
 Overall, majority of the submission times using \noah were faster compared to Excel---19 out of the 20 participants completed at least four tasks in less time using \noah compared to Excel.
@@
 Moreover, for all tasks except customize, the difference in submission times was statistically significant.
-Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.
+Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.}
 

### 2019-11-01 study.tex
s1-3 and letter comments

@@
 \item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
-\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of speed and accuracy?}}
+\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of navigation task submission speed and accuracy and quantitative survey metrics such as speed of use?}}
 \item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 

### 2019-11-01 usecases.tex
munzner ref added

@@
 and has been applied to a variety of 
-scenarios, including \newsaj{analytical reasoning during exploration~\cite{shrinivasan2008supporting}, multi-dimensional data analysis~\cite{valiati2006taxonomy}, graph comprehension~\cite{friel2001making}, cartography~\cite{roth2012empirically}, among others.}\agp{argue for
+scenarios, including \newsaj{visualization system design and validation models~\cite{meyer2015nested}, designing task taxonomy for cartograms~\cite{mcnabb2017survey}, and defining the scope of tasks in a number of domains, \eg interactive task authoring~\cite{fulda2015timelinecurator}, document mining~\cite{brehmer2014overview}, multivariate network analysis~\cite{pretorius2014tasks}, mass cytometry~\cite{hollt2016cytosplore}, among others.}\agp{argue for
  utility with citations.} \agp{This doesn't cut it because i wanted to argue that it HAS been used, not that it can be used. So can we find citations post 2013?.}

### 2019-11-11 design.tex
aditya comments on S5, S6

--- 2019-10-30 design.tex
@@
+%!TEX root = main.tex
+
 \section{\noah: Design Considerations}
@@
 informed by prior work on
-information visualization~\cite{shneiderman2003eyes,brehmer2013multi}, overview+detail interfaces~\cite{cockburn2009review}, multiple-coordinated views~\cite{wang2000guidelines},
+information visualization~\cite{shneiderman2003eyes,brehmer2013multi}, 
+overview+detail interfaces~\cite{cockburn2009review}, multiple-coordinated views~\cite{wang2000guidelines},
 and refined through our experiences
@@
 However, maintaining the overview in a separate
-location from the data can \newsaj{lead to loss of context};
+location from the data can \new{lead to loss of context};
 instead, having it co-located with the data
@@
 and study their properties in detail, \eg via steering.
-Navigating back and forth \saj{between different subsets of data}
+Navigating back and forth \new{between different subsets of data}
 can lead to increased visual discontinuity.
@@
 we need to consider the trade-off between visual discontinuity (\emph{DC2}) and clarity.
-\saj{Displaying a fine-grained overview improves visual clarity while increasing visual discontinuity---users need to scroll through the overview to access distant
+\new{Displaying a fine-grained overview improves visual clarity 
+while increasing visual discontinuity---users need to scroll through the overview to access distant
 subsets of data.
-Displaying a coarse-grained overview decreases visual discontinuity at cost of reduced visual clarity---the overview may span too many data subsets and appear visually cluttered}.
+Displaying a coarse-grained overview decreases visual discontinuity 
+at cost of reduced visual clarity---the overview may 
+span too many data subsets and appear visually cluttered.}
 The interface should further allow users to control the screen-space allocated
@@
 semantic zooming in/out~\cite{perlin1993pad}.
-\noah supports other interactions,
+\noah supports other interactions \new{atop this multi-granularity overview,}
 \eg customization and aggregation.
@@
 \stitle{Why a Multi-granularity Binned Overview?}
-\saj{A standard overview approach
-is to show a spatially partitioned
-collection of thumbnails \saj{on the left of the detailed view},
-similar to Microsoft PowerPoint or
+\new{A conventional design for overviews within
+popular interfaces is as a spatially partitioned
+collection of thumbnails 
+on the left of the standard detailed view,
+similar to Microsoft Power Point or
 Adobe Reader.
@@
 that abstracts the data at varying levels of detail.
-Multi-granularity interfaces
-provide an alternative to \newsaj{the spatially partitioned single-granularity representation of the data space, \eg in Powerpoint, by allowing users to control the scale at which the overview should be displayed~\cite{cockburn2009review}.}
+Multi-granularity representations have been shown to
+scale better to 
+larger datasets----presenting information 
+at multiple granularities
+makes visual representations
+more perceptually scalable
+and less cluttered~\cite{elmqvist2009hierarchical}.
+Thus, the multi-granularity overview of \noah
+provides an alternative to the aforementioned conventional 
+spatially partitioned 
+single-granularity representation of the data space, \eg in Power Point, 
+by allowing users to control the scale at 
+which the overview should be displayed~\cite{cockburn2009review}.
 Users can resize the overview to control the amount of spreadsheet data that remains visible.
@@
 
-\newsaj{The data structure underlying the overview
+\new{The data structure underlying the overview
 is an equi-depth histogram constructed
 on the values
-in the navigation attribute column.
+in the navigation attribute column.\agp{Briefly describe what equi-depth histograms are; or don't mention it here, just say histogram -- we can defer details later}
 Equi-depth histograms are commonly
@@
 of data, with applications such as
-database query optimizers~\cite{chaudhuri1998random}.
-The binned representation of the overview
+database query optimizers~\cite{chaudhuri1998random}. \agp{only one application?}
+The binned representation 
+of the overview
 using equi-depth histograms 
-enables faster navigation
-by reducing visual discontinuity. 
-The bins act as landmarks
+aggregates data 
+and visualizes density 
+by counting 
+the number of data points 
+falling within each bin. \agp{This previous sentence feels repetitive.}
+Binned aggregation 
+highlights both high level (\eg densities) and low level (e.g., outliers)
+details of large datasets, 
+while enabling
+multi-granularity representation of data
+via the choice of bin size~\cite{liu2013immens}. \agp{Clarify what is known from prior work and what
+you are proposing. Here, it is intermixed.}
+Moreover, binning reduces 
+visual discontinuity during navigation
+as users are able to 
+view an overview 
+that fits in the computer screen.
+As a result, users can 
+quickly navigate the data---the bins act as landmarks
 in the overview, enabling 
@@
 \stitle{Overview Construction.}
-\newsaj{The equi-depth histogram 
+\new{The equi-depth histogram 
 can be constructed on 
@@
 \saj{Saj: can we add a pseudocode in appendix and refer to that?}
+\agp{Yes please.}
 
@@
 One issue with zooming interactions
-is discoverability \saj{of the zoom operation}~\cite{cockburn2009review}.
+is discoverability \new{of the zoom operation}~\cite{cockburn2009review}.
 We circumvent this (see Figure~\ref{fig:concept}c)
@@
 may be reflected on the spreadsheet and vice-versa.
-\newsaj{One example of this coordinations is,} indicating the navigation attribute on the spreadsheet
+\new{One example of this coordinations is} indicating the navigation attribute on the spreadsheet
 using color (see the lime green column in Figure~\ref{fig:ux}c)
@@
 The coupling depends on the user's current
-focus---\emph{\saj{to ensure consistency between the overview and the spreadsheet,} any interaction on
+focus---\emph{\new{to ensure consistency between the overview and the spreadsheet,} any interaction on
 either interface
@@
 and provide the necessary formula parameters to a form.
-\newsaj{The aggregate column can employ any} statistical
-and mathematical formulae that
+\new{The aggregate column can employ any statistical
+or mathematical formulae that
 operate over a range of data.
-\newsaj{Therefore, creating an aggregate column
+Therefore, creating an aggregate column
 is equivalent to selecting subsets of data

### 2019-11-11 letter.tex
aditya comments on S5, S6

@@
 \end{itemize} 
-\new{We thank the reviewers and meta-reviewer for their detailed,
+We thank the reviewers and meta-reviewer for their detailed,
 constructive feedback. 
@@
 Every section in the paper has been revised;
-the changes are highlighted in blue.
+the changes are highlighted in \new{blue}.
 We briefly describe the changes to each section
@@
 with nearly a billion users.
-Furthermore, in Section~\ref{sec:intro},  
+We further augment   
 the definitions of spreadsheet 
-interactions like scrolling and steering have been augmented, and 
-we now clearly articulate the challenges of 
+interactions like scrolling and steering in Section~\ref{sec:intro}, 
+and now clearly articulate the challenges of 
 designing a general-purpose plug-in for spreadsheets.
@@
 using Brehmer and Munzner's typology~\cite{brehmer2013multi} to
-clarify the scope of tasks supported by spreadsheets and enhanced
-by a spreadsheet with a \noah plug-in; our new Table~\ref{tab:scope}
+clarify the scope of tasks supported by spreadsheets, as well
+as those that are enhanced
+when using a spreadsheet with a \noah plug-in; our new Table~\ref{tab:scope}
 provides use-cases supported within \noah
@@
 and explain why 
-we opted for histograms as an overview representation. 
+we opted for histograms as an overview representation, \new{by contextualizing our
+approach using prior work on multi-scale aggregation~\cite{elmqvist2009hierarchical}.} 
 \item 
@@
 Our research questions haven't been substantially altered;
-rather, they have been grouped together and made less
-ambiguous. In this version,
+rather, they have been grouped together and made more precise. 
+In this version,
 we now explain our choice
@@
 \subsection*{Changes Organized by Reviewer Concerns (C1--C6)}
-}
 
@@
 \item \textit{Merits over other tools} (\textbf{C1}): 
-\new{Throughout the paper (and especially in 
+Throughout the paper (and especially in 
 Sections~\ref{sec:intro}, \ref{sec:related},
@@
 spreadsheet users to adopt an entirely new tool, something
-that they are clearly loath to do.}
+that they are clearly loath to do.
 
 \item \textit{Explanation of the binning mechanism} (\textbf{C2}): 
-\new{We now explain our choice
+We now explain our choice
 of the binning mechanism in 
@@
 for that issue. 
-\agp{Possibly argue using the immens paper?}
-\newsaj{We cite related work on 
+We cite related work on 
 multi-scale aggregation~\cite{elmqvist2009hierarchical} 
@@
 a multi-granularity overview
-and show that binned aggregation via histogram~\cite{liu2013immens}
-is a suitable representation of such an overview.}
+and show that binned aggregation via histograms~\cite{liu2013immens}
+is a suitable representation of such an overview.
 However, we acknowledge limitations 
@@
 in the context of our user study findings.
-}
 
 \item \textit{Justification of Excel as a Baseline} (\textbf{C3}):
-\new{Since our goal is to improve spreadsheet
+Since our goal is to improve spreadsheet
 user experience---and there are very good reasons for doing so,
@@
 is to present the raw data as is, 
-amenable to editing, formula computation, and comparison of derived data, all done in-situ,
+amenable to editing, formula computation, 
+and comparison of derived data, all done in-situ,
 unlike TDA tools that hide the data 
@@
 that of TDA tools (Section~\ref{sec:study}).
-}
 
@@
 (\textbf{C4, C5}):
-\agp{Revise to clarify where these changes are made
-and HOW these changes solve the issue. Just saying
-we merged doesn't necessarily address the issue.}
 We have added additional 
-explanations regarding the goal 
+explanations regarding the goals 
 of the study while better articulating 
-the research questions, provided a better justification for the choice of the tasks for the study (see Section~\ref{sec:study_design}) and added a detailed discussion on the limitations of the study.
-\newsaj{Specifically, we merged our previous \emph{RQ1} (evaluating the quiz task performance) with \emph{RQ4} (summarizing user feedback on performance via survey) into a unified research question (new \emph{RQ1}) on participants' navigation performance.
-Moreover, we merged a previous \emph{RQ2} and \emph{RQ3} that explored how specific features of \noah affected participants' navigation experience into one broad question (new \emph{RQ2}) which explores their overall experience with \noah and its components}.
-Moreover, we have justified the choice of the quiz tasks by a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), and b) explaining how these tasks can reveal how people may use \noah for spreadsheet navigation.
-Finally, we have identified three limitations of the study, resulting from the sample size and the study design, which we justify based on the scope of the study.
+the research questions, 
+provided a better justification 
+for the choice of the tasks for the study (see Section~\ref{sec:study_design}) 
+and added a detailed discussion on the limitations of the study. 
+Specifically, we merged our previous \emph{RQ1} (evaluating the quiz task performance) 
+with \emph{RQ4} (summarizing user feedback on performance via a survey) 
+into a unified research question (new \emph{RQ1}) 
+on the participants' navigation performance. 
+Moreover, we merged previous \emph{RQ2} and \emph{RQ3} 
+that explored how specific features of \noah 
+affected participants' navigation experience 
+into one broad question (new \emph{RQ2}) 
+that explores their overall experience with \noah and its components. 
+Moreover, we have justified the choice of the quiz tasks by 
+a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), 
+and b) explaining how these tasks can reveal how people may use \noah 
+for spreadsheet navigation. 
+Finally, we have identified three limitations of the study, 
+resulting from the sample size and the study design, which we justify based on the scope of the study.
 

### 2019-11-11 usecases.tex
aditya comments on S5, S6

--- 2019-10-31 usecases.tex
@@
 and has been applied to a variety of 
-scenarios, including \newsaj{analytical reasoning during exploration~\cite{shrinivasan2008supporting}, multi-dimensional data analysis~\cite{valiati2006taxonomy}, graph comprehension~\cite{friel2001making}, cartography~\cite{roth2012empirically}, among others.}\agp{argue for
- utility with citations.} \agp{This doesn't cut it because i wanted to argue that it HAS been used, not that it can be used. So can we find citations post 2013?.}
+scenarios, including
+developing models for visualization system design~\cite{meyer2015nested}, 
+designing task taxonomies for cartograms~\cite{mcnabb2017survey}, 
+and defining the scope of tasks in a number of domains, 
+\eg interactive task authoring~\cite{fulda2015timelinecurator},
+ document mining~\cite{brehmer2014overview}, 
+ multivariate network analysis~\cite{pretorius2014tasks}, 
+ mass cytometry~\cite{hollt2016cytosplore}, among others.
 While all the tasks in Table~\ref{tab:scope} 

### 2019-11-11 design.tex
addressed Aditya s5 and 6 comments

@@
 
-\new{The data structure underlying the overview
-is an equi-depth histogram constructed
+\newsaj{The data structure underlying the overview
+is a histogram constructed
 on the values
-in the navigation attribute column.\agp{Briefly describe what equi-depth histograms are; or don't mention it here, just say histogram -- we can defer details later}
+in the navigation attribute column.
+Histograms are binned aggregates
+where each bin represents a collection (group) of data points and the frequency of the data points, \ie \code{COUNT} aggregate, visualizes the density of the corresponding group.} \agp{This previous sentence feels repetitive.} \saj{rephrased.
+wanted to explain that constructing histogram is a type of binned aggregation.}
+\newsaj{In addition to providing high level (\eg densities) and low level (e.g., outliers) details, binned aggregation techniques 
+enables multi-granularity visual representation of data
+by varying the bin size and
+have been deployed in 
+interactive visualization of 
+large scale datasets, \eg in \emph{imMens}~\cite{liu2013immens}.}. \agp{Clarify what is known from prior work and what
+you are proposing. Here, it is intermixed.}
+\newsaj{An additional benefit of deploying a binned overview for spreadsheets is a decrease in
+visual discontinuity during navigation. As users are able to 
+view an overview 
+that fits in the computer screen, they can 
+quickly navigate the data}---the bins act as landmarks
+in the overview, enabling 
+users to skip irrelevant bins
+and quickly navigate to the desired subset of data. We now discuss how the overview is constructed.
+
+\stitle{Overview Construction.}
+\newsaj{To generate the overview, we construct an equi-histogram. 
 Equi-depth histograms are commonly
 used for summarizing the statistical properties
-of data, with applications such as
-database query optimizers~\cite{chaudhuri1998random}. \agp{only one application?}
-The binned representation 
-of the overview
-using equi-depth histograms 
-aggregates data 
-and visualizes density 
-by counting 
-the number of data points 
-falling within each bin. \agp{This previous sentence feels repetitive.}
-Binned aggregation 
-highlights both high level (\eg densities) and low level (e.g., outliers)
-details of large datasets, 
-while enabling
-multi-granularity representation of data
-via the choice of bin size~\cite{liu2013immens}. \agp{Clarify what is known from prior work and what
-you are proposing. Here, it is intermixed.}
-Moreover, binning reduces 
-visual discontinuity during navigation
-as users are able to 
-view an overview 
-that fits in the computer screen.
-As a result, users can 
-quickly navigate the data---the bins act as landmarks
-in the overview, enabling 
-users to skip irrelevant bins
-and quickly navigate to the desired subset of data. We now discuss how the overview is constructed.}
-
-\stitle{Overview Construction.}
+of data, with applications in database systems for query optimization and approximate query processing~\cite{ioannidis2003history} and in data mining applications for distribution fitting in data streams~\cite{mousavi2011fast}, among others.} \agp{only one application?}
 \new{The equi-depth histogram 

### 2019-11-11 study.tex
addressed Aditya s5 and 6 comments

--- 2019-10-30 study.tex
@@
 \label{sec:study_design}
-\newsaj{The goal of our study is to evaluate whether \noah addresses the navigational challenges in traditional spreadsheets.
-Furthermore, we wanted understand how the presence of the features introduced by a plugin like \noah impacts a spreadsheet user's navigation experience as they explore and make sense of data.
-Therefore, we decided to compare a \noah-integrated spreadsheet with a typical, popular one, Excel, for various navigational tasks in the spreadsheet domain.} \saj{Similar domain specific-evaluations have been performed for measuring the performance of various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or
+\new{Our goal is to study the impact of an in-situ navigation plugin for spreadsheets, \noah, on navigation and exploration of data.
+Therefore, we decided to compare a \noah-integrated spreadsheet system with a typical, popular one, Excel, across various tasks.
+Similar domain specific-evaluations have been performed for evaluating various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or
 tree navigation~\cite{beard1990navigational}.
-As explained in Section~\ref{sec:related}, the goals \newsaj{and user populations} of spreadsheets and TDA tools are quite different.
+As explained in Section~\ref{sec:related}, the goals and user populations of spreadsheets and TDA tools are quite different.
 Therefore, we did not consider TDA tools for the comparative study.}
 Our study was designed to answer the following questions:
-\agp{The RQs feel repetitive. What did Karrie think.}
-\newsaj{She was fine with it.
-I think there are subtle differences.
-the RQ1 tries corresponds to quantifiable performance metric like accuracy, latency.
-RQ2 is qualitative feedback of user experience.}
 \squishlist
 \item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
-\saj{How did the addition of an overview like \noah impact \newsaj{a user's spreadsheet navigation performance  measured in terms of speed and accuracy?}}
+\new{How does integration of an overview plugin like \noah impact a) the completion time and accuracy of various spreadsheet navigational tasks\newsaj{, and b) the usability of spreadsheet systems?}} \saj{SAJ: the survey results are also part of this RQ}
 \item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
-\saj{How did \noah and it's components affect spreadsheet users'
-navigational experiences?}
+\new{How do the various components of \noah impact users' navigational experiences?}
 %\item {\bf RQ3.} \sout{How does bin customization help users inject domain knowledge in addressing their specific exploration questions compared to the automated overview?} \sout{\saj{How do users utilize additional features of \noah, \eg bin customization, during navigation?}
@@
 The two tools used in the study were: Microsoft Excel, and \noah integrated within {\scshape DataSpread}~\cite{dataspread}.
-\saj{We chose Excel for our comparative study because it is the most popular spreadsheet in use today.}
+\new{As mentioned previously, we chose Excel for our comparative study because it is the most popular spreadsheet in use today.}
 The study consisted of three phases: (a) an introductory phase explaining
@@
 to perform targeted tasks on two different datasets (described later) 
-followed by a survey to provide feedback on their impressions about both Excel and \noah, and
-(c) a semi-structured interview to collect qualitative data regarding the quiz phase.
+followed by a survey to provide feedback on their impressions about Excel and \noah, and
+(c) a semi-structured interview to collect qualitative feedback regarding the quiz phase.
 
-\stitle{Datasets.} We used two datasets---the birdstrikes (used in Keshif~\cite{yalccin2018keshif} and Voyager~\cite{wongsuphasawat2016voyager}), and the Airbnb~\cite{web:airbnb} datasets.
+\stitle{Datasets.} We used two datasets---the birdstrikes (\new{used for evaluating visual data exploration or recommendation tools like Keshif~\cite{yalccin2018keshif} and Voyager~\cite{wongsuphasawat2016voyager}}), and the Airbnb~\cite{web:airbnb} datasets.
 These datasets were chosen for their understandability to a general audience.
@@
 prospective participants filled out an interest form
-where they answered questions about their spreadsheet expertise, 
-types of spreadsheet tasks performed, and usage of spreadsheet operations. 
+where they answered questions about their spreadsheet expertise\newsaj{, spreadsheet usage purposes, and spreadsheet operations they typically use}.
 Participants were asked to rate their expertise with
@@
  \code{SUBTOTAL}, and conditional formatting. 
-To ensure that participants’ experience with spreadsheets 
-didn’t affect their performance during the quiz phase, 
+\new{To ensure that prior experience with spreadsheets 
+didn't affect the performance of participants during the quiz phase}, 
 we only recruited participants who rated their experience 
@@
 various tasks on spreadsheets, \eg maintaining, tracking, and
-analyzing, making predictions, and performing comparisons. 
+analyzing data, making predictions, and performing comparisons. 
 All of the participants were familiar with the basic mathematical
@@
 The quiz phase began as soon as the participants finished their exploration.
-\toappendix{Note that we recruited only experienced Excel users for the study and therefore, we did not provide any introduction to Excel.
-}
+\toappendix{Note that we recruited only experienced Excel users for the study and therefore, we did not provide any introduction to Excel.}
 
@@
 Category & Question & Use case    \\ \hline
-steer                & Organize the data by State. How many& \\
+steer                & Organize the data by State. How many& \code{locate} $\rightarrow$\\
                      & flights that had damages (damage = 1)& \code{summarize}\\
@@
                      & have damages? & \\
-steer                & How many flights that had damages & \code{summarize}\\
-                     & originated from California?                         & \\ 
+steer                & How many flights that had damages & \code{locate} $\rightarrow$\\
+                     & originated from California?                         & \code{summarize}\\ 
 \cmpA                & Which state between Florida and &\\
@@
 \stitle{Phase 2: The Quiz Phase.}
-\saj{The purpose of the quiz phase was to evaluate the effectiveness of the features supported by \noah in addressing spreadsheet limitations}. During the quiz phase, 
+\new{The purpose of the quiz phase was to evaluate the effectiveness of \noah in addressing spreadsheet navigation limitations}. During the quiz phase, 
 each participant performed specific tasks on the two datasets in two sessions, using Excel for one and \noah for the other. 
@@
 
-\emph{Quiz Tasks.} \saj{We designed six tasks grouped into four categories: 
-steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} uses cases \code{summarize}, \code{compare}, \code{identify}, and \code{generate}, respectively.
-All of the tasks except the customize task, are representative of the navigation interactions required for the most frequently issued spreadsheet formulae
-~\cite{bradbard2014spreadsheet, lawson2009comparison}.
-The customize tasks was designed to mimic the scenario where a user would group data subsets of interest and required participants to utilize the bin customization feature.
-The tasks were presented in the same order as shown in Table~\ref{tab:questions} for the birdstrikes dataset.
-The questions for the Airbnb dataset were similar.
-The order of the questions mimics the spreadsheet navigation scenario presented in Section~\ref{sec:usage}.}
+\emph{Quiz Tasks.} \new{We designed six tasks grouped into four categories: 
+steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} use cases \code{locate}, \code{summarize}, \code{compare}, \code{identify}, and \code{generate}, respectively.
+\newsaj{These use cases are representative of the navigation interactions required for the most frequently issued spreadsheet formulae
+~\cite{bradbard2014spreadsheet, lawson2009comparison}.} The tasks were presented in the same order as shown in Table~\ref{tab:questions} for the birdstrikes dataset.
+The tasks for the Airbnb dataset were similar.
+The order of the tasks mimics the spreadsheet navigation scenario presented in Section~\ref{sec:usage} \newsaj{where the task complexity increases as the user continues to navigate and analyze the data.}}
+%The customize tasks was designed to mimic the scenario where a user would group data subsets of interest and required participants to utilize the bin customization feature. 
 %All of these tasks require navigation of the spreadsheet data in the form of either scrolling or steering.
 
-\saj{The steer tasks required participants to use the aggregate column feature in \noah as opposed to steering in Excel.
-The identify task required participants to interact with the detailed view and relate the aggregate result of the first steer task with the raw spreadsheet data.
-The \cmpA task asked the participants to compare the results of the first steer task (\eg Florida) with that of the second steer task (\eg California) which would require them to use the context bar to revisit a previously visited bin.
-The \cmpB task involved comparing statistics of all the data subsets.
+\newsaj{The steer tasks required the participants to first \code{locate} a data subset and then \code{summarize} the subset by issuing an aggregate operation, \eg \code{COUNTIF}.
+Following summarization, the identify task required participants to interact with the detailed view and \code{identify} data points within the spreadsheet that satisfy the condition of the first steer task.
+The purpose of the second steer tasks was to enable the \cmpA task---given the two summaries obtained from the two steer tasks, the \cmpA task asked the participants to \code{compare} the results of both.
+Then the participants were asked to perform an even more complex comparison, \ie the \cmpB task, which involved comparing the same statistics computed in the two steer tasks of all the data subsets.
 The purpose of the \cmpB task was to see how increasing the number of subsets would affect navigation.
-As explained earlier, the customize task was designed to evaluate the utility of bin customization feature.}
+Finally, the customize task was designed to mimic the scenario where the \noah-generated overview doesn't reflect the user preference and a user would group data subsets of interest to create a new overview.
+(accomplished by utilizing the bin customization feature in \noah).}
+
+%\newsaj{The steer tasks required participants to first \code{locate} a data subset and then issue a formula, \eg \code{COUNTIF}, via steering in Excel to \code{summarize} the subset (accomplished by the aggregate column feature in \noah).
+Following summarization, the identify task required participants to interact with the detailed view and \code{identify} data points within the spreadsheet that satisfy the condition of the first steer task (accomplished by inspecting the colored spreadsheet cells in \noah).
+The purpose of the second steer tasks was to enable the subsequent comparison tasks.
+Given the two summaries obtained from the two steer tasks, the participants were then asked to \code{compare} the results of both, \ie the \cmpA task (accomplished by utilizing the context bar feature of \noah to revisit the results of the first steer task).
+The \cmpB task involved comparing the same statistics computed in the two steer tasks of all the data subsets.
+The purpose of the \cmpB task was to see how increasing the number of subsets would affect navigation.
+Finally, the customize task was designed to mimic the scenario where the \noah-generated overview doesn't reflect the user preference and a user would group data subsets of interest to create a new overview.
+(accomplished by utilizing the bin customization feature in \noah).}
 
@@
 and negative aspects of both tools. 
-The survey was designed to evaluate \emph{RQ4}.
+\newsaj{The survey was designed to evaluate \emph{RQ1b}}.
 
 \emph{Evaluation.} 
-We evaluated the accuracy (either 0 or 1) and
+We evaluated the accuracy and
 completion time for each of the six tasks. 
-\saj{We combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be explained with multiple information sources.}
-For example, 
-we analyzed the video recordings of participants' interaction with the tools during the quiz phase. 
-We further analyzed the survey responses to quantify the usability of both the tools.
+\newsaj{We combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be corroborated across multiple information sources.
+Moreover, we analyzed the survey responses to quantify the usability of both Excel and \noah-integrated {\scshape DataSpread}.}
+%For example, we analyzed the video recordings of participants' interaction with the tools during the quiz phase.
 
@@
 \subsection{Study Limitations.}
-\saj{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies.
-One limitation arises from our participant demographics.
-We conducted the study with $20$ participants from four different backgrounds.
-Our participant pool demographics partially represent the demographics of the general audience intended for \noah.
-%is in sample size and participant demographics: we conducted the study with $20$ participants from four different backgrounds.Therefore, our participant pool demographics partially represents the demographics of the general audience intended for \noah.
-A larger sample with more participants that better represented the spreadsheet user population would have provided more ecological validity to generalize our findings.
+\newsaj{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies.
+Firstly, our participant pool demographics partially represents the demographics of the general audience intended for \noah.
+A larger sample with more participants that better represents the spreadsheet user population would have provided more ecological validity to generalize our findings.
 }
@@
 
-\saj{Another limitation of the study is that we only compared the performance of a \noah-integrated spreadsheet with a traditional spreadsheet.
-We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} due to their limitations in the types of interactions supported (see Section~\ref{sec:related}).
-Moreover, instead of forcing the participants use unfamiliar spreadsheet operations, we wanted them to utilize the typical spreadsheet operations that they are comfortable with, to complete the tasks and see how introduction of \noah affected their navigation experience.}
+\newsaj{Next, we only compared the performance of a \noah-integrated spreadsheet with a traditional spreadsheet.
+We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} as they violate most of the design considerations proposed in Section~\ref{sec:design}.
+We discussed their limitations in Section~\ref{sec:related}.
+Moreover, allowing the participants to utilize the typical spreadsheet operations that they were comfortable with, enabled us to observe how introduction of \noah affected their navigation experience.}
 
-\saj{Finally, we did not isolate the effects of the individual features of \noah to better understand the implications of those.
+\new{Finally, we did not isolate the effects of the individual features of \noah to better understand the implications of those.
 For example, the effect of binned overview (visual clarity versus visual continuity), display layout (screen space trade-off), and
@@
 However, the goal of the study was to understand participants' navigation experience in the presence and absence of \noah.
-A more fine-grained study can be conducted in the future, to inspect the contribution of individual components of \noah in further detail.}
+A more fine-grained study that teases apart the contribution of individual components of \noah is warranted.}
 

### 2019-11-25 appendix.tex
final tvcg

@@
-In this supplementary material we highlight the use cases that \noah
-is suited for, the in-situ chart representations within
+%!TEX root = main.tex
+
+In this section we present the in-situ chart representations within
 the overview, \noah’s underlying system architecture,
 and statistical significance results from our user study.
-\section*{\noah Use Cases}
-As mentioned in Section~\ref{sec:design}, the goal of designing 
-an overview interface is to 
-not only enable seamless navigation 
-but also to facilitate functionalities to help users visually seek information about the data.
-Therefore, we draw parallels to the typology of abstract visualization tasks~\cite{brehmer2013multi} to identify the use cases that \noah is best suited for.
-We are specifically interested in exploring the different task purposes mentioned in the typology.
-The task purposes can be specified at three different levels: consume and produce (high-level), search (mid-level), and query (low-level).
-Table~\ref{tab:scope} captures all three levels and highlights where \noah can be used.
-
-At a very high level, users may utilize \noah to verify some phenomenon in the data (\code{discover}) or consume the data for casual interest (\code{enjoy}).
-Example of \code{discover} includes a journalist trying to verify whether listings in Airbnb are run as hotels (see Section~\ref{sec:usage}).
-Moreover, users may utilize the overview to communicate details about the data (\code{present}). 
-However, visualization tools are more suitable for such a purpose. 
-Finally, users can customize (\code{create}) the overview to fit their exploration purpose. Here, the high-level goal is to \emph{Produce} a new representation. 
-However, the present version of \noah does not allow users to add annotations, \eg visual cues, texts, to the overview (\code{annotate}).
-
-Regardless of the high-level purposes, users may want to find elements of interest in the overview (\emph{search}) and seek additional details about those elements (\emph{query}). 
-In Section~\ref{sec:ui}, we discussed the features and interactions supported by \noah and showed how these interactions enable users to \emph{search} and \emph{query} the data.
-Users can \code{browse} and \code{explore} the data using navigational operations, \eg clicking and zooming, to find elements of interest.
-Note that if users know either the identity or location of the information they are searching (\eg the name of the city in the Airbnb data), 
-they can simply use traditional spreadsheet operations like \emph{VLOOKUP} (\code{lookup}) or \emph{CTRL+F} (\code{locate}) (see Table~\ref{tab:scope}). 
-On the other hand, clicking or zooming is suitable when the identity or location of the information is unknown.
-Therefore, NOAH complements spreadsheets in accomplishing users’ \emph{search} goal via navigational operations.
-Once the user finds the elements of interest, they can further \code{summarize} the elements via aggregate columns, \code{compare} those elements, and \code{identify} the raw data corresponding to the summaries in the spreadsheet.
-
 \section*{Explaining the Chart Representations}
@@
 The bar represents
-the range of the data subset the bin spans
+the range of the data subset the bin spans,
 with the minimum and maximum values
@@
 
-\stitle{Conditional.} The result of a \emph{conditional} formulae,
+\stitle{Conditional.} The result of a \emph{conditional} formula,
 \eg \code{COUNTIF}, is depicted using a histogram.
-The histogram captures the distribution of the attribute,
+The histogram captures the distribution of the attribute
 on which the formula has been applied,
@@
 \eg \code{mode}, is depicted using a histogram.
-A similar shading technique as the \emph{conditional} formula is used to highlight the standard deviation. The mean of the is highlighted using a vertical line.
+A similar shading technique as the \emph{conditional} formula is used to highlight the standard deviation. The mean is highlighted using a vertical line.
 
@@
 The darker the color, the higher the value corresponding to that entry. 
-Users can utilize the color intensity
-to compare the results among different aggregate column entries.
+Users can use the color intensity
+to compare results among different aggregate column entries.
 
@@
 In this section, we provide an overview of the infrastructure of \noah. 
-We integrate \noah as a data exploration plugin within{\scshape DataSpread}~\cite{dataspread}, 
+We integrate \noah as a data exploration plugin within {\scshape DataSpread}~\cite{dataspread}, 
 an open-source scalable web-based spreadsheet.
 
-\subsection{Underlying Data Structures}
+\subsection*{Underlying Data Structures}
 As explained in Section~\ref{sec:ui}, the underlying data structure representing the overview is an in-memory equi-depth histogram. 
-\noah constructs the histogram on demand based on the navigation attribute. 
+\noah constructs the histogram on-demand based on the navigation attribute. 
 In the beginning, only the highest granularity bins are constructed. 
@@
 order statistics tree~\cite{datamodels} 
-built on the position (\eg row number) of the spreadsheet data. 
-For any given navigation attribute, a new positional index is constructed first. 
+built on the position (\eg row number) of the spreadsheet data. For any given navigation attribute, a new positional index is constructed first. 
 \noah then leverages the positional mapping to access the underlying data corresponding to the navigation attribute and constructs the histogram depicting the overview. 
@@
     \includegraphics[trim=0 4cm 0 3cm,clip,width=\linewidth]{images/arch.pdf}
-   \caption{System architecture.}
+   \caption{\noah System architecture.}
    \label{fig:arch}
@@
 We now explain the system architecture of \noah. 
-The \noah client is a web-based front-end that captures user input and renders both the navigation interface and the spreadsheet 
+The \noah client is a web-based front-end that captures user input and renders both the navigation plug-in and the spreadsheet 
 based on the results returned by the back-end. 
-The front-end is responsible for capturing user input and rendering components of the navigation interface, \ie overview, aggregate column, and context bar. 
+The front-end is responsible for capturing user input and rendering components of the navigation interface, \ie the overview, aggregate column, and context bar. 
 Given any interaction by the user on the front end, 
-the \emph{request processor} issues a request to back end.
-The back end navigation controller receives the request from the front end. 
+the \emph{request processor} issues a request to back-end.
+The back end navigation controller receives the request from the front-end. 
 After processing the request that corresponds to some front-end user interaction, 
@@
     %\vspace{-10pt}
-    \includegraphics[trim=0 57 0 145,clip,width=\linewidth]{images/userinfo-ip-1.pdf}
+    \includegraphics[trim=0 50 0 145,clip,width=\linewidth]{images/userinfo-ip-1.pdf}
    \caption{Intra-participant submission time differences between \noah and Excel across the quiz tasks.}
@@
 
-\subsection{Intra-participant differences.}
-Figure~\ref{fig:ipdiff} depicts the intra-participant submission time differences between \noah and Excel, across all the quiz tasks.
-For a given task and a participant, the corresponding circle denotes by what percentage (between $0.30\%$ to $94.33\%$) a system is faster than the competing system.
-The color of the circle denotes which system was faster (green: \noah faster, yellow: Excel faster).
-The larger the circle the faster the corresponding system is.
+\subsection*{Intra-participant differences.}
+\new{Figure~\ref{fig:ipdiff} depicts the intra-participant submission time differences between \noah and Excel, across all the quiz tasks.
+For a given task and a participant, the corresponding circle denotes by what percentage (between $0.30\%$ to $94.33\%$) a participant on a given tool was faster than on the other tool, when submitting a response.
+The color of the circle denotes on which tool the participant's submission time was faster (green: \noah faster, yellow: Excel faster).
+The larger the circle, the faster the corresponding tool is.
 For example, participant $P9$'s submission time was $91.78\%$ faster with \noah for the \cmpB task and $72.62\%$ faster with Excel for the \cmpA task.
 Figure~\ref{fig:ipdiff} confirms that majority of the submission times using \noah were faster compared to Excel---nineteen out of the twenty participants completed at least four tasks in less time using \noah compared to Excel.
-The submission time difference is even more apparent for the steer, identify, and \cmpB tasks.
+The submission time difference is even more apparent for the steer, find, and \cmpB tasks.}
 
-For \cmpA, customize, and the second steer task, less than a third of the participants' submission times were faster using Excel. 
-For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah, utilized the \emph{autosum} shortcut for the birdstrikes dataset.
-For the \cmpA tasks, a number of participants ($N=3$) that required more time to submit answer using \noah, first used the bin customization feature to view all the unique bins which contributed to the higher submission time compared to Excel.
-On the other hand, as explained earlier, the unfamiliarity with the bin customization feature contributed to higher submission time using \noah for six participants (yellow circles).
+\new{For \cmpA, customize, and the second steer task, fewer than a third of the participants' submission times were faster using Excel. 
+For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah used the \emph{autosum} shortcut for the birdstrikes dataset.
+For the \cmpA tasks, a number of participants ($N=3$) that required more time to submit an answer using \noah, first used the bin customization feature to view all the unique bins,  contributing to a higher submission time compared to Excel.
+On the other hand, as explained earlier, the unfamiliarity with  bin customization contributed to higher submission time using \noah for six participants (yellow circles).}
 
-\subsection{Statistical Significance: Task Performance}
-Since we conducted our user study on a small population (20 participants), we further evaluated the statistical significance of the task performance results, \ie accuracy and completion time. 
+\subsection*{Statistical Significance: Task Performance}
+Since we conducted our user study on a relatively small population (20 participants), we further evaluated the statistical significance of the task performance results, \ie accuracy and completion time.
 To measure the significance of the task completion times, we ran \emph{Mann-Whitney's U test} (as completion times did not follow a normal distribution).
 \emph{For all of the tasks except the customize task, we found a significant effect of the tools, \ie the response times for the tasks significantly differed by the choice of the tool} (see Table~\ref{tab:statTest}).
-On the other hand, we ran the \emph{Fisher's exact test} that measures the statistical significance of categorical data (0/1 accuracy): \emph{only for steer tasks, the percentage of the accuracy of submissions significantly differed by the choice of the tool}.
+We ran the \emph{Fisher's exact test} that measures the statistical significance of categorical data to asses the difference in accuracies (0/1 accuracy) : \emph{the percentage of accurate of submissions significantly differed by the choice of the tool only for the steer tasks}.
 
@@
 & & ($p$ value) & ($p$ value) \\ \hline
-$Q1$ & Steer & $\mathbf{0.0007}$ (*) &  $\mathbf{0.0033}$ (*)  \\
-$Q2$ & Identify &   $\mathbf{2.49 \times 10^{-5}}$ (*) & $0.7475$ \\  
-$Q3$ & Steer & $\mathbf{0.0043}$ (*) & $\mathbf{0.0202}$ (*) \\
+$Q1$ & steer & $\mathbf{0.0007}$ (*) &  $\mathbf{0.0033}$ (*)  \\
+$Q2$ & find &   $\mathbf{2.49 \times 10^{-5}}$ (*) & $0.7475$ \\  
+$Q3$ & steer & $\mathbf{0.0043}$ (*) & $\mathbf{0.0202}$ (*) \\
 $Q4$  & \cmpA & $\mathbf{0.0154}$ (*) & $1$ \\  
 $Q5$ & \cmpB & $\mathbf{5.83 \times 10^{-6}}$ (*) & $0.48$\\
-$Q6$ & Customize & $0.1207$ &    $0.0959$ \\\hline
+$Q6$ & customize & $0.1207$ &    $0.0959$ \\\hline
 \end{tabular}
@@
 
-\subsection{Statistical Significance: Subjective Ratings}
-We further conducted a statistical significance test---the \emph{Wilcoxon Signed-rank} test---on the survey responses which showed that for all the metrics, the ratings significantly differed by the choice of the tool, \ie \noah or Excel.
+\subsection*{Statistical Significance: Subjective Ratings}
+We further conducted a statistical significance test---the \emph{Wilcoxon Signed-rank} test---on the survey responses, with results shown in Table~\ref{tab:survey}.
+For all the metrics, the ratings significantly differed by the choice of the tool, \ie \noah or Excel.
 The distribution of the ratings for none of the criteria followed a normal distribution.

### 2019-11-25 design.tex
final tvcg

@@
 
-\newsaj{The data structure underlying the overview
+\new{The data structure underlying the overview
 is a histogram constructed
@@
 in the navigation attribute column.
-Histograms are binned aggregates
-where each bin represents a collection (group) of data points and the frequency of the data points, \ie \code{COUNT} aggregate, visualizes the density of the corresponding group.} \agp{This previous sentence feels repetitive.} \saj{rephrased.
-wanted to explain that constructing histogram is a type of binned aggregation.}
-\newsaj{In addition to providing high level (\eg densities) and low level (e.g., outliers) details, binned aggregation techniques 
-enables multi-granularity visual representation of data
+Histograms result from {\em binned aggregation}---consecutive data points are grouped into bins (or groups),
+where each bin represents a collection (group) 
+of data points and is associated with a count aggregate,
+capturing the number of data points that fall in that group.
+In addition to providing high level (\eg densities) 
+and low level (e.g., outliers) details, 
+binned aggregation techniques 
+enable a multi-granularity visual representation of data
 by varying the bin size and
-have been deployed in 
+have therefore been deployed in 
 interactive visualization of 
-large scale datasets, \eg in \emph{imMens}~\cite{liu2013immens}.}. \agp{Clarify what is known from prior work and what
-you are proposing. Here, it is intermixed.}
-\newsaj{An additional benefit of deploying a binned overview for spreadsheets is a decrease in
+large scale datasets, \eg in \emph{imMens}~\cite{liu2013immens}.
+An additional benefit of a binned overview for spreadsheets 
+is a decrease in
 visual discontinuity during navigation. As users are able to 
@@
 \stitle{Overview Construction.}
-\newsaj{To generate the overview, we construct an equi-histogram. 
+\new{To generate the overview, we construct an equi-histogram. 
 Equi-depth histograms are commonly
-used for summarizing the statistical properties
-of data, with applications in database systems for query optimization and approximate query processing~\cite{ioannidis2003history} and in data mining applications for distribution fitting in data streams~\cite{mousavi2011fast}, among others.} \agp{only one application?}
-\new{The equi-depth histogram 
+used for summarizing statistical properties
+of data, with applications in database systems 
+for query optimization and approximate 
+query processing~\cite{ioannidis2003history} 
+and in data mining applications for distribution fitting in data streams~\cite{mousavi2011fast}, among others.
+The equi-depth histogram 
 can be constructed on 
@@
 that the ranges represented by each bin have no overlap.
-\agp{Aditya: would be valuable to describe the algorithm to increase
-technical depth.}
-\saj{Saj: can we add a pseudocode in appendix and refer to that?}
-\agp{Yes please.}
+% \agp{Aditya: would be valuable to describe the algorithm to increase
+% technical depth.}
+% \saj{Saj: can we add a pseudocode in appendix and refer to that?}
+% \agp{Yes please.}
 
@@
 may be reflected on the spreadsheet and vice-versa.
-\new{One example of this coordinations is} indicating the navigation attribute on the spreadsheet
+\new{One example of this coordination is} indicating the navigation attribute on the spreadsheet
 using color (see the lime green column in Figure~\ref{fig:ux}c)
@@
 \subsection{Aggregate Columns}
+\label{sec:agg_col}
 Users can issue spreadsheet formulae
@@
 We have integrated \noah with {\scshape DataSpread}~\cite{datamodels}, a web-based spreadsheet.
-The {\scshape DataSpread} back-end maintains the histogram data structure and  supports the aggregate column computation via its built-in formula engine.\toappendix{ To enable seamless integration of the overview with the spreadsheet data, we leverage the hierarchical positional indexes used by {\scshape DataSpread} to access
-the spreadsheet data.
-The index is essentially an order statistics tree~\cite{datamodels} built on the position (\eg row number) of the spreadsheet data.
-For any given navigation attribute, a new positional index is constructed first.
-\noah then leverages the positional mapping to access the underlying data corresponding to the navigation attribute and constructs the histogram depicting the overview.
-Therefore, \noah can be integrated to any spreadsheet and requires only access to the positional mapping structure of that spreadsheet.} The \noah front-end is built with HTML/CSS/JS technologies along with the D3 framework~\cite{d3} for generating charts.
+The {\scshape DataSpread} back-end maintains the histogram data structure and  supports the aggregate column computation via its built-in formula engine.
+The \noah front-end is built with HTML/CSS/JS technologies along with the D3 framework~\cite{d3} for generating charts. We explain the \noah architecture in detail in the Appendix.

### 2019-11-25 discussion.tex
final tvcg

@@
+%!TEX root = main.tex
 \section{Discussion and Conclusion}
 \label{sec:discuss}
-\saj{In this paper, we have introduced \noah, an in-situ navigation interface, designed as a spreadsheet plugin.
-Using \noah, users can get a bird’s-eye view of the data, with the ability to scroll or seek
-additional details on demand, using a multi-granularity overview, as well as
-aggregate columns that eliminate cumbersome steering operations.
-Quantitatively, we find \noah to speed up navigation without compromising the accuracy of the tasks.
+\new{\noah represents our 
+first step towards a general purpose spreadsheet navigation plug-in to make spreadsheets more effective
+when exploring datasets that are increasingly the norm. 
+Using \noah, users can get a bird's eye view of the data, with the ability to scroll or seek
+additional details on demand via a multi-granularity overview, as well as employ
+aggregattion in-situ, which eliminates cumbersome steering operations.
+Quantitatively, we find that \noah speeds up navigation without compromising accuracy.
 Qualitatively, study participants identify it as positively impacting their experience while overviewing and navigating large datasets, and issuing formulae.
-However, participants identified unfamiliarity and lack of transparency of operations as some of the limitations of \noah. In this section, we discuss
-how to address these limitations
-and highlight other future enhancement opportunities for \noah.}
+However, the user study revealed some limitations of \noah.
+Moreover, the study design itself had some limitations.
+In this section, we discuss these limitations and propose extensions to address these limitations as well as enhance the capabilities of \noah.}
 
-\stitle{Transparency and documentation.} \saj{Several operations in \noah 
-are quite different from typical spreadsheet interactions, \eg zooming, bin customization.
-Moreover, participants found some of the terminologies quite different from typical spreadsheet terminologies (see Section~\ref{sec:rq2}).
+\subsection{User Study Limitations}
+Our study has a few limitations that can be strengthened by future larger-scale and finer-grained studies.
+
+\stitle{\new{Insufficient coverage of participant demographics.}} \new{Our participant pool demographics only partially represents the demographics of the general audience intended for \noah.
+A larger sample with more  participants with a range of skill-sets and backgrounds that better represents the spreadsheet user population would have provided more ecological validity to generalize our findings.}
+%While a larger sample with more diverse backgrounds would have allowed us to perform more definitive quantitative analysis, we combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be explained with multiple information sources.
+
+\stitle{\new{Lack of targeted comparisons with advanced spreadsheet features.}} \new{We only compared the performance of a \noah-integrated spreadsheet with a traditional spreadsheet.
+We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} as they violate most of the design considerations proposed in Section~\ref{sec:design}; we discussed their limitations in Section~\ref{sec:related}.
+We instead allowed the participants to freely use any spreadsheet operations that they were comfortable with (including the advanced ones), enabling us to observe how introducing a plug-in like \noah affected their navigation experience.
+However, a future study targeted at evaluating the pros and cons of these features for spreadsheet navigation would be valuable.}
+
+\stitle{\new{Isolated evaluation of \noah components.}} \new{While we did present the impact of various components of \noah in Section~\ref{sec:rq2}, we did not isolate the effects of individual features during our study.
+For example, we did not study the effects of the binned overview (visual clarity versus visual continuity), display layout (screen space trade-off), and
+contextual presentation of data (raw text versus chart representation of aggregate columns) in isolation.
+A more fine-grained study that teases apart the contribution of individual components of \noah is warranted.}
+
+\subsection{Limitations of \noah and Possible Enhancements}
+While we alluded to some limitations of \noah in Section~\ref{sec:results}, we now discuss these in more detail and how we can possibly overcome them.
+
+\stitle{\new{Lack of transparency and documentation of new interactions.}} \new{Several operations in \noah 
+are quite different from typical spreadsheet interactions, \eg zooming or bin customization.
+Moreover, participants found some of the terminologies, \eg explore or bin, to be rather different from typical spreadsheet terminologies, which took some getting used to (see Section~\ref{sec:rq2}).
 In addition, some participants complained about the lack of explanation surrounding the overview construction and aggregate column computation.
@@
 
-\stitle{Maintaining spreadsheet look and feel.}
-\saj{In subsequent versions of \noah, we can further 
+\stitle{\new{Deviation from the spreadsheet look and feel.}}
+\new{A couple of participants ($P2$ and $P11$) mentioned that the aggregate column results hides the actual spreadsheet formula and they would prefer some visual cues that highlight the corresponding formula underlying each bin.
+Subsequent versions of \noah can
 display the appropriate formula for each bin as users hover
-over the corresponding cell on the aggregate column. 
-Moreover, participants ($N=5$) noted the fact that \noah currently 
-does not support user defined formulae, another possible future enhancement.
-Bin customization in \noah is performed using a menu bar which adds an additional step.
-In Excel, the cell splitting and merging operations are direct and only requires a single click.
-Similar direct adjustment of data grouping strategies have been explored for visualization tools~\cite{sarvghad2018embedded} and can be introduced in a future version of \noah.}
+over the corresponding cell on the aggregate column\cut{making the output more relatable to typical spreadsheet formula results}.
+The aggregate column feature can be further enhanced by enabling users to issue user defined formulae, a feature requested by a number of participants ($N=5$).
+Another feature that can be made more similar to spreadsheet interaction semantics is the bin customization operation.
+Currently, this operation in \noah is performed from a menu bar, adding an additional step.
+The bin splitting and merging operations can be made more similar to how spreadsheet cells are split or merged---in Excel, these operations are direct and only require a single click.
+Similar direct adjustment of data grouping strategies have been explored for visualization tools~\cite{sarvghad2018embedded} and can be adapted to this setting.}
 
-\stitle{Binning for different data types.} \saj{The experience 
+\stitle{\new{Absence of bespoke overview representations for various data types.}} \new{The experience 
 surrounding the construction
 of the overview
-can be further improved, specially for categorical data.
+can be further improved, especially for categorical data.
 Currently, the bins of the overview
@@
 Providing the users
-the capability to
+the ability to
 select the representation
 (similar to bin customization)
-of the overview could have addressed this issue.
+of the overview at the outset could have possibly addressed this issue.
 Understanding the impact of these
 representation choices for the overview
-can be an interesting future research.}
+is an interesting open question.}
 
-\stitle{Scope of overview-spreadsheet coordination.} 
-\saj{Spreadsheet users may perform various edit operations, 
+\subsection{Additional enhancements}
+We now discuss other functionalities that can be introduced to further enhance the capabilities of \noah.
+
+\stitle{\new{Broadening the scope of overview-spreadsheet coordination.}} 
+\new{Spreadsheet users often perform various edit operations, 
 \eg updating values, adding/deleting rows/columns. 
-However, \noah currently assumes the data to be read-only. 
+However, our current \noah implementation assumes the data to be read-only. 
 In our next version, we can add support for
-propagating the spreadsheet updates to the overview. 
+propagating spreadsheet updates to the overview. 
 Moreover, the charts displayed in an \emph{aggregate column} are non-interactive, 
@@
 
-\stitle{Enhancing the navigation experience.} \noah currently constructs the overview on a single attribute.
+\stitle{\new{Enabling more flexible overview binning.}} \noah currently constructs the overview on a single attribute.
 We can add support for multi-attribute navigation (\eg explore the Airbnb data by city and neighborhood), and multi-level navigation (\eg explore the neighborhoods after zooming into a specific city in the Airbnb data).
-Furthermore, bin customization currently supports changing the bin boundaries only while maintaining the current order.
-Supporting user defined ordering to allow arbitrary reshuffling of the bins can be another enhancement.\toappendix{ However, such modification will require a redesign of the histogram based data structure since the ordering property will not hold anymore.Finally, a complete characterization of the spreadsheet operations that can be supported by \noah is another open question, for example, other operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting, can also be supported by \noah.}
+Furthermore, bin customization currently supports changing the bin boundaries only, while maintaining the current order.
+Supporting user-defined ordering to allow arbitrary reshuffling of the bins can be another enhancement.\toappendix{ However, such modification will require a redesign of the histogram based data structure since the ordering property will not hold anymore.Finally, a complete characterization of the spreadsheet operations that can be supported by \noah is another open question, for example, other operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting, can also be supported by \noah.}
 
-\stitle{Expanding the scope of supported operations.} \saj{Spreadsheet operations that involve working with subsets of spreadsheet data, \eg sorting, filtering, copy-pasting, can also be supported by \noah.
-Therefore, it is necessary to  completely characterize the scope of spreadsheet operations that can be supported by \noah.
-Other enhancement includes adding annotations, \eg visual cues, texts, to the overview and then exporting the customized overview---a required feature for information seeking tools~\cite{shneiderman2003eyes}.}
+\stitle{\new{Supporting other spreadsheet operations.}} \new{Spreadsheet operations that involve working with subsets of the data, \eg sorting, filtering, copy-pasting, can be supported by \noah, but our current implementation does not support them.
+Future versions would support such operations as well.
+Other enhancements include adding annotations, \eg visual cues, text, to the overview and then exporting the customized overview for presentation or reporting---a required feature for information seeking tools~\cite{shneiderman2003eyes}.}
 
-\stitle{Beyond Tabular Data.} \noah operates only on tabular data.
+\stitle{\new{Supporting navigation for semi-structured data.}} \noah operates only on tabular data.
 However, spreadsheets can be semi-structured---formulae 
 and text can be interspersed with tabular data. 
-We can extend \noah to act as a map highlighting heterogeneous regions on such complex spreadsheets which users can utilize to navigate the spreadsheet. 
+\new{In such cases, NOAH can be used to support exploration and zooming for each such tabular region independently, supported by an overall overview (akin to a map-based panning tool) for users to select which tabular region they want to explore in detail.}
 We can leverage existing work on spreadsheet table detection~\cite{tablesense}, 
 and property identification~\cite{chen2017spreadsheet} 
-to construct the map.
+to \new{support such extensions}.
 
-\stitle{From Perceptual to Interactive Scalability.} 
+\stitle{\new{Achieving interactive scalability beyond traditional spreadsheet limits.}} 
 The current version of \noah addresses the perceptual 
@@
 This opens the door to a new set of research challenges that may range from approximate query processing to progressive data analytics.
-
-\noah represents our 
-first step towards a general purpose spreadsheet navigation plug-in that can make spreadsheets more effective
-on large datasets that are increasingly the norm.

### 2019-11-25 introduction.tex
final tvcg

@@
 \stitle{\noah: \new{a navigation plug-in for spreadsheets}.} We address the aforementioned challenges
-in \noah,
+in \noah\footnote{\new{A video demo of \noah can be found at https://www.youtube.com/watch?v=iZsboe3x680}},
 an in-situ

### 2019-11-25 letter.tex
final tvcg

@@
 
-A previous version of this paper was submitted to InfoVis 2019.
+A previous version of this paper was submitted to InfoVis 2019 (submission ID: 1043).
 This version was rejected as the paper co-chairs deemed that revising the paper would require more time than allocated to meet reviewer expectations.
 The paper co-chairs mentioned the following: ``\textit{We particularly encourage such revisions where submissions were positively received by reviewers, but the revisions required were deemed to be beyond the scope of the conference review cycle.
-If you address the issues raised and subsequently submit to TVCG, please make reference to this  InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}'' The entire set of reviews are attached.
-In particular, the reviewers asked for the following changes for a revision:
+If you address the issues raised and subsequently submit to TVCG, please make reference to this InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}'' The entire set of reviews and the old INFOVIS submission are attached  at the end of this document.
+The reviewers asked for the following changes for a revision:
 \begin{itemize}
@@
 stronger as a result. 
-Every section in the paper has been revised;
-the changes are highlighted in \new{blue}.
+Every section in the paper has been revised. We have attached the new version of the paper with this ``summary of changes'' document, where
+the changes are highlighted in \new{blue} for reviewers' convenience.
 We briefly describe the changes to each section
@@
 and explain why 
-we opted for histograms as an overview representation, \new{by contextualizing our
-approach using prior work on multi-scale aggregation~\cite{elmqvist2009hierarchical}.} 
+we opted for histograms as an overview representation, by contextualizing our
+approach using prior work on multi-scale aggregation~\cite{elmqvist2009hierarchical}. 
 \item 
@@
 we now explain our choice
-for the quiz tasks and discuss 
-three limitations of our study in detail. 
+for the quiz tasks in detail. 
 \item Section~\ref{sec:results} has been completely
@@
 include additional analysis results on intra-participant differences. 
-\item In Section~\ref{sec:discuss}, we have added a summary of our takeaways from the evaluation study, while discussing additional limitations of \noah and future enhancement opportunities. 
+\item In Section~\ref{sec:discuss}, we have added a summary of our takeaways from the evaluation study, while discussing limitations of our evaluation study and \noah's design while highlighting future enhancement opportunities.
 \end{itemize}
@@
 is its \emph{design as a general-purpose navigation plug-in} 
-to any existing spreadsheet system. 
+to any existing spreadsheet tool. 
 As spreadsheets have a massive user-base, 
@@
 we have provided a thorough justification
-for what spreadsheet systems like Excel offer
+for what spreadsheet tools like Excel offer
 relative to tabular data analysis tools (TDA),
@@
 \item \textit{Reporting Intra-participant Differences} (\textbf{C6}):
-\newsaj{We performed additional analysis of the study data to report intra-participant submission time differences for the quiz phase tasks.
-While we summarize the results of the analysis in Section~\ref{sec:time_acc}, the detailed result can be found in Appendix}\agp{pls add; insights on trends sounds vague let's be more clear}.
-\newsaj{We find that, despite a few exceptions, across all tasks, participants' task submission times were faster using \noah---all but one participants completed at least four tasks in less time using \noah compared to Excel.} However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}.
-Therefore, we have included the results of the intra-participant analysis in Appendix.
+We performed additional analysis of the study data to report intra-participant submission time differences for the quiz phase tasks.
+While we summarize the results of the analysis in Section~\ref{sec:time_acc}, the detailed results can be found in Appendix.
+We find that despite a few exceptions, across all tasks, participants' task submission times were faster using \noah---all but one participants completed at least four tasks in less time using \noah compared to Excel.
+However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}.
+Therefore, we decided to include the results of the intra-participant analysis in the Appendix as opposed to the main paper.
 \end{enumerate}
+
+We realize that these extensive changes throughout the paper have increased the paper length.
+As the TVCG submission guidelines indicate ``\emph{The regular paper page length limit is defined at 12 formatted
+pages, including references and author biographies. Any pages or
+fraction thereof exceeding this limit are charged \textdollar220 per page. Regular
+papers may not exceed 18 formatted pages}'', so we have decided to let the paper length stand since our paper is at 16 pages currently — and we would be happy to pay for the extra pages.
+We would also be happy to reduce length by moving appropriate content from the main body to the supplementary material as suggested by reviewers.
+The following are attached with the ``Summary of changes'' document:
+\begin{itemize}
+    \item The resubmitted TVCG paper with changes highlighted in \new{blue}.
+    \item The entire set of reviews of the INFOVIS submission.
+    \item The previous INFOVIS submission.
+\end{itemize}

### 2019-11-25 results.tex
final tvcg

@@
+%!TEX root = main.tex
+
 \section{Results}
@@
 
-\subsection{RQ1. Navigation performance of participants}
+\subsection{RQ1. Impact of overview-spreadsheet integration on navigation performance and spreadsheet usability}
 \label{sec:rq1}
-To answer RQ1, we first compare our participants’ task completion times  
-and accuracies in \noah and Excel while analyzing the survey response.  
+To answer RQ1, we first compare task completion times  
+and accuracies in \noah and Excel and \new{then analyze the survey responses that evaluate the usability of the tools.}  
 \begin{figure}[t]
@@
 \begin{tabular}{c c}  %trim=left bottom right top
- \includegraphics[width=0.23\textwidth,trim={18 37 20 15},clip]{images/bird_box.pdf} &
-   \includegraphics[width=0.23\textwidth,trim={18 37 20 15},clip]{images/airbnb_box.pdf} \\
+ \includegraphics[width=0.23\textwidth,trim={18 20 20 15},clip]{images/bird_box.pdf} &
+   \includegraphics[width=0.23\textwidth,trim={18 20 20 15},clip]{images/airbnb_box.pdf} \\
    \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
@@
 we show the distribution of submission times of participants 
-for the four task categories, for birdstrikes and Airbnb respectively. 
+for the five task categories, for birdstrikes and Airbnb respectively. 
 For most categories, 
 \emph{participants' median submission times using \noah 
-were less than the fastest submission times using Excel}. This suggests that
+were less than the fastest submission times using Excel}. This observation suggests that
 the new capabilities offered by \noah
 made spreadsheet navigation faster for these tasks.
-\saj{We also analyzed the intra-participant submission time difference, and that supported the aforementioned observations.
-Overall, majority of the submission times using \noah were faster compared to Excel---19 out of the 20 participants completed at least four tasks in less time using \noah compared to Excel.
-The submission time differences were more prominent for the steer, identify, and \cmpB tasks.
-Moreover, for all tasks except customize, the difference in submission times was statistically significant.
-Both the intra-participant difference and the statistical significance test results are discussed in detail in the Appendix.}
+\new{We analyzed the intra-participant differences in submission times, which also supported these observations: the majority of submission times using \noah were faster than Excel---19 out of the 20 participants completed at least four tasks in less time using \noah.
+The submission time differences were more prominent for the steer, find, and \cmpB tasks.
+\new{In Section~\ref{sec:rq2}, we explain these outcomes in detail.
+For example, the aggregate column feature provides a faster alternative to steering, the overview-spreadsheet coordination accelerates raw data inspection, and the binned overview coupled with the context bar enables faster comparison.} The differences in submission times were statistically significant for all of the tasks except customize.
+Both the intra-participant differences and the statistical significance test results are discussed in detail in the Appendix.}
 
@@
 \begin{tabular}{c c} %trim=left bottom right top
-   \includegraphics[width=0.23\textwidth,trim={18 35 20 20},clip]{images/bird_accC.pdf} &
-   \includegraphics[width=0.23\textwidth,trim={18 35 20 10},clip]{images/airbnb_accC.pdf} \\
+   \includegraphics[width=0.23\textwidth,trim={18 20 20 20},clip]{images/bird_accC.pdf} &
+   \includegraphics[width=0.23\textwidth,trim={18 20 20 10},clip]{images/airbnb_accC.pdf} \\
    \textbf{(a) birdstrikes} & \textbf{(b) airbnb} \\
@@
 attained slightly higher accuracy
-with \noah compared to Excel. However, the difference in accuracies was statistically significant 
-for the steer tasks only (see Appendix). 
+with \noah compared to Excel. However, the differences in accuracies were statistically significant 
+for the steer tasks only (see Appendix). \new{We evaluate the usability of Excel and \noah next.} 
 \toappendix{Analysis of screen capture for \cmpA 
@@
 
-\subsubsection{Participants' preferences: \noah VS Excel} 
-\saj{Figure~\ref{fig:survey} shows a diverging stacked bar chart representation of the survey results in which participants rated their experience with Excel and \noah. 
+\subsubsection{Participants preferred \noah to Excel} 
+\new{Figure~\ref{fig:survey} shows a diverging stacked bar chart representation of the survey results in which participants rated their experience with Excel and \noah. 
 For each metric mentioned in Section~\ref{sec:study}, there are two stacked bar charts, one for Excel and one for \noah.
@@
 
-\subsection{RQ2. Impact of \noah and its components on spreadsheet navigation}
+\subsection{RQ2. Impact of \noah and its components on spreadsheet navigation experience}
 \label{sec:rq2}
-To answer RQ2, we assessed how \noah components impacted navigation in our study.
+\new{To answer RQ2, we assess how \noah's components, \ie the binned overview, aggregate column, and context bar, impacted participants' navigation.
+For each observation, we present participant feedback from the interview phase.}
 %We now discuss how \noah impacted participants' navigation experience compared to Excel.
 
-\subsubsection{Binned Overview}
-\stitle{Data navigation at scale.}
-\saj{Without an overview, participants found it difficult to 
-perform various tasks in Excel. One participant ($P11$) commented---\response{Excel can get overwhelming 
+\subsubsection{Binned Overview: Customizable Hierarchical Organization}
+\new{Overall, the binned overview prevented participants from being overwhelmed during navigation, especially at scale.
+Personalizing the overview enabled participants to define their own grouping of the data, resulting in a more meaningful overview presentation.
+However, the newer interactions at times deviated from spreadsheet semantics, contributing to a steeper learning curve.}
+
+\stitle{Overviews aid navigation at scale.}
+\new{Participants found it difficult to 
+perform various navigation tasks in Excel, especially
+at scale; \noah, on the other hand, helped participants
+avoid endless scrolling via clicking and semantically zooming on the overview, and provided cues for what to explore next via the bins of the overview. 
+One participant ($P11$) commented---\response{Excel can get overwhelming 
 if you have a lot of data in it and sometimes with that data 
 finding things can be difficult}. Participants ($N=6$) mentioned 
-that they would prefer \noah over Excel when the dataset is large: 
+that they would prefer \noah when the dataset is large: 
 \response{If I just had a large amount of data 
@@
 the overall structure of the data better 
-and prioritize the bin they want to visit next, enabling faster navigation. One participant ($P5$) commented: 
+and prioritize the bin they want to visit next. 
+One participant ($P5$) commented: 
 \response{I think it was just a little bit easier to navigate 
-and find where things were because you could already see what bins had what.} Another participant ($P1$) said:  
-\response{I like \noah a lot better. It was a lot easier to look up different data 
+and find where things were because you 
+could already see what bins had what.} Another participant ($P1$) said:  
+\response{I like \noah a lot better. 
+It was a lot easier to look up different data 
 and it was a lot quicker too}.}
@@
 
-\stitle{Enhanced navigation with guided bin customization.} 
-\saj{The bin customization feature enabled users to 
+\stitle{\new{Overview customization enables related data to be analyzed
+together in task-specific ways.}} 
+\new{Bin customization enabled participants to 
 personalize the overview based on their specific needs. 
@@
 \response{I did like the fact that it lets you take a data sheet and, in some way, containerize the stuff you care and the stuff you don't care about.} 
-14 out of 20 
-participants found the bin customization feature to be useful. 
-Participants preferred the feature to Excel's filtering feature 
+Participants (14 out of 20) 
+preferred the feature to Excel's filtering feature 
 when working with numeric data---\response{That was so much easier in \noah 
-than it was in Excel to be able to specify the range that you wanted it to go in} ($P17$).
-Our analysis of the video recordings revealed that, for the birdstrikes dataset in Excel, the customize task involved filtering out the desired values from a total of 451 unique values.
-As a result, participants had to manually filter a large number of values and took more time
-to submit their responses compared to \noah when they were able to use the bin customization feature. However, the time taken for this task was higher than  other tasks in \noah, as
+than it was in Excel to be able to 
+specify the range that you wanted it to go in} ($P17$). 
+Our analysis of the video recordings revealed that 
+for the birdstrikes dataset in Excel,
+the customize task involved filtering out 
+certain values from a total of 451 unique values. 
+This manual filtering led to a significant delays in
+task completion, compared to \noah where they were able to use the bin customization feature. 
+However, the time taken for this task was higher than  other tasks in \noah, as
 it required participants to restructure the overview
-before any calculation could be performed.
-Unfamiliarity with customization
-operations also contributed to higher task completion times.}
+before any calculation could be performed.}
 
-\stitle{Unfamiliar interactions.} \saj{Participants were unfamiliar with the bin customization feature.
-This unfamiliarity led to some participants ($N=5$ out of 20) 
-preferring Excel over \noah.
-One participant ($P11$) commented: \emph{Since I'm not used to spreadsheet data being presented that way, it took a little bit of getting used to.} Participants found some of the terminology
+\stitle{\new{Overview customization interactions have a steeper learning curve.}}
+\new{Unfamiliarity with the interactions required
+during customization in \noah
+also contributed to higher 
+task completion times for the customize task  
+compared to other tasks.
+The unfamiliarity led to some participants ($N=5$ out of 20) 
+preferring Excel over \noah for this task. 
+One participant ($P11$) commented: \response{Since I'm not used to spreadsheet data being presented that way, it took a little bit of getting used to.} Participants found some of the terminology 
 used in the interface---\eg explore, bin---quite 
 unfamiliar ($N=14$). 
-Moreover, two participants didn’t understand how the bins were constructed.}
+Moreover, two participants didn't understand how the bins were constructed and
+requested implementation details during the interview.}
 
-\stitle{Overview presentation across data types.}
-\saj{Although participants appreciated the binned representation of the overview for numeric data, a number of participants ($N=6$) stated that they would've preferred a pivot table-like single level overview for categorical data where each bin corresponds to one item.
+\stitle{\new{Tradeoffs between hierarchical and flat overviews.}}
+\new{While participants generally appreciated 
+the binned representation of the overview for numeric data, 
+a number of participants ($N=6$) stated that they would have preferred a non-hierarchical overview for categorical data, 
+where each bin corresponds to one item.
 One participant ($P13$) commented: \response{I would prefer it start with all the bins split, and then I can merge them as I want.} Another participant ($P4$) said---\response{When I started, it (\noah) had already grouped them, I think, alphabetically.
@@
 
-\subsubsection{Aggregate Column}
-\stitle{Ease of issuing formulae.} 
-\saj{The steer tasks required participants 
+\subsubsection{Aggregate Column: In-situ Steering-free Computation}
+\new{The aggregate column feature enabled participants to avoid cumbersome steering interactions, resulting in  faster and more accurate analysis compared to Excel.
+However, comparing the analysis results of more than two data subsets resulted in increased visual discontinuity and consequent errors.}
+
+\stitle{\new{Cumbersome steering replaced by a few button clicks with aggregate columns.}} 
+\new{The steer tasks required participants 
 to issue a \code{COUNTIF} formula on a data subset. 
@@
 the scrolling is a little bit cumbersome} ($P4$). 
-With \noah, participants avoided
-(a) scrolling by using clicking or zooming operations, 
-and (b) steering by performing aggregate operations on the overview. Participants ($N=13$) found it easier to issue formulae using the aggregate column feature.
+With \noah, participants avoided steering by \new{using aggregate column feature on the menu-bar and selecting the appropriate formula}.
+Multiple participants ($N=13$) found it easier to issue formulae using this feature.
 One participant ($P3$) commented: 
@@
 
-\stitle{Higher efficiency in issuing formulae.} 
-\saj{While the accuracies and submission times for the steer tasks in Excel varied significantly across datasets, using \noah, participants exhibited higher accuracies and faster submission times irrespective of the dataset (see Figure~\ref{fig:timeBox} and~\ref{fig:acc}).
+\stitle{\new{Issuing formulae is faster and more accurate with aggregate columns.}} 
+\new{While the accuracies and submission times for the steer tasks in Excel varied significantly across datasets, using \noah, participants exhibited higher accuracies and faster submission times irrespective of the dataset (see Figure~\ref{fig:timeBox} and~\ref{fig:acc}).
 The automated and steering-free aggregate column feature of \noah contributed to high accuracies (100$\%$) for the steer tasks.
@@
 you don't have to highlight every number versus Excel where you actually have to select everything.} All of the 14
-inaccurate submissions with Excel involved steering incorrect spreadsheets regions.
-11 of the inaccurate submissions were with the Airbnb dataset.
-However, in \noah, participants were able to avoid steering by utilizing the aggregate column feature.
-Analysis of screen recordings of Excel usage
-revealed that for birdstrikes dataset, several participants 
+inaccurate submissions with Excel involved steering an incorrect spreadsheet region; 11 of the inaccurate submissions were with the Airbnb dataset.
+In \noah, participants were able to avoid steering by using the aggregate column feature.
+Analysis of screen recordings
+revealed that, for Excel, for the birdstrikes dataset, several participants
 used the \emph{autosum}
-feature to quickly count the number of $1$'s in a binary-valued column that was involved in the steering task. Summing up
+feature to quickly count the number of $1$'s in the binary-valued column involved in the steering task. Summing up
 binary values is equal to the number of $1$'s in the collection. 
@@
 of the cells in the selected column. 
-In both cases, participants avoided steering the data. 
-On the other hand, for Airbnb dataset, participants 
-could not use these shortcuts as the column that was involved in the steering task was non-binary 
-(it had 365 different values). Participants, at times, steered incorrect regions, resulting in inaccurate responses for the task.  
-Therefore, the participants’ ability to avoid steering 
-depended on the data type. 
-Failure to avoid steering often led participants
-to selecting an incorrect range of data ($N=14$), resulting in incorrect
-responses.}
+In both cases, participants avoided steering the data resulting in fewer errors. 
+On the other hand, for the Airbnb dataset, participants 
+could not use these shortcuts as the column involved in the steering task was non-binary 
+(it had 365 different values). Failure to avoid steering often led participants
+to select an incorrect range of data ($N=14$ cases), resulting in incorrect
+responses.  
+Therefore, the participants' ability to avoid steering 
+depended on the data type.}
 
-\subsubsection{Detailed View}
-\stitle{Accelerated data inspection.} 
-\saj{For the \emph{identify} task, 
-participants had to skim through all the cells 
-in the current window in Excel, 
-resulting in higher completion times. 
-%Again, participants had to skim over binary values when counting in birdstrikes versus non-binary values for Airbnb, resulting in even higher completion times for the latter. 
-Even though Excel provides a conditional formatting feature, that adds one additional step when performing the \emph{idenitfy} task. 
-In \noah, participants benefited from having visual cues in the form of colored cells, helping them relate the aggregate column with the raw data---\emph{You didn't have to do any additional steps and it was a visual cue right there, made it very quick to count it up ($P17$).} Another participant ($P9$) commented---\response{In Excel, you would have to do your own condition of formatting.
-But you have to build that every time  you need to ask a question.
-This one (\noah) at least something is pre-built in, and you can easily count.} However, one participant ($P3$) pointed out the fact that, when the data corresponding to the bin does not fit in the screen, they had to scroll through the data to identify relevant information.}
-%Out of 20 participants, 16 participants preferred the automatic data highlighting feature of \noah while performing the \emph{identify} task.
-Moreover, the colored cells focused their attention to relevant regions within the spreadsheet Another participant ($P5$) mentioned the fact that some of the color choices would make \noah unusable for the colorblind population.}
+\stitle{\new{Visual discontinuity during comparison while reduced, was not completely eliminated.}} 
+\new{For \cmpB tasks, participants 
+had to perform $N$ comparisons in \noah while issuing the aggregate column operation once.
+However, the comparison among $N$ bins resulted in increased visual discontinuity.
+This lead to some ($N=4$ out of 20) incorrect submissions.
+In Excel, the experience was worse, as the participants had to perform $N$ separate steering tasks.
+As a result, in Excel, the \cmpB task submission times were very high compared to \cmpA tasks (see  Figure~\ref{fig:timeBox}).
+In addition, the accuracies of the \cmpB task in Excel were lower ($N=7$ out of 20 submissions were inaccurate).}
+\cut{aggregate  column  is  kept  in
+sync  with  the  bins  a}
 
-\subsubsection{Context Bar}
-\stitle{Utilizing history to avoid repeated execution.}
-\saj{For the \cmpA task, participants utilized the context bar to navigate to the previously visited bin for the first steer task.
-As \noah automatically materialized the aggregate summaries of a previously visited bin, participants were able to view the aggregate column values instantly without having to repeat the same command.
-However, as Excel did not preserve any navigation history, participants had to re-execute the first steering operation.
-As a result, the submission times for \cmpA tasks were lower in \noah compared to Excel (see Figure~\ref{fig:timeBox}).
+\subsubsection{\new{History, Context, and Coordination}}
+\new{The context bar enabled participants to revisit previously explored bins.
+The aggregation results corresponding to that bin were automatically updated in the aggregate column, due to automatic syncing between the binned overview and aggregate column (see Section~\ref{sec:agg_col}).
+ The coordination between the overview and the raw spreadsheet data further helped participants relate the aggregate column results with the raw data.}
+
+\stitle{\new{History helps avoid repeated interactions.}}
+\new{For the \cmpA task in \noah, all of the participants used the context bar to navigate to a bin previously visited for the first steer task.
+As the bin currently being displayed was changed, the aggregate column was automatically updated to display values corresponding to that bin, enabling participants to view the aggregate column values instantly without having to reissue the operation.} \new{On the other hand, as Excel did not preserve any navigation history, participants had to re-execute the first steering operation.
+As a result, the submission times for \cmpA tasks were faster in \noah compared to Excel (see Figure~\ref{fig:timeBox}).
 One participant ($P16$) commented---\response{Once I got familiar with the interface, it was easy to just say, I want to see this state, and I like that fact that like automatically it goes into the bins on NOAH, gave me summary information.} Another participant ($P9$) said---\response{Noah was easy to find and compare and toggle in between.} }
-
-\stitle{Visual discontinuity during navigation.} 
-\saj{For \cmpB tasks, participants 
-had to perform $N$ comparisons in \noah while issuing the aggregate column operation once.
-However, comparison among $N$ bins resulted in increased visual discontinuity.
-This lead to some ($N=4$ out of 20) incorrect submissions.
-In excel, the experience was worse, as the participants had to perform $N$ steering tasks.
-As a result, in Excel, \cmpB task submission times were very high compared to \cmpA tasks (see  Figure~\ref{fig:timeBox}).
-In addition, the accuracies of the \cmpB task in Excel were extremely low.}
 
@@
 Even in \noah, comparisons between multiple values resulted in increased visual discontinuity leading to some ($N=4$ out of 20) incorrect submissions.
+\stitle{\new{Overview-spreadsheet coordination helps relate interactions on the overview with the raw data.}}
+\new{The coordination between the overview and spreadsheet in \noah enabled users to quickly relate their interactions on the overview with the raw spreadsheet data. For example, for the find task, 
+participants had to find all the cells within the spreadsheet that satisfy a condition corresponding to the preceding steer task.
+To do so, they had to
+skim through all the cells 
+in the current window in Excel, 
+resulting in higher completion times.
+%Again, participants had to skim over binary values when counting in birdstrikes versus non-binary values for Airbnb, resulting in even higher completion times for the latter. 
+Even though Excel supports a conditional formatting feature\footnote{\new{The conditional formatting feature in Excel enables users to request cells that satisfy certain conditions to be colored.}}, issuing the feature added one additional step when performing the find task.}
+\new{In \noah, participants benefited from having visual cues in the form of automatically colored cells, helping them relate the aggregate column with the raw data---\response{You didn't have to do any additional steps and it was a visual cue right there, made it very quick to count it up ($P17$).} Another participant ($P9$) commented---\response{In Excel, you would have to add your own condition for formatting.
+But you have to build that (conditional formatting) every time  you need to ask a question.
+This one (\noah) at least something is pre-built in, and you can easily count.} However, one participant ($P3$) pointed out that, when the data corresponding to the bin does not fit in the screen, they had to scroll through to identify relevant information.}
+%Out of 20 participants, 16 participants preferred the automatic data highlighting feature of \noah while performing the \emph{identify} task.
+Moreover, the colored cells focused their attention to relevant regions within the spreadsheet Another participant ($P5$) mentioned the fact that some of the color choices would make \noah unusable for the colorblind population.}

### 2019-11-25 study.tex
final tvcg

@@
+%!TEX root = main.tex
+
 \section{Evaluation Study Design}
@@
 \label{sec:study_design}
-\new{Our goal is to study the impact of an in-situ navigation plugin for spreadsheets, \noah, on navigation and exploration of data.
-Therefore, we decided to compare a \noah-integrated spreadsheet system with a typical, popular one, Excel, across various tasks.
+\new{Our goal is to study the impact of an in-situ navigation 
+plugin for spreadsheets, \noah, on navigation and exploration of data.
+Therefore, we decided to compare a \noah-integrated 
+spreadsheet system with a typical, popular one, Excel, across various tasks.
 Similar domain specific-evaluations have been performed for evaluating various  overview+detail interfaces, \eg database browsing~\cite{north2000snap} or
@@
 \item {\bf RQ1.} %\sout{How does the addition of an overview help address spreadsheet navigational challenges, with respect to quantitative performance metrics such as speed and accuracy?}
-\new{How does integration of an overview plugin like \noah impact a) the completion time and accuracy of various spreadsheet navigational tasks\newsaj{, and b) the usability of spreadsheet systems?}} \saj{SAJ: the survey results are also part of this RQ}
+\new{How does the integration of an overview plugin like \noah impact the efficiency of navigation
+within and the usability of spreadsheet systems?}
 \item {\bf RQ2.} %\sout{How do users make use of the aggregate column results in conjunction with the raw spreadsheet data to analyze data?} 
 \new{How do the various components of \noah impact users' navigational experiences?}
+
 %\item {\bf RQ3.} \sout{How does bin customization help users inject domain knowledge in addressing their specific exploration questions compared to the automated overview?} \sout{\saj{How do users utilize additional features of \noah, \eg bin customization, during navigation?}
@@
 \stitle{Study Design.} We conducted a $2 \times 2$ (2 datasets, 2 tools) mixed design within-subject study.
-The two tools used in the study were: Microsoft Excel, and \noah integrated within {\scshape DataSpread}~\cite{dataspread}.
-\new{As mentioned previously, we chose Excel for our comparative study because it is the most popular spreadsheet in use today.}
+The two tools used in the study were: Microsoft Excel, and \noah integrated within {\scshape DataSpread}~\cite{dataspread} \new{(henceforth, referred to as \noah for succinctness).
+As mentioned previously, we chose Excel for our comparative study because it is the most popular spreadsheet in use today.}
 The study consisted of three phases: (a) an introductory phase explaining
@@
 
-\stitle{Datasets.} We used two datasets---the birdstrikes (\new{used for evaluating visual data exploration or recommendation tools like Keshif~\cite{yalccin2018keshif} and Voyager~\cite{wongsuphasawat2016voyager}}), and the Airbnb~\cite{web:airbnb} datasets.
+\stitle{Datasets.} We used two datasets---the birdstrikes (\new{used for evaluating visual data exploration tools like Keshif~\cite{yalccin2018keshif} and Voyager~\cite{wongsuphasawat2016voyager}}), and the Airbnb~\cite{web:airbnb} datasets.
 These datasets were chosen for their understandability to a general audience.
@@
 prospective participants filled out an interest form
-where they answered questions about their spreadsheet expertise\newsaj{, spreadsheet usage purposes, and spreadsheet operations they typically use}.
+where they answered questions about their spreadsheet expertise\new{, their typical goals when using spreadsheets, and the spreadsheet operations they typically use}.
 Participants were asked to rate their expertise with
-different spreadsheet software, \eg Excel and Google Sheets, 
+different spreadsheet systems, \eg Excel and Google Sheets, 
 and their frequency of using various spreadsheet 
@@
 We also asked participants about their familiarity 
-with basic mathematical and statistical spreadsheet functions
+with basic mathematical and statistical spreadsheet functions,
 as well as advanced operations, \eg pivot table,
@@
  \vspace{-10pt}
-\caption{\newsaj{Quiz tasks for the birdstrikes dataset. The use cases correspond to the task typology discussed in Section~\ref{sec:usage}}}
+\caption{\new{Quiz tasks for the birdstrikes dataset. The task purposes and use cases correspond to the task typology discussed in Section~\ref{sec:usage}.}}
 \label{tab:questions}
 \centering
-\begin{tabular}{l l l}
+\new{
+\begin{tabular}{l l}
 \hline
-Category & Question & Use case    \\ \hline
-steer                & Organize the data by State. How many& \code{locate} $\rightarrow$\\
-                     & flights that had damages (damage = 1)& \code{summarize}\\
-                     & originated from Florida? \\ 
-identify             & How many flights in the currently &\\
-                     & visible spreadsheet window  &  \code{identify}\\ 
-                     & have damages? & \\
-steer                & How many flights that had damages & \code{locate} $\rightarrow$\\
-                     & originated from California?                         & \code{summarize}\\ 
-\cmpA                & Which state between Florida and &\\
-                     & California  has a higher number of & \code{compare}\\ 
-                     & flights with damages?    &\\
-\cmpB                & Find the state with the most & \code{compare}\\
-                     & birdstrike occurrences.  &\\ 
-customize            & Organize the data by \emph{altitude}. &\\
-                     & What is the average cost of damages & \code{generate}\\
-                     &    for altitude bin 0-450?         & \\ \hline
+Category & Question (Q), Purpose (P), Use case (U)    \\ \hline
+steer                & \textbf{Q}: Organize the data by State. How many flights that had \\
+                     & damages (damage = 1) originated from Florida?, \\
+                     & \textbf{P}: \emph{Search}$\rightarrow$\emph{Query}, \textbf{U}: \code{lookup}$\rightarrow$\code{identify}\\ \hline
+find                 & \textbf{Q}: How many flights in the currently visible spreadsheet \\
+                     & window have damages?, \textbf{P}: \emph{Search}, \textbf{U}:  \code{browse}\\ \hline 
+steer                & \textbf{Q}: Organize the data by State. How many flights that had \\
+                     & damages (damage = 1) originated from California?, \\
+                     & \textbf{P}: \emph{Search}$\rightarrow$\emph{Query}, \textbf{U}: \code{lookup}$\rightarrow$\code{identify}\\ \hline
+\cmpA                & \textbf{Q}: Which state between Florida and California has a higher\\
+                     & number of flights with damages?, \textbf{P}: \emph{Query}, \textbf{U}:  \code{compare}\\ \hline
+\cmpB                & \textbf{Q}: Find the state with the most birdstrike occurrences,\\
+                     & \textbf{P}: \emph{Query}$\rightarrow$\emph{Search}, \textbf{U}: \code{summarize}$\rightarrow$\code{locate}\\ \hline
+customize            & \textbf{Q}: Organize the data by \emph{altitude}. What is the average cost of\\
+                     & damages for altitude bin 0-450?, \textbf{P}: \emph{Query}$\rightarrow$\emph{Search}$\rightarrow$\emph{Produce}, \\
+                     & \textbf{U}: \code{generate}$\rightarrow$\code{summarize}$\rightarrow$\code{lookup}\\
 \end{tabular}
+}
 \end{table}
@@
 We developed an online JavaScript-based quiz system that recorded user responses and submission times. 
-We also recorded the participants’ interactions
+We also recorded the participants' interactions
 with both tools using screen capture software. 
@@
 
-\emph{Quiz Tasks.} \new{We designed six tasks grouped into four categories: 
-steer (two tasks), compare (two tasks), identify (one task), and customize (one task), representing the Table~\ref{tab:scope} use cases \code{locate}, \code{summarize}, \code{compare}, \code{identify}, and \code{generate}, respectively.
-\newsaj{These use cases are representative of the navigation interactions required for the most frequently issued spreadsheet formulae
-~\cite{bradbard2014spreadsheet, lawson2009comparison}.} The tasks were presented in the same order as shown in Table~\ref{tab:questions} for the birdstrikes dataset.
-The tasks for the Airbnb dataset were similar.
-The order of the tasks mimics the spreadsheet navigation scenario presented in Section~\ref{sec:usage} \newsaj{where the task complexity increases as the user continues to navigate and analyze the data.}}
-%The customize tasks was designed to mimic the scenario where a user would group data subsets of interest and required participants to utilize the bin customization feature. 
-%All of these tasks require navigation of the spreadsheet data in the form of either scrolling or steering.
+\emph{Quiz Tasks.} 
+\new{We designed six tasks across five categories: 
+steer (two tasks), find (one task), \cmpA (one task), \cmpB (one task), and customize (one task), encompassing six of the seven task typology use cases 
+underlying the \emph{Search}, \emph{Query}, 
+and \emph{Produce} purposes: \code{lookup}/\code{locate}, \code{identify},
+\code{browse}, \code{compare}, \code{summarize}, and \code{generate} 
+(see Table~\ref{tab:scope})\footnote{\new{We omitted the \code{export/save} use-case under the \emph{Produce} purpose since
+\noah is not targeted at improving that use case. 
+We also did not
+study the \emph{Consume} purpose so as to focus our evaluation on completion 
+of tasks as opposed to an open-ended exploration setting, which is beyond
+the scope of our study.}}.
+These tasks were selected to mimic a typical spreadsheet analysis
+workflow and are representative of navigation interactions 
+required for the most frequently issued spreadsheet operations~\cite{bradbard2014spreadsheet, lawson2009comparison}.
+The tasks were presented in the same order as shown in Table~\ref{tab:questions} for the birdstrikes dataset.
+%The order of the tasks mimics a spreadsheet navigation scenario where a user begins by casually exploring a dataset (\emph{Consume}), \eg organizing the birdstrike data by state, and then engaging in more focused set of tasks to satisfy various purposes, \eg \emph{Search} and \emph{Query} data, \emph{Produce} new information.
+The tasks for the Airbnb dataset mimic a scenario similar to the example in Section~\ref{sec:usage}. We explain the scenario in the context of the birdstrikes dataset next.}
 
-\newsaj{The steer tasks required the participants to first \code{locate} a data subset and then \code{summarize} the subset by issuing an aggregate operation, \eg \code{COUNTIF}.
-Following summarization, the identify task required participants to interact with the detailed view and \code{identify} data points within the spreadsheet that satisfy the condition of the first steer task.
-The purpose of the second steer tasks was to enable the \cmpA task---given the two summaries obtained from the two steer tasks, the \cmpA task asked the participants to \code{compare} the results of both.
-Then the participants were asked to perform an even more complex comparison, \ie the \cmpB task, which involved comparing the same statistics computed in the two steer tasks of all the data subsets.
-The purpose of the \cmpB task was to see how increasing the number of subsets would affect navigation.
-Finally, the customize task was designed to mimic the scenario where the \noah-generated overview doesn't reflect the user preference and a user would group data subsets of interest to create a new overview.
-(accomplished by utilizing the bin customization feature in \noah).}
+\new{Say a user is interested in analyzing bird-strike statistics across US states.
+As the user is from Florida, she starts by computing bird-strike occurrences for that state (steer) and finds the number of occurrences to be quite high.
+After looking at the aggregate statistics, she decides to examine specific instances of bird-strike occurrences to inspect other attributes, \eg the altitude where the strike happened or the species of the bird (find).
+She notices a bird-strike occurrence at 50ft, which is surprising.
+She decides to investigate this issue later.
+For now, she focuses her attention on analyzing the state-wise statistics.
+She computes the same occurrence statistics for another large state, say California (steer) and then compares the statistics between the two large states (\cmpA).
+At this point, she becomes interested in learning the occurrence statistics across all states and in finding the state with the highest bird-strike occurrences (\cmpB).
+With the state-wise comparison completed, she decides to revisit low altitude bird-strikes.
+So she organizes the data by \emph{altitude} and computes the occurrence statistics at low altitudes, along with the associated damages (customize).
+For this final task, the bins generated by \noah did not correspond to the given altitude range and would require bin customization to compute the statistics}.
 
-%\newsaj{The steer tasks required participants to first \code{locate} a data subset and then issue a formula, \eg \code{COUNTIF}, via steering in Excel to \code{summarize} the subset (accomplished by the aggregate column feature in \noah).
-Following summarization, the identify task required participants to interact with the detailed view and \code{identify} data points within the spreadsheet that satisfy the condition of the first steer task (accomplished by inspecting the colored spreadsheet cells in \noah).
-The purpose of the second steer tasks was to enable the subsequent comparison tasks.
-Given the two summaries obtained from the two steer tasks, the participants were then asked to \code{compare} the results of both, \ie the \cmpA task (accomplished by utilizing the context bar feature of \noah to revisit the results of the first steer task).
-The \cmpB task involved comparing the same statistics computed in the two steer tasks of all the data subsets.
-The purpose of the \cmpB task was to see how increasing the number of subsets would affect navigation.
-Finally, the customize task was designed to mimic the scenario where the \noah-generated overview doesn't reflect the user preference and a user would group data subsets of interest to create a new overview.
-(accomplished by utilizing the bin customization feature in \noah).}
-
-%As participants use both \noah and Excel to perform these tasks, by examining participants' task performance, we can understand a) how the addition of an overview affect participants' navigation experience (\emph{RQ1}), b) what are the impacts of the individual features of \noah in spreadsheet navigation (\emph{RQ2}), and c) how the participants utilize the features supported by \noah during navigation (\emph{RQ3})
+%\hide{First,  the  user  is  interested  in  bird-strike  statistics  in  a state  of  interest,  say  Florida  (steer: lookup state  and  then identify the number of bird-strike occurrences).
+Then the user wants to find the actual bird-strike instances (find: \code{browse} data) within the spreadsheet region currently visible in the screen to obtain other details about the occurrence, for example, the altitude at which the strikes happened.
+Next, the user wants to verify whether the strike occurrences are similar for larger states (steer: \code{lookup} another state say California and then \code{identify} the number of bird-strike occurrences).
+Then the user \code{compare}s the occurrence statistics of the two states (\cmpA).
+Eventually the user decides to compare the same statistics for all the states in the dataset and find the state with the highest bird-strike occurrences (\cmpB: \code{summarize} all the states and then \code{locate} the maximum value state).
+Finally, the user wants to \code{summarize} the data by another attribute say altitude and \code{lookup} the bird-strike occurrence for a specific altitude range (customize).
+Note that in this scenario, the bins generated by \noah did not correspond to the given altitude range and would require the participant to utilize the bin customization feature to \code{generate} a customized overview first and then \code{summarize} and \code{lookup}.}
 
@@
 15 in total, to ensure reliability. 
-Participants were also asked to mention the positive 
+Participants were also asked to describe the positive 
 and negative aspects of both tools. 
-\newsaj{The survey was designed to evaluate \emph{RQ1b}}.
+%\newsaj{The survey was designed to evaluate \emph{RQ1}}.
 
@@
 completion time for each of the six tasks. 
-\newsaj{We combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be corroborated across multiple information sources.
-Moreover, we analyzed the survey responses to quantify the usability of both Excel and \noah-integrated {\scshape DataSpread}.}
+\new{We combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be corroborated across multiple sources.
+Moreover, we analyzed the survey responses to quantify the usability of both Excel and the \noah-integrated spreadsheet.}
 %For example, we analyzed the video recordings of participants' interaction with the tools during the quiz phase.
@@
 different features provided by \noah and Excel.
-
-\subsection{Study Limitations.}
-\newsaj{Our study has several limitations that can be strengthened by future larger-scale and more fine-grained studies.
-Firstly, our participant pool demographics partially represents the demographics of the general audience intended for \noah.
-A larger sample with more participants that better represents the spreadsheet user population would have provided more ecological validity to generalize our findings.
-}
-%While a larger sample with more diverse backgrounds would have allowed us to perform more definitive quantitative analysis, we combined this analysis with qualitative survey, interview, and screen/audio recording data to provide insights that can be explained with multiple information sources.
-
-\newsaj{Next, we only compared the performance of a \noah-integrated spreadsheet with a traditional spreadsheet.
-We did not evaluate specific spreadsheet features like pivot table and \code{SUBTOTAL} as they violate most of the design considerations proposed in Section~\ref{sec:design}.
-We discussed their limitations in Section~\ref{sec:related}.
-Moreover, allowing the participants to utilize the typical spreadsheet operations that they were comfortable with, enabled us to observe how introduction of \noah affected their navigation experience.}
-
-\new{Finally, we did not isolate the effects of the individual features of \noah to better understand the implications of those.
-For example, the effect of binned overview (visual clarity versus visual continuity), display layout (screen space trade-off), and
-contextual presentation of data (raw text versus chart representation of aggregate columns).
-However, the goal of the study was to understand participants' navigation experience in the presence and absence of \noah.
-A more fine-grained study that teases apart the contribution of individual components of \noah is warranted.}
-
-\begin{comment}
-\infovis{
-\begin{itemize}
-    \item Justify the choice of comparing to Excel and note the limitations of that approach (see section 6.1 intro and the {\bf study design paragraph})
-    \item Better justify the choice of tasks in the study (see section 6.2 \emph{quiz tasks})
-    \item Explain the choice of the four research questions. (see section 6.1 intro)
-    \item Provide more details about the study, including reporting intra-participant
-    differences (see section 7.2)
-\end{itemize}
-}
-\end{comment}

### 2020-03-10 rebuttal.tex
Updates from Overleaf

%!TEX root = main.tex

\section*{Rebuttal to reviews of InfoVis 2019 submission ``Extending Spreadsheets to Support Seamless
Navigation at Scale''}

A previous version of this paper was submitted to InfoVis 2019 (submission ID: 1043). This version was rejected as the paper co-chairs deemed that revising the paper would require more time than allocated to meet reviewer expectations. The paper co-chairs mentioned the following: ``\textit{We particularly encourage such revisions where submissions were positively received by reviewers, but the revisions required were deemed to be beyond the scope of the conference review cycle. If you address the issues raised and subsequently submit to TVCG, please make reference to this InfoVis submission and include a description of how you addressed the InfoVis reviewers' comments.}'' 
The summary review asked for the following changes in a revision:
\begin{itemize}
    \item[\textbf{C1:}] {\em More clearly articulate the merits over alternative tools, such as Keshif.} 
     \item[\textbf{C2:}] {\em Convincingly explain the design choice around the binning mechanism or revise
    that approach.}
    \item[\textbf{C3:}] {\em Justify the choice of comparing with Excel and note the limitations of that
    approach.}
     \item[\textbf{C4:}] {\em Better justify the choice of tasks in the study.} 
     \item[\textbf{C5:}]{\em Explain the choice of the four research questions.}
     \item[\textbf{C6:}] {\em Provide more details about the study, including reporting intra-participant
    differences.}
\end{itemize} 
We thank the reviewers and meta-reviewer for their detailed and
constructive feedback.
Taking this feedback into account, 
we have spent the last four months 
preparing a revised version of this paper.
We believe the paper is substantially
stronger as a result. 
Every section in the paper has been revised. The changes are explained in the ``summary of changes'' document, where
the changes are highlighted in \new{blue} for reviewers' convenience.

We realize that these extensive changes throughout the paper have increased the paper length. As the TVCG submission guidelines indicate ``\emph{The regular paper page length limit is defined at 12 formatted
pages, including references and author biographies. Any pages or
fraction thereof exceeding this limit are charged \textdollar220 per page. Regular
papers may not exceed 18 formatted pages}'', so we have decided to let the paper length stand since our paper is at 16 pages currently — and we would be happy to pay for the extra pages. We would also be happy to reduce length by moving appropriate content from the main body to the supplementary material as suggested by reviewers.

We briefly describe the changes to each section
before returning to the changes required in the summary review  
({\bf C1}--{\bf C6}). Finally, we provide a point-by-point response to the reviewer comments.

\subsection*{Changes Organized by Section}
\begin{itemize}
	\leftmargin=25pt \rightmargin=0pt \labelsep=5pt \labelwidth=20pt \itemindent=0pt \listparindent=0pt \topsep=0pt plus 2pt minus 4pt \partopsep=0pt plus 1pt minus 1pt \parsep=0pt plus 1pt \itemsep=\parsep
\item[\textbf{S1}]
In Section~\ref{sec:intro}, we now cite
work that highlights the widespread adoption
of spreadsheets even among users of 
advanced enterprise solutions, further justifying 
why we focused on addressing navigation challenges within
spreadsheets. In fact, recent 
debate within the visualization community following
VIS 2019 also echoes this view\footnote{https://twitter.com/FILWD/status/1187411664611749888}---instead of designing a new sophisticated tool from scratch that 
may cater to a small population, 
we aim to enhance the user experience for
existing spreadsheet tools
with nearly a billion users.
We further augment   
the definitions of spreadsheet 
interactions like scrolling and steering in Section~\ref{sec:intro}, 
and now clearly articulate the challenges of 
designing a general-purpose plug-in for spreadsheets.
\item[\textbf{S2}]
In Section~\ref{sec:usage}, 
our usage scenario has been substantially revised 
using Brehmer and Munzner's typology~\cite{brehmer2013multi} to
clarify the scope of tasks supported by spreadsheets, as well
as those that are enhanced
when using a spreadsheet with a \noah plug-in; our new Table~\ref{tab:scope}
provides use-cases supported within \noah
for each category.
\item[\textbf{S3}]
In Section~\ref{sec:related},
we now clearly
articulate the differences in goals
of spreadsheets and tabular data analysis tools (TDA),
further justifying our choice of Excel,
the most widely used spreadsheet tool,
as our baseline in the evaluation study.
\item[\textbf{S4}]
In Section~\ref{sec:design},
we now explain the  
third design consideration 
for \noah more clearly
(\textbf{DC3})---focusing on motivating the binning
mechanism.
\item[\textbf{S5}] In Section~\ref{sec:ui}, 
we now justify why we are developing a 
multi-granularity overview,
and explain why 
we opted for histograms as an overview representation, by contextualizing our
approach using prior work on multi-scale aggregation~\cite{elmqvist2009hierarchical}. 
\item[\textbf{S6}]
In Section~\ref{sec:study}, 
we further clarified the goals of our study
and reframed our research questions to 
better reflect those goals. 
Our research questions haven't been substantially altered;
rather, they have been grouped together and made more precise. 
In this version,
we now explain our choice
for the quiz tasks in detail. 
\item[\textbf{S7}] Section~\ref{sec:results} has been completely
revamped to better reflect our research questions,
while adding more qualitative observations 
regarding the user experience with \noah. 
We now
include additional analysis results on intra-participant differences. 
\item[\textbf{S8}] In Section~\ref{sec:discuss}, we have added a summary of our takeaways from the evaluation study, while discussing limitations of our evaluation study and \noah's design while highlighting future enhancement opportunities. 
\end{itemize}

\subsection*{Response to summary review (C1--C6)}

\begin{itemize}
\leftmargin=25pt \rightmargin=0pt \labelsep=5pt \labelwidth=20pt \itemindent=0pt \listparindent=0pt \topsep=0pt plus 2pt minus 4pt \partopsep=0pt plus 1pt minus 1pt \parsep=0pt plus 1pt \itemsep=\parsep

\item[\textbf{R(C1)}] \textit{Merits over other tools} (\textbf{C1}): 
Throughout the paper (and especially in 
Sections~\ref{sec:intro}, \ref{sec:related},
and \ref{sec:study}), we have emphasized
that the main contribution of \noah
is its \emph{design as a general-purpose navigation plug-in} 
to any existing spreadsheet tool. 
As spreadsheets have a massive user-base, 
enhancing exploration and formula computation 
on large datasets while maintaining 
the spreadsheet look-and-feel as much as possible, 
has the potential to impact hundreds of millions 
of spreadsheet users. 
Most of these users employ spreadsheets 
as their primary data management and analysis 
tool while shunning enterprise solutions 
with more advanced features. 
Therefore, our goal is to develop a 
solution to improve navigation within spreadsheets.
Indeed, we could have tried to enhance navigation
in other tools, such as Keshif, or Tableau, like
the reviewers suggest---but we would be forcing
spreadsheet users to adopt an entirely new tool, something
that they are clearly loath to do.

\item[\textbf{R(C2)}] \textit{Explanation of the binning mechanism} (\textbf{C2}): 
We now explain our choice
of the binning mechanism in 
the context of the third design consideration (\emph{DC3}),
as detailed in Section~\ref{sec:design}.
We further expand on this in the blurb titled 
``Why a Multi-granularity Binned Overview'' 
in Section~\ref{sec:ui}.
In brief, we opted for binning to 
provide a clear and concise 
representation of the overall data 
distribution while minimizing user's back 
and forth movement across multiple screens 
during navigation. 
At some level, we do need to limit the number
of values displayed so that the overview
fits on the screen: binning is a natural solution
for that issue. 
We cite related work on 
multi-scale aggregation~\cite{elmqvist2009hierarchical} 
to motivate the choice of 
a multi-granularity overview
and show that binned aggregation via histograms~\cite{liu2013immens}
is a suitable representation of such an overview.
However, we acknowledge limitations 
for this choice for binning categorical data 
and discuss possible solutions 
in Section~\ref{sec:discuss} 
in the context of our user study findings.

\item[\textbf{R(C3)}] \textit{Justification of Excel as a Baseline} (\textbf{C3}):
Since our goal is to improve spreadsheet
user experience---and there are very good reasons for doing so,
as outlined in item 1) above---the natural comparison point
is Excel. 
That said, beyond popularity, 
we have provided a thorough justification
for what spreadsheet tools like Excel offer
relative to tabular data analysis tools (TDA),
making it a more appropriate point of comparison.
Our justification for using Excel over 
tabular data analysis (TDA) tools involves 
a) highlighting the appeal of spreadsheets 
among users who shun more advanced enterprise 
tools (Section~\ref{sec:intro}), 
b) identifying the scope of tasks 
supported by spreadsheets and 
TDA tools (Section~\ref{sec:usage} and~\ref{sec:related}), 
c) explaining that the goal of spreadsheets 
is to present the raw data as is, 
amenable to editing, formula computation, 
and comparison of derived data, all done in-situ,
unlike TDA tools that hide the data 
while providing summarized statistics 
(Section~\ref{sec:related}), and 
d) highlighting the importance of 
addressing the shortcomings of a tool, 
the user base of which far exceeds 
that of TDA tools (Section~\ref{sec:study}).

\item[\textbf{R(C4, C5)}] \textit{Enhancing the user study design} 
(\textbf{C4, C5}):
We have added additional 
explanations regarding the goals 
of the study while better articulating 
the research questions, 
provided a better justification 
for the choice of the tasks for the study (see Section~\ref{sec:study_design}) 
and added a detailed discussion on the limitations of the study. 
Specifically, we merged our previous \emph{RQ1} (evaluating the quiz task performance) 
with \emph{RQ4} (summarizing user feedback on performance via a survey) 
into a unified research question (new \emph{RQ1}) 
on the participants' navigation performance. 
Moreover, we merged previous \emph{RQ2} and \emph{RQ3} 
that explored how specific features of \noah 
affected participants' navigation experience 
into one broad question (new \emph{RQ2}) 
that explores their overall experience with \noah and its components. 
Moreover, we have justified the choice of the quiz tasks by 
a) relating their goals with the scope of tasks supported by spreadsheets (see Section~\ref{sec:usage}), 
and b) explaining how these tasks can reveal how people may use \noah 
for spreadsheet navigation. 
Finally, we have identified three limitations of the study, 
resulting from the sample size and the study design, which we justify based on the scope of the study.

\item[\textbf{R(C6)}] \textit{Reporting Intra-participant Differences} (\textbf{C6}):
We performed additional analysis of the study data to report intra-participant submission time differences for the quiz phase tasks. While we summarize the results of the analysis in Section~\ref{sec:time_acc}, the detailed results can be found in Appendix. We find that despite a few exceptions, across all tasks, participants' task submission times were faster using \noah---all but one participants completed at least four tasks in less time using \noah compared to Excel. However, we already obtained similar observations from Figure~\ref{fig:timeBox} in Section~\ref{sec:rq1}. Therefore, we decided to include the results of the intra-participant analysis in the Appendix as opposed to the main paper.
\end{itemize}

\subsection*{Point-by-point response to reviews}

\subsection*{Primary Reviewer (PR) review}

\stitle{PR1a.} \textit{The
    treatment of Tabular Data Analysis tools is to short, and the argument that they
    are "not co-located with the underlying data, introducing visual discontinuities
    and lack of context" is IMO not true. This is also the main weakness of the paper: A tool like Keshif supports most things that NOAH supports, it certainly supports all the things that the study
    tests.}

\stitle{Response.} 
\resp{In our revised version, 
we now clearly articulate the difference 
in goals between spreadsheets and TDA tools 
in Section 2.
Specifically, regarding Keshif,
the goal of Keshif is to auto-generate
sophisticated dashboards with customizability
built-in to maximize the number of insights gained.
However, with Keshif, a user cannot
simultaneously examine a summary dashboard alongside
the raw data---the lowest level of detail 
a user can see is the unique values for a given column,
and not the raw data underlying each unique value. 
There are many other tools like Keshif that support
advanced analytics capabilities. 
However, 
as we argue in {\bf R(C1)} and {\bf R(C3)} and 
in changes to the paper listed in {\bf S1} and {\bf S3},
these advanced tools do not have the adoption that spreadsheets have.
So instead of developing yet another analytics tool
that is not likely to be used by many, we decided
to address the shortcomings of spreadsheets directly
by building a navigation plugin. 
See also our response to {\bf PR1b}.
% In our revised version, 
% we have clearly articulated the difference 
% in goals between Spreadsheets and TDA tools 
% in Section 2
%  and again mention these differences when responding to PR1b. As for comparison with Keshif, the goal for this tool is to auto-generate complex dashboards with customizability built-in and maximize the number of insights gained. However, Keshif cannot relate the summary of a dashboard to the underlying raw data---the lowest level of detail a user can obtain is to view the unique values for a given column and not the the entire data underlying each unique value. Besides Keshif, there are a number of other tools that also support advanced analytical features. However, we have already commented in ``Changes Organized by Section'' under \textbf{S1} and again in \textbf{R(C3)} that such advanced tools have not gained adoption the same way spreadsheets have. Therefore, instead of developing a new tool with a small user-base we opted to address the shortcomings of spreadsheet, a widely used tool.
 }

\stitle{PR1b.} \textit{A strength of NOAH compared to TDA tools is that it's in an actual
    spreadsheet, where it's easy to derive new data based on formulas or expressions,
    but that's not well articulated in the paper and not really part of the study.}

\stitle{Response.} 
\resp{We have already addressed this comment 
in {\bf R(C1)} and {\bf R(C3)}, explaining that 
the goal of spreadsheets 
is to present the raw data as is, 
amenable to editing, formula computation, 
and comparison of derived data, all done in-situ,
unlike TDA tools that hide the data 
while providing summarized statistics
(Section~\ref{sec:related}). 
Our study does explore the derivation of new data---specifically,
the aggregate column feature allows users to derive new 
data via steering tasks, and perform comparisons via 
the compare tasks. 
We evaluate this feature 
both qualitatively and quantitatively 
in the user study results section (Section~\ref{sec:results}).}

\stitle{PR2.} \textit{A design decision that I think is not optimal is the binning of multiple unrelated
items into larger bins, supposedly to ensure scalability (e.g., Ash-Bos). It seems
like an odd choice to group semantically unrelated items (lexicographic ordering
isn't super helpful). It also doesn't seem to help much in terms of scalability, as each individual city is represented as a fairly large bar chart. It would
probably be more effective to just bin separately, which also some participants of the study noted.}

\stitle{Response.} 
\resp{We have addressed this comment in 
in the summary reviews section under the discussion point \textbf{R(C2)} titled \textit{Explanation  of  the  binning  mechanism (C2)}.}

\stitle{PR3.} \textit{By
using Excel vs another tool, it's very obvious to the users which of the tool the experimenter wants to do well, which is a potential source for bias. That could
have potentially been mitigated by having them use the open source spreadsheet tool in both cases (which they are unlikely to be familiar with), which would also
remove some additional confounders.}

\stitle{Response.} \resp{
While we agree with reviewer's comment, 
this bias is unavoidable and is present
in the evaluation of any tool with new features
the participants haven't encountered before
(\ie novelty bias).
The reason we used Excel as opposed to the version
of {\sc DataSpread} without \noah is because
{\sc DataSpread} lacks support for most ``advanced''
spreadsheet operations such as sort, filter,
conditional formatting, pivot table,
and subtotals, limiting the participants' ability
to efficiently complete tasks and negatively
impacting their accuracy and completion times---so it would
not be a fair comparison.
% In fact, one of our goals was to evaluate
% whether 
% the use of these advanced spreadsheet operations
% renders \noah capabilities to be redundant.
}

\stitle{PR4.} \textit{The selection of participants for the study indicates that all were expert
    spreadsheet users. I wonder how NOAH can help non-expert spreadsheet users
    navigate large datasets easily. Another study with users rating their spreadsheet
    skills 2 and above will help answer this question.}

\stitle{Response.} \resp{We have acknowledged this limitation
of our study in our revised submission---see discussion points 
\textbf{R(C4, C5)}. 
The reason we focus on expert users is 
because we did not want expertise
to be the reason for poor task performance, 
and instead wanted
to focus our evaluation on the capabilities
provided by the tools. 
}

\stitle{Minor Comments:}

    * \textit{In Fig 1. the label for breadcrumb covers the actual breadcrumb view. Moving the
    label to the right will make it easy to get a sense of how breadcrumbs look in the
    final tool.}

\stitle{Response.} \resp{In Figure 1, user is at the highest level of the granularity. Therefore, the current path is ``Home''.}

    * \textit{Is it users that rate themselves as greater than 4, or at least 4 in terms of
    skills with spreadsheets?}

\stitle{Response.} \resp{At least 4.}

\subsection*{Secondary Reviewer (SR) review}

\stitle{SR1.} 
\textit{In general, I agree with the motivation and overarching goal of this work: it
    could be useful to enable people to “explore” large tabular data “holistically” by
    showing the data rows and visualizations on the same page with tight coupling
    (i.e., without visual discontinuities in the authors’ term). However, the focus of
    this paper is too narrow and as a result the significance of research contribution
    is too small for a VIS publication.}

\stitle{Response.} 
\resp{As we explained in 
discussion point 
{\bf R(C1)} and changes detailed in {\bf S1}, 
\textit{instead of designing a new sophisticated tool 
from scratch that may cater 
to a small population, 
we aim to enhance the user experience for existing
spreadsheet tools with nearly a billion users}. 
Therefore, the focus of the paper, 
\ie developing a general-purpose plug-in 
to improve navigation within spreadsheets, 
is, on the contrary, 
quite broad and has the 
potential to impact billions of users.
We therefore believe this is a problem 
well-worth investigating for the VIS community.
We have further clarified the scope of tasks 
supported by \noah in changes detailed in {\bf S2},
using Brehmer and Munzner's typology.}

\stitle{SR2.} \textit{As Shneiderman’s mantra---Overview first, zoom and filter, then detail on demand---indicates, seeing an overview is an important and useful step for data exploration. It is a neat idea to provide an overview alongside the table and use the visual elements in the overview as a shortcut to the raw data in the table. However, because this overview is just a histogram for only one data column, even with some additional features, NOAH’s exploration capability seems limited.}

\stitle{Response.} \resp{
\noah does enhance most use cases covered by 
Brehmer and Munzner's typology~\cite{brehmer2013multi}. 
We have addressed 
this comment in changes detailed in \textbf{S2} 
where we identify the scope of tasks supported by \noah.}

\stitle{SR3.} \textit{I note that the comparative study is designed to evaluate whether NOAH helps
    address spreadsheet “navigational challenges” instead of demonstrating NOAH’s
    value in facilitating holistic data exploration. It seems that the six quiz tasks
    are somewhat arbitrary or even biased to emphasize the navigational issue of
    Excel: I do agree that Excel’s navigation is not ideal.}

\stitle{Response.} 
\resp{We have addressed this comment in the discussion point titled \textbf{R(C4, C5)}. We now explain the quiz tasks in the context
of the typology discussed previously, while also providing
a detailed end-to-end motivating scenario that captures
the quiz tasks.}

\stitle{SR4.} \textit{For the Steer and
    Customize tasks, why do the quiz tasks “instruct” participants to organize the
    data by some attributes? For the ``find'' task, why is it asking participants to
    answer the question from “the currently visible spreadsheet window?”}

\stitle{Response.} \resp{Both these tasks 
simulate the scenario where a spreadsheet user 
chooses to navigate the data based on 
a specific attribute, as a first step 
towards answering a specific question regarding the dataset.
In the revised version, 
we provide more details as to why the 
quiz tasks were selected in the context of an 
end-to-end motivating example scenario in Section~\ref{sec:study}.}

\stitle{SR5.} \textit{More
    importantly, people with a basic level of expertise can easily answer these tasks
    using existing BI tools such as Tableau, Pivot Table, and Power BI. 
    I know that, as
    the authors pointed out, these tabular data analysis tools do not show the tabular
    data and visualizations together on the same sheet.}

\stitle{Response.} 
\resp{We address this comment 
in the changes detailed in point {\bf S1},
and explain how spreadsheets cater to a 
much larger user-base than the aforementioned tools.
In fact, users typically shun these ``advanced'' BI tools
and instead use spreadsheets. 
So, it is incumbent upon us to improve their experience
when using spreadsheets, rather than forcing
them to adopt a new tool. 
We further explain how the goals of BI tools 
and spreadsheets are different 
in discussion point \textbf{R(C3)}.}

\stitle{SR6.} \textit{However, Excel is not designed to best support navigation and instead it provides
    additional flexible features that the authors acknowledge as limitations. To me,
    it is unfair to choose the tool with a limitation as a baseline (excluding other
    more power existing tools) along with some questionable tasks, and show that NOAH
    is better than the baseline for the tasks demonstrate the limitation. Therefore,
    the study results are neither interesting nor informative.}

\stitle{Response.} 
\resp{See our response to {\bf SR1}, {\bf SR2}, {\bf SR3}, and {\bf SR5}.
We now justify our quiz tasks better and 
discuss the motivation for comparing
against Excel. 
We would argue that understanding the user experience
for a tool that approximately 10\% of the planet uses
is well worth the effort---and our study results will help inform
the development of future tools that eliminate barriers
in spreadsheet data analysis on increasingly large datasets.}

\stitle{SR7.} \textit{I am not sure
    why the authors wanted to treat the dataset as a main factor when the four main
    research questions do not involve dataset characteristics? I expect the
    explanation of the comparison of dataset characteristics and the discussion of
    interaction effect, which would make it worth introducing it as a main factor
    instead of running the study as a simple A/B comparison.}

\stitle{Response.} \resp{We 
address this comment in the discussion point \textbf{R(C4, C5)}, 
where we explain how, in our revised version,
we now clearly articulate the 
goals of the study while restructuring our research questions. 
Moreover, we have completely 
revamped Section 7 (the results section) 
to better reflect
our  research  questions,  
while  adding  more  qualitative  observations  
regarding  the  user  experience  with \noah.}

\stitle{SR8.} \textit{While I appreciate the authors’ effort on conducting a comparative study, it would
    have been a lot more interesting and informative if the authors conducted a
    qualitative study with high-level questions (which can be combined with a few
    meaningful low-level ones) that would encourage participants to utilize NOAH’s
    most novel features to see *IF* and *HOW* people use NOAH in their data
    exploration process.}

\stitle{Response.} \resp{As explained in 
the changes detailed in point {\bf S7}, 
in the revised paper, we have added more qualitative observations 
regarding the user experience with \noah in Section 7.2.
Overall, Section 7 has been completely restructured
and rewritten. }

\stitle{SR9.} \textit{Even though the authors described their design considerations, I have some
    doubts about NOAH’s design. For example, despite the authors’ justification on the
    overview construction process using the common summarization technique, the number
    of items in each group seems small to me. Also, it is weird to see the different
    bar height between different groups when they have different number of cities and
    to see the completely different representation when a group has only one city
    (Fig. 2).}

\stitle{Response.} \resp{See discussion point \textbf{R(C2)}.}

\subsection*{External Reviewer 1 (ERA) review}
\stitle{ERA1.} \textit{I have little comments on the first part of the paper, which I find very well
    executed. The comments I have are: What about DC2 in the implementation section?}

\stitle{Response.} \resp{We address DC2 by 
introducing the Aggregate column feature (Section 5.2).}

\stitle{ERA2.} \textit{I wonder whether the authors considered more direct interactions for some of the
    operations. For example, one could use Embedded Merge and Split technique to
    facilitate merging and splitting bins in histograms [A].}

 \stitle{Response.} \resp{We have added this citation [66] to motivate future work on improving the bin customization feature (Section 8.2).}

\stitle{ERA3.} \textit{I would have liked to see more explanations regarding how the four research
    questions were selected. Why these ones? Are there others that were considered?}

\stitle{Response.} \resp{We have addressed this comment in 
discussion point \textbf{R(C4, C5)}.}

\stitle{ERA4.} \textit{The input method for participants to specify their answers is not described. Not
    that it is a big problem, because task completion time is several dozens of
    seconds, but still it should be specified.}

\stitle{Response.} \resp{We had already discussed the procedure in the original submission in Phase 2 of Section 6.2 \textit{We developed an online JavaScript-based quiz system that recorded user responses and submission times. 
We also recorded the participants' interactions
with both tools using screen capture software}.}

\stitle{ERA5.} \textit{I think the fact that participants subjectively indicated preferring NOAH over Excel should be tone-downed a bit. Because it was clear to them what the
    experimenter's solution is (NOAH), they were largely sensitive to bias in
    assessing the two different conditions.}

\stitle{Response.} \resp{We have revised the paper to acknowledge
the issue of bias; also see response to primary reviewer comment {\bf PR3}. 
We provide qualitative insights to highlight the pros and cons of using \noah for spreadsheet navigation in Section~\ref{sec:results},
and have de-emphasized  
the quantitative insights that show \noah outperforming Excel.}

\stitle{ERA6.} \textit{Given the within-subject design, and given that the authors assume that the two datasets lead to comparable results, I was expecting to see intra-participant
    differences being reported. Still, this is not a deal-breaker as the authors report the results separately for each dataset.}

\stitle{Response.} \resp{We have addressed this comment in the
discussion point \textbf{R(C6)} 
titled \textit{Reporting Intra-participant Differences (C6)}.}

\stitle{ERA7.} \textit{I find the choice of datasets a bit limiting. If the real challenge is the
    difficulty to navigate in very large datasets (the authors mention several
    millions of cells in the paper), then why selecting datasets with only 14 or 15 columns? This easily fits on a single screen and only requires scrolling
    vertically. This should at least be discussed in the discussion (the experiment targets scalability in terms of rows, but not in terms of columns).}

\stitle{Response.} 
\resp{
%We didn't discuss this. Was trying to avoid providing a direct answer. We can say the following:
We acknowledge the reviewer's comment. In the revised submission,
we discuss how future iterations of \noah can support bespoke user-defined customizations of the overview (Section~\ref{sec:discuss_additionational}). This may be as simple as defining the orientation of the overview to selecting the binning criteria.}

\subsection*{External Reviewer 2 (ERB) review}

\stitle{ERB1.} \textit{The title says “At Scale”  but there is no discussion of scalability. Does it work
    with the 2M cells mentioned in the intro?  No test with different sizes of large
    datasets.  Nothing is done or said to convince me that the testing warrants saying
    “at scale” in the title of the paper.
    REQUIRED: remove or justify.}

\stitle{Response.} \resp{Overview histogram construction 
in \noah is as expensive as the spreadsheet sort operation. The only other computational operation is formulae calculation for the aggregate column operation which leverages the formula engine of {\sc DataSpread}. Therefore, \noah is as fast as the underlying spreadsheet system. We provide the implementation details in the Appendix.}

\stitle{ERB2.} \textit{We usually talk about scrolling and zooming for navigation.  Is steering the same
    as zooming?  Or is steering Sorting and/or Filtering?  Something else?  The paper
    needs to clarify what operations are actually included in the “steering”.}

\stitle{Response.} \resp{We have better articulated the 
definition of these operations with examples in our revised 
Section 1. (Also see summary of changes point {\bf S1}.)}

\stitle{ERB3.} \textit{I found the “Ashville-Boston bin” confusing…. Why were several cities binned
    together by default? \ldots I really do not understand the benefit of binning cities i.e. categorical
    variables… The individual bars for each city are shown anyway so adding binning in
    the overview only increases the length of the overview (and the need for
    scrolling) so why not just show a bar chart with one bar for each city first, and
    then allow binning only as needed?
    REQUIRED: Clarify the binning by default of categorical variables.}

\stitle{Response.} \resp{We have addressed this comment 
in discussion point \textbf{R(C2)} 
titled \textit{Explanation  of  the  binning  mechanism (C2)}.}

\stitle{ERB4.} \textit{The contextual and historical information are not particularly novel but are
    appropriately included (and authors should be commended for doing so).   Can the
    paper clarify how the history is kept?  E.g. keeping the complete state info or
    recompute from start? Any challenges worth reporting?}

\stitle{Response.} \resp{As mentioned in Section 5.3, \noah simply documents each new bin that the user navigates to in the context bar. \noah preserves the complete state information as explained in the final paragraph of Section 5.2: \textit{we note that the aggregate column
is kept in sync with the bins
as users zoom in and out,
eliminating repeated steering operations}.}

\stitle{ERB5.} \textit{I am not sure I really understand the difference with SUBTOTAL. Not clear…
    The paper says: “the size of the summary itself can become very large and
    span multiple screens, e.g., for numeric data. “ is it because there is no
    binning? Or because too many attributes are summarized in extra columns with a
    visual representation (and it add too many columns)?   Not clear.   Can the paper
    also clarify what level of coordination there is in Subtotal? In short: is the
    difference a fundamental difference or only a superficial layout one?
    REQUIRED: Clarify difference with Subtotal.}

\stitle{Response.} 
% \resp{OLD:} \newsaj{We have further clarified how the Subtotal feature operates in Section 3.1. Subtotal \agp{This is not a sentence} similar limitations regarding navigation as Pivot Tables: a) unlike \noah, the summary groups in Subtotal are not binned and impact visual continuity when there are too many groups to explore, b) the summaries are inserted as a collection of new rows within the spreadsheet and lack coordination with the raw data. } \resp{NEW: } 
\resp{We have further clarified how the Subtotal feature operates in Section 3.1. 
The difference between Subtotal and 
\noah's overview are: 
a) the summary groups within 
Subtotal are not binned like in \noah,
resulting in visual discontinuity when
there are too many groups to explore and
b) the summaries are inserted as a collection of new rows within the spreadsheet 
rather than as a separate plugin like \noah 
and lack 
coordination with the raw data.}

\stitle{ERB6.} \textit{In 5.1.1
    The paper says that clicking is not filtering but what is it?  A novel sort?
    Describe the algorithm used E.g. When it says “as the user clicks on the bin, the
    data corresponding to Boston is displayed”: what really is done on the table side?
    REQUIRED: clarify operations done in the detail view}

\stitle{Response.} 
%\resp{OLD: The overview-spreadsheet coordination is discussed in Section 5.1.2. \agp{Is this revised? They say this is required. Say that something is done in the new version.} }
%\resp{NEW:} 
\resp{The overview-spreadsheet coordination is discussed in Section \ref{sec:design_coord}.  In the revised submission, we discuss the implementation details in Appendix and mention that \noah leverages {\sc DataSpread}'s built-in positional representation of the data to enable such coordination.}

\stitle{ERB7.} \textit{In general: 
    I would like to see at least one example of the use of a continuous attribute e.g.
    numerical or date .
    Aren’t they some drawbacks in allowing only a single overview attribute?  E..g
    doesn’t it make it harder to see correlations between attributes?}

\stitle{Response.} \resp{The customize task required participants to navigate a numeric attribute. Moreover, the attached video demo in the supplemental material also demonstrates how this task is performed for numeric data at 3 minutes 44 seconds.}

\stitle{ERB8.} \textit{User study: 
    It looks like the user study had no task using the details part of the
    overview+details (while coordination is a main contribution).  Strange.}

\stitle{Response.} \resp{The find task is on the detailed view and asks the participants to relate the results of the aggregate column with the raw spreadsheet data.}

\stitle{ERB9.} \textit{While I understand that it is easier to compare a new tool to Excel I wondered why
    it was not compared with Subtotal, which seems to be a close state of the art.}

\stitle{Response.} 
%\resp{OLD:} \newsaj{Participants were free to use any feature they were comfortable with while completing the quiz tasks. However, the Subtotal feature is not very effective for a number of the quiz tasks, \eg both steer tasks (due to not supporting COUNTIF), find and compare (for not \agp{maybe start with due to for consistency} being able to perform the steer task\agp{?}), customize (due to a lack of customization functionality).}
\resp{Participants were free to use any feature they were comfortable with while completing the quiz tasks, and therefore, could use subtotals if they wished. However, the subtotal feature is not very effective for a number of the quiz tasks, \eg both steer tasks (due to not supporting COUNTIF) and customize (due to a lack of customization functionality). Moreover, the find and compare tasks were dependant on the results of the steer tasks, and, therefore, could not 
be performed without completing the steer
tasks first.}

\stitle{ERB10.} \textit{In
    the current study we have so many things changing between the 2 interfaces that we
    will never know what really helped the most: it is the presence of an overview? or
    the bin mechanism which reduces overview scrolling? or fact that only a single
    overview can be seen at once?, or the tight coupling between details and overview?
    I cannot tell…}

\stitle{Response.} \resp{We acknowledge this as a limitation of our study. In the revision, we have revamped Section 8 and discussed a number of limitations, including the one mentioned by the reviewer, in Section 8.1 under \textit{Isolated evaluation of \noah components.}. See also summary of changes point {\bf S8}.}

\stitle{ERB11.} \textit{If all the tasks “only” involve counts (and no need to see the details) then a
    better candidate for comparison may be Keshif, where (if I remember correctly) all
    count distributions are shown by default for all attributes making easy to find
    counts, min and max.  Comparisons are possible between 2 selections.}

\stitle{Response.} 
\resp{Please see our response to \textbf{PR1a} where we explain our reasoning for not considering Keshif. 
Our tasks went beyond just counts; for example,
the customize task required users to invoke an average formula. In our revised submission, we mention the scope of formulae supported by the aggregate column feature in Section~\ref{sec:agg_col}: \textit{``the aggregate column can employ any statistical
or mathematical formulae that
operate over a range of data''}.}

### 2020-03-10 discussion.tex
Updates from Overleaf

@@
 additional details on demand via a multi-granularity overview, as well as employ
-aggregattion in-situ, which eliminates cumbersome steering operations.
+aggregation
+in-situ, which eliminates cumbersome steering operations.
 Quantitatively, we find that \noah speeds up navigation without compromising accuracy.
@@
 \subsection{Additional enhancements}
+\label{sec:discuss_additionational}
 We now discuss other functionalities that can be introduced to further enhance the capabilities of \noah.

### 2020-03-10 related.tex
s1-3 and letter comments
final tvcg
Updates from Overleaf

@@
 compare subsets of data without having to 
-provides a summary view, enabling users
-to compare subsets of data without
-having to navigate to various locations
+navigate to various locations
 within the sheet.
@@
 auto-generating charts of data statistics.
-Users can specify queries to the system
+Users can specify queries to the tool
 (similar to a web search)
@@
 lose the ability to directly manipulate raw data, derive new data, and issue formulae for free-form analysis.  
-\cut{For example, using these tools users cannot \code{browse} subsets of data and \code{lookup} or \code{locate} data points of interest (see Table~\ref{tab:scope}).
-Aditya: previous line problematic: they can...
-because all these tools support to varying extents the typology.
-Even users of such tools want to inspect the underlying data to verify the summary statistics and visualizations~\cite{2017trust}.
-Aditya:this line is problematic too...
-because the citation is to an AQP tool rather than some general purpose study}
 Therefore, the goals of spreadsheets differ from TDA 
@@
 using various operations involving navigation,
-\eg formulae or expressions\agp{expressions?}.
+\eg issuing formulae.
 \noah being a plug-in to spreadsheets, 
